---
title: "Analyses"
format: html
editor: visual
---

**Response variables:**

-   New_nests_os: number of new osmia nests for each trapnest and date combination.
-   New_nests_meg: number of new megachile nests for each trapnest and date combination.
-   Prop_filled: proportion of the straw filled with cells, used as a measure of investment into a nest.

**Predictor variables:**

-   Available_cavities_prior: number of available cavities that could have had a new nest built between visits.
-   Osmia_ps_prior: number of existing osmia in each trapnest per date, including pseudo nests.
-   Osmia_no_ps_prior: number of existing osmia in each trapnest per date, excluding pseudo nests.
-   Meg_ps_prior: number of existing megachile in each trapnest per date, including pseudo nests.
-   Meg_no_ps_prior: number of existing megachile in each trapnest per date, excluding pseudo nests.
-   Par_in_trapnest: number of parasites present in the trapnest when I log new bee nests (Bee_choice). \#######

**Additional variables:**

-   Good_days: number of days between visits with good enough weather for bee activity.
-   Day_of_year: the day of the year, a stand in for date.
-   Site: three sites in total.
-   Subsite: six subsites for each site, unique to each site.
-   Trapnest: 72 trapnests total, ach subsite contains three, unique to each subsite.

```{r Prep, include=FALSE}

## Loading data and packages

library(glmmTMB)
library(DHARMa)
library(lme4)
library(car)
library(see)
library(performance)
library(sjPlot) # for plot_model
library(ggplot2)
library(ggeffects)
library(effects)
library(tidyverse)
library(interactions) # for interact_plot
library(betareg)

nesting <- read_csv('data/nesting_cleaned_all_bees.csv')
# nesting_os <- read_csv('data/nesting_cleaned_osmia.csv')
# nesting_meg <- read_csv('data/nesting_cleaned_megachile.csv') # three columns with parasite and cell info blank, not dissected
nesting_os_xs <- read_csv('data/nesting_cleaned_osmia_xs.csv')
nesting_meg_xs <- read_csv('data/nesting_cleaned_megachile_xs.csv')
nesting_cells_os <- read_csv('data/nesting_cells_cleaned_osia.csv')
nesting_investment <- read_csv('data/investment_cleaned.csv')
investment_os <- nesting_investment %>% filter(Occupant == 'osmia') 

nesting_cells_os$Last_cell <- as.factor(nesting_cells_os$Last_cell)


```

### Nesting per Available Cavities

```{r Models: Nest per Available Cavities, include=TRUE}


# MODEL 1 
### New nests per available cavities Osmia ###
mod_choice_cav_os <- glmer(New_nests_os ~ scale(Available_cavities_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_os_xs, family = poisson()) 

plot(simulateResiduals(mod_choice_cav_os)) # great
check_model(mod_choice_cav_os) # fine
testDispersion(mod_choice_cav_os, type = 'DHARMa') # good
testZeroInflation(mod_choice_cav_os) # good
summary(mod_choice_cav_os) # Site significant
 # anova to see site differences

# Plot model
ggpredict(mod_choice_cav_os, c('Available_cavities_prior')) %>% plot(rawdata = TRUE, jitter = .04)


# MODEL 2
### New nests per available cavities Megachile ###
mod_choice_cav_meg <- glmer(New_nests_meg ~ scale(Available_cavities_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_meg_xs, family = poisson()) 

plot(simulateResiduals(mod_choice_cav_meg)) # good
testDispersion(mod_choice_cav_meg, type = 'DHARMa') # fine
testZeroInflation(mod_choice_cav_meg) # good
check_model(mod_choice_cav_meg) # fine
summary(mod_choice_cav_meg) # Day of year and Site significant
# anova

# Plot model
ggpredict(mod_choice_cav_os, c('Available_cavities_prior')) %>% plot(rawdata = TRUE, jitter = .04) 

```

### Nesting per Conspecific Presence

```{r Models: Nest per Conspecifics Osmia, include=TRUE}


# MODEL 3 
### New nests per conspecifics Osmia, with pseudo ###

mod_choice_consp_os_ps <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Osmia_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_os_xs, family = binomial) # ignore boundary singular warning, means little to no influence of random effects

plot(simulateResiduals(mod_choice_consp_os_ps)) # fine
check_model(mod_choice_consp_os_ps) # fine
testDispersion(mod_choice_consp_os_ps, type = 'DHARMa')  
testZeroInflation(mod_choice_consp_os_ps) # zero-inflation, but not detrimental
summary(mod_choice_consp_os_ps) # more osmia nests when there's higher osmia densities


ggpredict(mod_choice_consp_os_ps, c('Osmia_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 


# MODEL 4 
### New nests per conspecifics Osmia, without pseudo ###

mod_choice_consp_os_no_ps_BINOMIAL <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Trapnest), data = nesting_os_xs, family = binomial(),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=1000000)))

mod_choice_consp_os_no_ps_POISSON <- glmer(New_nests_os ~ offset(log(Available_cavities_prior - New_nests_sum))+ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Trapnest), data = nesting_os_xs, family = poisson())

mod_choice_consp_os_no_ps_POISSON_subsite <- glmer(New_nests_os ~ offset(log(Available_cavities_prior - New_nests_sum))+ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite) + (1|Trapnest), data = nesting_os_xs, family = poisson(), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=1000000)))


plot(simulateResiduals(mod_choice_consp_os_no_ps_BINOMIAL)) # 
plot(simulateResiduals(mod_choice_consp_os_no_ps_POISSON)) # 
plot(simulateResiduals(mod_choice_consp_os_no_ps_POISSON_subsite)) # 

summary(mod_choice_consp_os_no_ps_BINOMIAL)
summary(mod_choice_consp_os_no_ps_POISSON)
summary(mod_choice_consp_os_no_ps_POISSON_subsite)

ggpredict(mod_choice_consp_os_no_ps_BINOMIAL, c('Osmia_no_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 
ggpredict(mod_choice_consp_os_no_ps_POISSON, c('Osmia_no_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 
ggpredict(mod_choice_consp_os_no_ps_POISSON_subsite, c('Osmia_no_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 



plot(simulateResiduals(mod_choice_consp_os_no_ps)) # fine
check_model(mod_choice_consp_os_no_ps) # good
testDispersion(mod_choice_consp_os_no_ps, type = 'DHARMa') # some overdispersion, not detrimental
testZeroInflation(mod_choice_consp_os_no_ps) # some zero inflation
summary(mod_choice_consp_os_no_ps) #### HELP
# anova

ggpredict(mod_choice_consp_os_no_ps_BINOMIAL, c('Osmia_no_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 
ggpredict(mod_choice_consp_os_no_ps_POISSON, c('Osmia_no_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 







plot(allEffects(mod_choice_consp_os_no_ps))

```

------------------------------------------------------------------------

```{r Models: Nest per Conspecifics Megachile}



# MODEL 5
### New nests per conspecifics Megachile, with pseudo ###
mod_choice_consp_meg_ps <- glmer(cbind(New_nests_meg, Available_cavities_prior - New_nests_sum) ~ scale(Meg_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_meg_xs, family = binomial())

plot(simulateResiduals(mod_choice_consp_meg_ps)) # great
check_model(mod_choice_consp_meg_ps) # good
testDispersion(mod_choice_consp_meg_ps, type = 'DHARMa') # good
testZeroInflation(mod_choice_consp_meg_ps) # good
summary(mod_choice_consp_meg_ps) # Day of year and Site significant
save(mod_choice_consp_meg_ps, file='Models/mod_choice_consp_meg_ps.Rdata')

ggpredict(mod_choice_consp_meg_ps, c('Meg_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 


# MODEL 6
### New nests per conspecifics Megachile, without pseudo ###
mod_choice_consp_meg_no_ps_BINOMIAL <- glmer(cbind(New_nests_meg, Available_cavities_prior - New_nests_sum) ~ scale(Meg_no_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_meg_xs, family = binomial())

mod_choice_consp_meg_no_ps_POISSON <- glmer(New_nests_meg ~ offset(log(Available_cavities_prior - New_nests_sum))+ scale(Meg_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Trapnest), data = nesting_os_xs, family = poisson())

plot(simulateResiduals(mod_choice_consp_meg_no_ps_BINOMIAL))
plot(simulateResiduals(mod_choice_consp_meg_no_ps_POISSON))

summary(mod_choice_consp_meg_no_ps_BINOMIAL) 
summary(mod_choice_consp_meg_no_ps_POISSON) 


plot(simulateResiduals(mod_choice_consp_meg_no_ps)) # great
check_model(mod_choice_consp_meg_no_ps) # fine
testDispersion(mod_choice_consp_meg_no_ps, type = 'DHARMa') # good 
testZeroInflation(mod_choice_consp_meg_no_ps) # good
summary(mod_choice_consp_meg_no_ps) # Day of year and Site significant
save(mod_choice_consp_meg_no_ps, file='Models/mod_choice_consp_meg_no_ps.Rdata')

ggpredict(mod_choice_consp_meg_no_ps, c('Meg_no_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 


```

### Parasitism

```{r Parasites and Aggregations, include=TRUE}

# Relation between parasites and aggregation densities

# MODEL 7
# Relation between parasites and aggregation densities, with pseudo

mod_par_agg_ps <- glmer(Cell_parasitized ~ scale(Osmia_ps_no_iw) + scale(Day_of_year) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_ps)) # good
check_model(mod_par_agg_ps) # fine
testDispersion(mod_par_agg_ps, type = 'DHARMa') # good
testZeroInflation(mod_par_agg_ps) # good
summary(mod_par_agg_ps) # no detected effects

ggpredict(mod_par_agg_ps, c('Osmia_ps_no_iw')) %>% plot(rawdata = TRUE, jitter = .04) 


# MODEL 8
# Relation between parasites and aggregation densities, without pseudo

mod_par_agg_no_ps <- glmer(Cell_parasitized ~ scale(Osmia_no_ps_no_iw) + scale(Day_of_year) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_no_ps)) # good
check_model(mod_par_agg_no_ps) # fine
testDispersion(mod_par_agg_no_ps, type = 'DHARMa') # good
testZeroInflation(mod_par_agg_no_ps) # good
summary(mod_par_agg_no_ps) # no detected effects

ggpredict(mod_par_agg_no_ps, c('Osmia_no_ps_no_iw')) %>% plot(rawdata = TRUE, jitter = .04) 


# MODEL 9
# Relation between parasites and aggregation densities, unsealed

mod_par_agg_unsealed_os <- glmer(Cell_parasitized ~ scale(Sum_unsealed_os_no_iw) + scale(Day_of_year) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_unsealed_os)) # good
check_model(mod_par_agg_unsealed_os) # fine
testDispersion(mod_par_agg_unsealed_os, type = 'DHARMa') # good
testZeroInflation(mod_par_agg_unsealed_os) # good
summary(mod_par_agg_unsealed_os) 

ggpredict(mod_par_agg_unsealed_os, c('Sum_unsealed_os_no_iw')) %>% plot(rawdata = TRUE, jitter = .04) 

```

```{r Preferences and Parasites, include=TRUE}


# MODEL 10
# How the presence of parasites effects nesting preferences, with pseudo

mod_choice_par_ps_os <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Par_nests_unsealed_os) * scale(Osmia_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting, family = binomial, control = glmerControl(optimizer ="nloptwrap")) ### MODEL FAILED TO CONVERGE


plot(simulateResiduals(mod_choice_par_ps_os)) # help
check_model(mod_choice_par_ps_os) # HELP 
testDispersion(mod_choice_par_ps_os, type = 'DHARMa') # 
testZeroInflation(mod_choice_par_ps_os) # 
summary(mod_choice_par_ps_os) #  almost everything sig, how to interpret interaction?

interact_plot(mod_choice_par_ps_os, pred = Osmia_ps_prior, modx = Par_nests_unsealed_os, data = nesting)


# Model 11
# How the presence of parasites effects nesting preferences, without pseudo

mod_choice_par_no_ps_os <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Par_nests_unsealed_os) * scale(Osmia_no_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting, family = binomial)

plot(simulateResiduals(mod_choice_par_no_ps_os)) # 
check_model(mod_choice_par_no_ps_os) # HELP  
testDispersion(mod_choice_par_no_ps_os, type = 'DHARMa') # fine
testZeroInflation(mod_choice_par_no_ps_os) # fine
summary(mod_choice_par_no_ps_os) # almost everything sig

interact_plot(mod_choice_par_no_ps_os, pred = Osmia_no_ps_prior, modx = Par_nests_unsealed_os, data = nesting)

```

```{r Investment, include=TRUE}


# MODEL 12
# How the presence of parasites effects nest investment
# Likelyhood of  of terminating nest investment after a cell was parasitized
# position of cell (last or not) ~ if cell is parasitized + day of year + site + (1|Subsite/Trapnest/Nest_ID)

mod_last_cell <- glmer(Last_cell ~ Cell_parasitized + Day_of_year + Site + (1|Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_last_cell)) # good
# check_model(mod_last_cell) # HELP doesnt work
summary(mod_last_cell) # site and day of year

ggpredict(mod_last_cell, c('Cell_parasitized')) %>% plot(rawdata = TRUE, jitter = .04) 




plot(I(New_nests_os/ (Available_cavities_prior)) ~ (Day_of_year), data=nesting)
plot(I(New_nests_os/ (Available_cavities_prior)) ~ (Day_of_year), data=nesting)

mod_choice_par_ps_os <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Par_nests_unsealed_os) * scale(Osmia_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting, family = binomial, control = glmerControl(optimizer ="nloptwrap"))








# MODEL 13
# Investment into a nest if other active nests or the nest itself is parasitized

mod_invest_filled <- glmer(cbind(Filled_length, Straw_length) ~ scale(Par_nests_unsealed_os) * Nest_parasitized + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = investment_os, family = binomial)

## MAKING PROP
investment_os <- investment_os %>% mutate(Prop_nest_filled = Filled_length_mm/Straw_length_mm)

# mod_invest_filled_2 <- glmer(Prop_nest_filled ~ scale(Par_nests_unsealed_os) * Nest_parasitized + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = investment_os, family = ordbeta(link = "logit"))

plot(simulateResiduals(mod_invest_filled)) # 
check_model(mod_invest_filled) # 
summary(mod_invest_filled) # 

ggpredict(mod_invest_filled, c('Par_nests_unsealed_os')) %>% plot(rawdata = TRUE, jitter = .04) 


```
