---
title: "Analyses"
format: html
editor: visual
---

**Response variables:**

-   New_nests_os: number of new osmia nests for each trapnest and date combination.
-   New_nests_meg: number of new megachile nests for each trapnest and date combination.
-   Prop_filled: proportion of the straw filled with cells, used as a measure of investment into a nest.

**Predictor variables:**

-   Available_cavities_prior: number of available cavities that could have had a new nest built between visits.
-   Osmia_ps_prior: number of existing osmia in each trapnest per date, including pseudo nests.
-   Osmia_no_ps_prior: number of existing osmia in each trapnest per date, excluding pseudo nests.
-   Meg_ps_prior: number of existing megachile in each trapnest per date, including pseudo nests.
-   Meg_no_ps_prior: number of existing megachile in each trapnest per date, excluding pseudo nests.
-   Par_in_trapnest: number of parasites present in the trapnest when I log new bee nests (Bee_choice). \#######

**Additional variables:**

-   Good_days: number of days between visits with good enough weather for bee activity.
-   Day_of_year: the day of the year, a stand in for date.
-   Site: three sites in total.
-   Subsite: six subsites for each site, unique to each site.
-   Trapnest: 72 trapnests total, ach subsite contains three, unique to each subsite.

```{r Prep, include=FALSE}

## Loading data and packages

library(glmmTMB)
library(DHARMa)
library(lme4)
library(car)
library(see)
library(performance)
library(sjPlot) # for plot_model
library(ggplot2)
library(ggeffects)
library(effects)
library(tidyverse)
library(interactions) # for interact_plot

# nesting <- read_csv('data/nesting_cleaned_all_bees.csv')
# nesting_os <- read_csv('data/nesting_cleaned_osmia.csv')
# nesting_meg <- read_csv('data/nesting_cleaned_megachile.csv') # three columns with parasite and cell info blank, not dissected
nesting_os_xs <- read_csv('data/nesting_cleaned_osmia_xs.csv')
nesting_meg_xs <- read_csv('data/nesting_cleaned_megachile_xs.csv')
investment <- read_csv('data/investment_cleaned.csv')
investment_os <- investment %>% filter(Occupant == 'osmia')

```

### Nesting per Available Cavities

```{r Models: Nest per Available Cavities, include=FALSE}


# MODEL 1 
### New nests per available cavities Osmia ###
mod_choice_cav_os <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Available_cavities_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_os_xs, family = binomial) 

plot(simulateResiduals(mod_choice_cav_os)) # great
check_model(mod_choice_cav_os) # fine
testDispersion(mod_choice_cav_os, type = 'DHARMa') # good
testZeroInflation(mod_choice_cav_os) # good
summary(mod_choice_cav_os) # Site significant
 # anova to see site differences
save(mod_choice_cav_os, file='Models/mod_choice_cav_os.Rdata') 

# Plot model
ggpredict(mod_choice_cav_os, c('Available_cavities_prior')) %>% plot(rawdata = TRUE, jitter = .04)


# MODEL 2
### New nests per available cavities Megachile ###
mod_choice_cav_meg <- glmer(cbind(New_nests_meg, Available_cavities_prior - New_nests_sum)  ~ scale(Available_cavities_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_meg_xs, family = binomial) 

plot(simulateResiduals(mod_choice_cav_meg)) # great
testDispersion(mod_choice_cav_meg, type = 'DHARMa') # good
testZeroInflation(mod_choice_cav_meg) # good
check_model(mod_choice_cav_meg) # fine
summary(mod_choice_cav_meg) # Day of year and Site significant
# anova
save(mod_choice_cav_meg, file='Models/mod_choice_cav_meg.Rdata')

# Plot model
ggpredict(mod_choice_cav_os, c('Available_cavities_prior')) %>% plot(rawdata = TRUE, jitter = .04) # + labs(x = 'Available cavities', y = 'Presence of new nest for each visit')

```

### Nesting per Conspecific Presence

```{r Models: Nest per Conspecifics Osmia, include=FALSE}


# MODEL 3 
### New nests per conspecifics Osmia, with pseudo ###

mod_choice_consp_os_ps <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Osmia_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_os_xs, family = binomial) # ignore boundary singular warning, means little to no influence of random effects

plot(simulateResiduals(mod_choice_consp_os_ps)) # fine
check_model(mod_choice_consp_os_ps) # fine
testDispersion(mod_choice_consp_os_ps, type = 'DHARMa')  
testZeroInflation(mod_choice_consp_os_ps) # zero-inflation, but not detrimental
summary(mod_choice_consp_os_ps) # more osmia nests when there's higher osmia densities
save(mod_choice_consp_os_ps, file='Models/mod_choice_consp_os_ps.Rdata')


ggpredict(mod_choice_consp_os_ps, c('Osmia_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 


# MODEL 4 
### New nests per conspecifics Osmia, without pseudo ###

mod_choice_consp_os_no_ps <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Osmia_no_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_os_xs, family = binomial())

plot(simulateResiduals(mod_choice_consp_os_no_ps)) # fine
check_model(mod_choice_consp_os_no_ps) # good
testDispersion(mod_choice_consp_os_no_ps, type = 'DHARMa') # some overdispersion, not detrimental
testZeroInflation(mod_choice_consp_os_no_ps) # some zero inflation
summary(mod_choice_consp_os_no_ps) #### HELP
# anova
save(mod_choice_consp_os_no_ps, file='Models/mod_choice_consp_os_no_ps.Rdata')


ggpredict(mod_choice_consp_os_no_ps, c('Osmia_no_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 
####### HELP





```

------------------------------------------------------------------------

```{r Models: Nest per Conspecifics Megachile}



# MODEL 5
### New nests per conspecifics Megachile, with pseudo ###
mod_choice_consp_meg_ps <- glmer(cbind(New_nests_meg, Available_cavities_prior - New_nests_sum) ~ scale(Meg_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_meg_xs, family = binomial())

plot(simulateResiduals(mod_choice_consp_meg_ps)) # great
check_model(mod_choice_consp_meg_ps) # good
testDispersion(mod_choice_consp_meg_ps, type = 'DHARMa') # good
testZeroInflation(mod_choice_consp_meg_ps) # good
summary(mod_choice_consp_meg_ps) # Day of year and Site significant
save(mod_choice_consp_meg_ps, file='Models/mod_choice_consp_meg_ps.Rdata')

ggpredict(mod_choice_consp_meg_ps, c('Meg_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 


# MODEL 6
### New nests per conspecifics Megachile, without pseudo ###
mod_choice_consp_meg_no_ps <- glmer(cbind(New_nests_meg, Available_cavities_prior - New_nests_sum) ~ scale(Meg_no_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_meg_xs, family = binomial())

plot(simulateResiduals(mod_choice_consp_meg_no_ps)) # great
check_model(mod_choice_consp_meg_no_ps) # fine
testDispersion(mod_choice_consp_meg_no_ps, type = 'DHARMa') # good 
testZeroInflation(mod_choice_consp_meg_no_ps) # good
summary(mod_choice_consp_meg_no_ps) # Day of year and Site significant
save(mod_choice_consp_meg_no_ps, file='Models/mod_choice_consp_meg_no_ps.Rdata')

ggpredict(mod_choice_consp_meg_no_ps, c('Meg_no_ps_prior')) %>% plot(rawdata = TRUE, jitter = .04) 


```

### Parasitism

```{r Parasites and Aggregations, include=FALSE}

# Relation between parasites and aggregation densities

# MODEL 7
# Relation between parasites and aggregation densities, with pseudo

mod_par_agg_ps <- glmer(Cell_parasitized ~ scale(Osmia_ps_post) + scale(Day_of_year) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_ps)) # 
check_model(mod_par_agg_ps) # 
testDispersion(mod_par_agg_ps, type = 'DHARMa') #  
testZeroInflation(mod_par_agg_ps) # 
summary(mod_par_agg_ps) # 

ggpredict(mod_par_agg_ps, c('Osmia_ps_post')) %>% plot(rawdata = TRUE, jitter = .04) 


# MODEL 8
# Relation between parasites and aggregation densities, without pseudo

mod_par_agg_no_ps <- glmer(Cell_parasitized ~ scale(Osmia_no_ps_post) + scale(Day_of_year) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_no_ps)) # 
check_model(mod_par_agg_no_ps) # 
testDispersion(mod_par_agg_no_ps, type = 'DHARMa') #  
testZeroInflation(mod_par_agg_no_ps) # 
summary(mod_par_agg_no_ps) # 

ggpredict(mod_par_agg_no_ps, c('Osmia_no_ps_post')) %>% plot(rawdata = TRUE, jitter = .04) 


# MODEL 9
# Relation between parasites and aggregation densities, unsealed

mod_par_agg_unsealed <- glmer(Cell_parasitized ~ scale(Sum_unsealed) + scale(Day_of_year) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_unsealed)) # 
check_model(mod_par_agg_unsealed) # 
testDispersion(mod_par_agg_unsealed, type = 'DHARMa') #  
testZeroInflation(mod_par_agg_unsealed) # 
summary(mod_par_agg_unsealed) # 

ggpredict(mod_par_agg_unsealed, c('Sum_unsealed')) %>% plot(rawdata = TRUE, jitter = .04) 

#### CBIND TO ACCOUNT FOR THE TOTAL NESTS?





# MODEL 10
# How the presence of parasites effects nesting preferences, with pseudo

mod_choice_par_ps <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Par_nests_unsealed) * scale(Osmia_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_1, family = binomial)

plot(simulateResiduals(mod_choice_par_ps)) # 
check_model(mod_choice_par_ps) # 
testDispersion(mod_choice_par_ps, type = 'DHARMa') #  
testZeroInflation(mod_choice_par_ps) # 
summary(mod_choice_par_ps) # 

interact_plot(mod_choice_par_ps, pred = Osmia_ps_prior, modx = Par_nests_unsealed, data = nesting_1)


# Model 11
# How the presence of parasites effects nesting preferences, without pseudo

mod_choice_par_no_ps <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Par_nests_unsealed) * scale(Osmia_no_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_1, family = binomial)

plot(simulateResiduals(mod_choice_par_no_ps)) # 
check_model(mod_choice_par_no_ps) # 
testDispersion(mod_choice_par_no_ps, type = 'DHARMa') #  
testZeroInflation(mod_choice_par_no_ps) # 
summary(mod_choice_par_no_ps) # 

interact_plot(mod_choice_par_no_ps, pred = Osmia_no_ps_prior, modx = Par_nests_unsealed, data = nesting_1)


```



```{r Investment, include=FALSE}


# MODEL 12
# How the presence of parasites effects nest investment
# Likelyhood of abandoning a nest after a cell was parasitized
# position of cell (last or not) ~ if cell is parasitized + day of year + site + (1|Subsite/Trapnest/Nest_ID)




# MODEL 13
# Investment into a nest if other active nests are parasitized

mod_invest_filled <- glmer(cbind(Filled_length, Straw_length) ~ scale(Par_nests_unsealed) * Nest_parasitized + scale(Day_of_year) + Site + (1|Subsite/Trapnest), data = nesting_investment, family = binomial)

plot(simulateResiduals(mod_invest_filled)) # weird
check_model(mod_invest_filled) # high variability?
summary(mod_invest_filled) # 


```
