---
title: "Analyses"
format: 
 html:
  self-contained: true
  toc: true
editor: visual
---

```{r Load, include=FALSE}

## Loading packages
library(glmmTMB)
library(DHARMa)
library(lme4)
library(car)
library(see)
library(performance)
library(sjPlot) # for plot_model
library(ggplot2)
library(ggeffects)
library(effects)
library(tidyverse)
library(interactions) # for interact_plot
library(betareg)


## Loading data
  # nest-level data for all cavities
nesting_os <- read_csv('data/nesting_cleaned_osmia.csv')
nesting_meg <- read_csv('data/nesting_cleaned_megachile.csv') 
  # trapnest-level data
trapnesting_os <- read_csv('data/trapnesting_cleaned_osmia.csv')
trapnesting_meg <- read_csv('data/trapnesting_cleaned_megachile.csv')
  # cell-level data
nesting_cells_os <- read_csv('data/nesting_cells_cleaned_osmia.csv')
nesting_cells_meg <- read_csv('data/nesting_cells_cleaned_megachile.csv')
  # nest-level data for investment measures
nesting_investment_os <- read_csv('data/investment_cleaned_osmia.csv')
nesting_investment_meg <- read_csv('data/investment_cleaned_megachile.csv')


## Inspect
# checking form of Day_of_year
plot(I(New_nests_os/ (Osmia_no_ps_prior)) ~ (Day_of_year), data = trapnesting_os)
plot(I(New_nests_meg/ (Meg_no_ps_prior)) ~ (Day_of_year), data = trapnesting_meg)
# inspect correlation for detectable and non detectable parasites
plot(jitter(nesting_os$NDP_os_prior) ~ jitter(nesting_os$DP_os_prior))
cor(nesting_os$NDP_os_prior, nesting_os$DP_os_prior) # good
plot(jitter(nesting_meg$NDP_meg_prior) ~ jitter(nesting_meg$DP_meg_prior))
cor(nesting_meg$NDP_meg_prior, nesting_meg$DP_meg_prior) # good


```

```{r Notes, include=TRUE}


## Notes
# in ggpredict(), the percentages on the y-axis are the predicted probabilities of the response variable (e.g., new nests) at each of the x axis values (e.g., number of osmia in the trapnest)
# in interact_plot(), the y-axis are the predicted values of the outcome variable (e.g., proportion of straw filled) with the different combinations of the two main predictors (e.g., number of parasitized unsealed nests and if the focal nest is parasitized)


```

```{r Prep_for_Simulations, include=FALSE}


## Create the function to simulate data

# simulate the null hypothesis that bees are treating each cavity equally, thus nesting would be based just on availability
# simulation without replacement so that if there are multiple new osmia nests, the number of available cavities goes down by one for each sequential new nest in that subsite-group combo

# create a function to simulate the distribution of nests based on the null hypothesis that bees are nesting according to available cavities 
simulate_nests <- function(avail_cavities, total_nests) {
  n <- length(avail_cavities) # number of available cavities
  nest_counts <- rep(0, n) # start with zero nests in each trapnest

  # keep as zeros is no nests are built
  if (total_nests == 0) {
    return(nest_counts)
  }

  # look through each of the new nests to be randomly assigned
  for (i in 1:total_nests) {
    if (sum(avail_cavities) == 0) break  # stop if all cavities used
    probability <- avail_cavities / sum(avail_cavities) # calculate probability proportional to the number of available cavities
    chosen <- sample(1:n, size = 1, prob = probability) # randomly choose a trapnest for the new nest based on available cavities
    nest_counts[chosen] <- nest_counts[chosen] + 1 # record it
    avail_cavities[chosen] <- avail_cavities[chosen] - 1 # reduce the number of available cavities by 1, when we apply the function the loop will run again until no new nests are left
  }
  return(nest_counts) # return the vector with how many nests were randomly assigned
}

# indicate how many simulated data columns we want for the model
n_sim <- 100 
```

## Congeneric Attraction - Osmia

Model 1_i - including pseudo

Model 1_ii - excluding pseudo

```{r Congeneric_Attraction_Osmia, include=TRUE}


## Apply the function

trapnesting_os_null <- trapnesting_os %>%
  group_by(Subsite, Date) %>%
  group_modify(~ { # simulate random nest distribution for subsite-date combo
    for (i in 1:n_sim){ # make each of the columns we indicated that we want 
    nest_counts <- simulate_nests(.x$Available_cavities_prior, sum(.x$New_nests_os)) # use available cavities as weights and sum the new nests
    .x[[paste0('Null_new_nests', i)]] <- nest_counts # make the new column
    }
    .x # return the modified group
  }
  ) %>%
  ungroup()


# MODEL 1.i - Osmia including pseudo #

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
sim_list_os_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Osmia_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_os_ps <- c(unlist(lapply(sim_list_os_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_choice_consp_os_ps <- glmer(New_nests_os ~ scale(Osmia_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_choice_consp_os_ps)

# extract my observed coefficient
obvs_coef_os_ps <- summary(mod_choice_consp_os_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_os_ps, c(0.025, 0.975))
obvs_coef_os_ps # falls outside of range

# convert to dataframe
df_coefs_os_ps <- data.frame(sim_coefs_os_ps = sim_coefs_os_ps)

ggplot(df_coefs_os_ps, aes(x = sim_coefs_os_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_os_ps, color = 'red', linetype = 'dashed', size = 1) 
# reject the null - Osmia are attracted to congenerics 


# MODEL 1.ii - Osmia excluding pseudo #

## Simulated data 

# run the model on each of the simulated data columns
sim_list_os_no_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_os_no_ps <- c(unlist(lapply(sim_list_os_no_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Observed data 

mod_choice_consp_os_no_ps <- glmer(New_nests_os ~ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_choice_consp_os_no_ps)

# extract my observed coefficient
obvs_coef_os_no_ps <- summary(mod_choice_consp_os_no_ps)$coefficients[2,1]

# does my observed value fall outside this range, and we reject the null?
quantile(sim_coefs_os_no_ps, c(0.025, 0.975))
obvs_coef_os_no_ps # falls outside of range

# convert to dataframe
df_coefs_os_no_ps <- data.frame(sim_coefs_os_no_ps = sim_coefs_os_no_ps)

# plot
ggplot(df_coefs_os_no_ps, aes(x = sim_coefs_os_no_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_os_no_ps, color = 'red', linetype = 'dashed', size = 1) 
# reject the null - Osmia are attracted to congenerics

```

## Congeneric Attraction - Megachile

Model 1_iii - including pseudo

Model 1_iv - excluding pseudo

```{r Congeneric_Attraction_Megachile, include=TRUE}


## Apply the function

trapnesting_meg_null <- trapnesting_meg %>%
  group_by(Subsite, Date) %>%
  group_modify(~ { # simulate random nest distribution for subsite-date combo
    for (i in 1:n_sim){ # make each of the columns we indicated that we want 
    nest_counts <- simulate_nests(.x$Available_cavities_prior, sum(.x$New_nests_meg)) # use available cavities as weights and sum the new nests
    .x[[paste0('Null_new_nests', i)]] <- nest_counts # make the new column
    }
    .x # return the modified group
  }
  ) %>%
  ungroup()


# MODEL 1.iii - Megachile including pseudo #

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns
sim_list_em_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Meg_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_meg_ps <- c(unlist(lapply(sim_list_em_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))

## Making the model part 2: Observed data 

mod_choice_consp_meg_ps <- glmer(New_nests_meg ~ scale(Meg_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
summary(mod_choice_consp_meg_ps)

# extract my observed coefficient
obvs_coef_meg_ps <- summary(mod_choice_consp_meg_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_meg_ps, c(0.025, 0.975))
obvs_coef_meg_ps # not within this range

# convert to dataframe
df_coefs_meg_ps <- data.frame(sim_coefs_meg_ps = sim_coefs_meg_ps)

ggplot(df_coefs_meg_ps, aes(x = sim_coefs_meg_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_meg_ps, color = 'red', linetype = 'dashed', size = 1) 
# reject the null - Megachile are attracted to congenerics 

# MODEL 1.iv - Megachile excluding pseudo #

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns
sim_list_em_no_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Meg_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_meg_no_ps <- c(unlist(lapply(sim_list_em_no_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_choice_consp_meg_no_ps <- glmer(New_nests_meg ~ scale(Meg_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
summary(mod_choice_consp_meg_no_ps)

# extract my observed coefficient
obvs_coef_meg_no_ps <- summary(mod_choice_consp_meg_no_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_meg_no_ps, c(0.025, 0.975))
obvs_coef_meg_no_ps # falls outside this range

# convert to dataframe
df_coefs_meg_no_ps <- data.frame(sim_coefs_meg_no_ps = sim_coefs_meg_no_ps)

ggplot(df_coefs_meg_no_ps, aes(x = sim_coefs_meg_no_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_meg_no_ps, color = 'red', linetype = 'dashed', size = 1) 
# reject the null - Megachile are attracted to congenerics 


```

## Parasite Preferences - Osmia (Sapyga)

Model 2_i - including pseudo

Model 2_ii - excluding pseudo

Model 2_iii - active nests

```{r Parasites_Attraction_Osmia, include=TRUE}

# Relation between parasites and aggregation densities

# MODEL 2_i - Osmia parasite aggregation preferences, including pseudo #

mod_par_agg_os_ps <- glmer(Cell_parasitized ~ scale(Osmia_ps) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_os_ps)) # good
check_model(mod_par_agg_os_ps) # good
testDispersion(mod_par_agg_os_ps, type = 'DHARMa') # good
testZeroInflation(mod_par_agg_os_ps) # good
summary(mod_par_agg_os_ps) # no detected effects
vif(mod_par_agg_os_ps) # good

# plot
values_osmia_ps <- seq(min(nesting_cells_os$Osmia_ps, na.rm = TRUE), 
               max(nesting_cells_os$Osmia_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_os_ps <- ggpredict(mod_par_agg_os_ps, 
               terms = "Osmia_ps [all]", 
               conditions = list(Osmia_ps = values_osmia_ps))
plot(marginal_effects_os_ps, jitter = 0.04)  # more parasitism for least and most osmia numbers


# MODEL 2_ii - Osmia parasite aggregation preferences, excluding pseudo #

mod_par_agg_os_no_ps <- glmer(Cell_parasitized ~ scale(Osmia_no_ps) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_os_no_ps)) # good
check_model(mod_par_agg_os_no_ps) # good
testDispersion(mod_par_agg_os_no_ps, type = 'DHARMa') # good
testZeroInflation(mod_par_agg_os_no_ps) # good
summary(mod_par_agg_os_no_ps) # no detected effects
vif(mod_par_agg_os_no_ps)

# plot
values_osmia_no_ps <- seq(min(nesting_cells_os$Osmia_no_ps, na.rm = TRUE), 
               max(nesting_cells_os$Osmia_no_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_os_no_ps <- ggpredict(mod_par_agg_os_no_ps, 
               terms = "Osmia_no_ps [all]", 
               conditions = list(Osmia_no_ps = values_osmia_no_ps))
plot(marginal_effects_os_no_ps, jitter = 0.04) # trend with less parasitism for more osmia 


# MODEL 2_iii - Osmia parasite aggregation preferences for active nests #

mod_par_agg_os_active <- glmer(Cell_parasitized ~ scale(Osmia_active_mothers) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_os_active)) # good
check_model(mod_par_agg_os_active) # good
testDispersion(mod_par_agg_os_active, type = 'DHARMa') # good
testZeroInflation(mod_par_agg_os_active) # good
summary(mod_par_agg_os_active) # no detected effects
vif(mod_par_agg_os_active) # good

# plot
values_osmia_active <- seq(min(nesting_cells_os$Osmia_active_mothers, na.rm = TRUE), 
               max(nesting_cells_os$Osmia_active_mothers, na.rm = TRUE), 
               length.out = 100)
marginal_effects_os_actives <- ggpredict(mod_par_agg_os_active, 
               terms = "Osmia_active_mothers [all]", 
               conditions = list(Osmia_active_mothers = values_osmia_active))
plot(marginal_effects_os_actives, jitter = 0.04) # trend for more parasitism with more active mothers 

```

## Parasite Preferences - Megachile (Meloid, Coelioxys, Chalcid)

Model 2_iv - including pseudo

Model 2_v - excluding pseudo

Model 2_vi - active nests

```{r Parasites_Attraction_Megachile, include=TRUE}

# Relation between parasites and aggregation densities (small sample size for Megachile)

# MODEL 2_iv - Megachile parasite aggregation preferences, including pseudo #

mod_par_agg_meg_ps <- glmer(Cell_parasitized ~ scale(Meg_ps) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_meg, family = binomial)

plot(simulateResiduals(mod_par_agg_meg_ps)) # fine
check_model(mod_par_agg_meg_ps) # good
testDispersion(mod_par_agg_meg_ps, type = 'DHARMa') # close to being underdisperesed
testZeroInflation(mod_par_agg_meg_ps) # good
summary(mod_par_agg_meg_ps) # no detected effects
vif(mod_par_agg_meg_ps) # good

# plot
values_meg_ps <- seq(min(nesting_cells_meg$Meg_ps, na.rm = TRUE), 
               max(nesting_cells_meg$Meg_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_meg_ps <- ggpredict(mod_par_agg_meg_ps, 
               terms = "Meg_ps [all]", 
               conditions = list(Meg_ps = values_meg_ps))
plot(marginal_effects_meg_ps, jitter = 0.04) # no trend


# MODEL 2_v - Megachile parasite aggregation preferences, excluding pseudo #

mod_par_agg_meg_no_ps <- glmer(Cell_parasitized ~ scale(Meg_no_ps) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_meg, family = binomial)

plot(simulateResiduals(mod_par_agg_meg_no_ps)) # good
check_model(mod_par_agg_meg_no_ps) # good
testDispersion(mod_par_agg_meg_no_ps, type = 'DHARMa') # good
testZeroInflation(mod_par_agg_meg_no_ps) # good
summary(mod_par_agg_meg_no_ps) # no detected effects
vif(mod_par_agg_meg_no_ps) # good

# plot
values_meg_no_ps <- seq(min(nesting_cells_meg$Meg_no_ps, na.rm = TRUE), 
               max(nesting_cells_meg$Meg_no_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_meg_no_ps <- ggpredict(mod_par_agg_meg_no_ps, 
               terms = "Meg_no_ps [all]", 
               conditions = list(Meg_no_ps = values_meg_no_ps))
plot(marginal_effects_meg_no_ps, jitter = 0.04) # no trend 


# MODEL 2_vi - Megachile parasite aggregation preferences for active nests #

mod_par_agg_meg_active <- glmer(Cell_parasitized ~ scale(Meg_active_mothers) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_meg, family = binomial)

plot(simulateResiduals(mod_par_agg_meg_active)) # not great
check_model(mod_par_agg_meg_active) # good
testDispersion(mod_par_agg_meg_active, type = 'DHARMa') # underdispersed
testZeroInflation(mod_par_agg_meg_active) # zero inflation 
summary(mod_par_agg_meg_active) # no detected effects
vif(mod_par_agg_meg_active) # good

# plot
values_meg_active <- seq(min(nesting_cells_meg$Meg_active_mothers, na.rm = TRUE), 
               max(nesting_cells_meg$Meg_active_mothers, na.rm = TRUE), 
               length.out = 100)
marginal_effects_meg_active <- ggpredict(mod_par_agg_meg_active, 
               terms = "Meg_active_mothers [all]", 
               conditions = list(Meg_active_mothers = values_meg_active))
plot(marginal_effects_meg_active, jitter = 0.04) # no trend


```

## Parasite Avoidance - Osmia

Model 3a_i - including pseudo

Model 3a_ii - excluding pseudo

```{r Parasite_Avoidance_Osmia, include=TRUE}

# MODEL 3a_i - Osmia avoidance of parasites, including pseudo #

# run the model on each of the simulated data columns
sim_list_os_ps_avoid <- lapply(1:n_sim, function(x){
  my_formula <- paste0("Null_new_nests", x, " ~ scale(DP_os_prior) * scale(NDP_ps_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)") 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson())
  summary(output)
} )

# take all of the simulated coefficients
sim_coefs_os_ps_avoid <- c(unlist(lapply(sim_list_os_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_choice_consp_par_os_ps <- glmer(New_nests_os ~ scale(DP_os_prior) * scale(NDP_ps_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_choice_consp_par_os_ps)

# extract my observed coefficient
obvs_coef_os_ps_avoid <- summary(mod_choice_consp_par_os_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_os_ps_avoid, c(0.025, 0.975))
obvs_coef_os_ps_avoid # close but falls within this range

# Convert to dataframe
df_coefs_os_ps_avoid <- data.frame(sim_coefs_os_ps_avoid = sim_coefs_os_ps_avoid)

ggplot(df_coefs_os_ps_avoid, aes(x = sim_coefs_os_ps_avoid)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_os_ps_avoid, color = "red2", linetype = "dashed", size = 1) 
# fail to reject the null - no evidence for Osmia avoiding nests with parasites


# MODEL 3a_ii - Osmia avoidance of parasites, excluding pseudo #

# run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
sim_list_os_no_ps_avoid <- lapply(1:n_sim, function(x){
  my_formula <- paste0("Null_new_nests", x, " ~ scale(DP_os_prior) * scale(NDP_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)") 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson())
  summary(output)
} )

# take all of the simulated coefficients
sim_coefs_os_no_ps_avoid <- c(unlist(lapply(sim_list_os_no_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_choice_consp_par_os_no_ps <- glmer(New_nests_os ~ scale(DP_os_prior) * scale(NDP_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_choice_consp_par_os_no_ps)

# extract my observed coefficient
obvs_coef_os_no_ps_avoid <- summary(mod_choice_consp_par_os_no_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_os_no_ps_avoid, c(0.025, 0.975))
obvs_coef_os_no_ps_avoid # falls outside this range

# Convert to dataframe
df_coefs_os_no_ps_avoid <- data.frame(sim_coefs_os_no_ps_avoid = sim_coefs_os_no_ps_avoid)

ggplot(df_coefs_os_no_ps_avoid, aes(x = sim_coefs_os_no_ps_avoid)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_os_no_ps_avoid, color = "red2", linetype = "dashed", size = 1) 
# small effect but reject the null? - Osmia avoid nesitng near parasites?


```

## Parasite Avoidance - Megachile

Model 3a_iii - including pseudo

Model 3a_iv - excluding pseudo

```{r Parasite_Avoidance_Megachile, include=TRUE}


# MODEL 3a_iii - Megachile avoidance of parasites, including pseudo #

# run the model on each of the simulated data columns
sim_list_meg_ps_avoid <- lapply(1:n_sim, function(x){
  my_formula <- paste0("Null_new_nests", x, " ~ scale(DP_meg_prior) * scale(NDP_ps_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)") 
  output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson())
  summary(output)
} )

# take all of the simulated coefficients
sim_coefs_meg_ps_avoid <- c(unlist(lapply(sim_list_meg_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))

## Making the model part 2: Observed data 

mod_choice_consp_par_meg_ps <- glmer(New_nests_meg ~ scale(DP_meg_prior) * scale(NDP_ps_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
summary(mod_choice_consp_par_meg_ps)

# extract my observed coefficient
obvs_coef_meg_ps_avoid <- summary(mod_choice_consp_par_meg_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_meg_ps_avoid, c(0.025, 0.975))
obvs_coef_meg_ps_avoid # falls within this range

# Convert to dataframe
df_coefs_meg_ps_avoid <- data.frame(sim_coefs_meg_ps_avoid = sim_coefs_meg_ps_avoid)

ggplot(df_coefs_meg_ps_avoid, aes(x = sim_coefs_meg_ps_avoid)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_meg_ps_avoid, color = "red2", linetype = "dashed", size = 1) 
# fail to reject the null - Megachile do not avoid nesitng near parasites


# MODEL 3a_iv - Megachile avoidance of parasites, excluding pseudo #

# running the simulation gets the error "Error: PIRLS loop resulted in NaN value"
# too little data to run it

# # run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
# sim_list_meg_no_ps_avoid <- lapply(1:n_sim, function(x){
#   my_formula <- paste0("Null_new_nests", x, " ~ scale(DP_meg_prior) * scale(NDP_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)")
#   output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson())
#   summary(output)
# } )
# 
# # take all of the simulated coefficients
# sim_coefs_meg_no_ps_avoid <- c(unlist(lapply(sim_list_meg_no_ps_avoid, function(x) {
#   x$coefficients[2,1] # extract second row, first column for the estimate
# })))
# 
# 
# ## Making the model part 2: Observed data
# 
# mod_choice_consp_par_meg_no_ps <- glmer(New_nests_meg ~ scale(DP_meg_prior) * scale(NDP_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
# summary(mod_choice_consp_par_meg_no_ps)
# 
# # extract my observed coefficient
# obvs_coef_meg_no_ps_avoid <- summary(mod_choice_consp_par_meg_no_ps)$coefficients[2,1]
# 
# # plot my observed values with the simulated coefficients
# # abline(v=obvs_coef_meg_no_ps_avoid)
# 
# # range of 95%, i.e., if my observed value falls within this range, its not different from the null
# quantile(sim_coefs_meg_no_ps_avoid, c(0.025, 0.975))
# 
# # Convert to dataframe
# df_coefs_meg_no_ps_avoid <- data.frame(sim_coefs_meg_no_ps_avoid = sim_coefs_meg_no_ps_avoid)
# 
# ggplot(df_coefs_meg_no_ps_avoid, aes(x = sim_coefs_meg_no_ps_avoid)) +
#   geom_histogram(binwidth = 0.1) +
#   geom_vline(xintercept = obvs_coef_meg_no_ps_avoid, color = "red2", linetype = "dashed", size = 1)


```

## Investment

Model 3b_i - length of straw filled, Osmia

Model 3b_ii - length of straw filled, Megachile

Model 3c_i - nest termination, Osmia

Model 3c_ii - nest termination, Megachile

Model 3d_i - abandoned inner wall, Osmia

Model 3d_ii - abandoned inner wall, Megachile

```{r Nest_Investment, include=TRUE}

# MODEL 3b_i - Osmia investment #

# Investment into a nest if other active nests or the nest itself is parasitized

mod_invest_filled_os <- glmer(cbind(Filled_length, Straw_length) ~ scale(DP_os_prior) * Nest_parasitized + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = nesting_investment_os, family = binomial, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1000000)))

plot(simulateResiduals(mod_invest_filled_os)) # fine
check_model(mod_invest_filled_os) # fine, some influential points
# check the influential points (red dots on the influential observations figure)
nesting_cells_os[rownames(nesting_cells_os) %in% c(89, 69), ] # 
summary(mod_invest_filled_os) # significant effect of if focal nest is parasitized, day of year, and year
vif(mod_invest_filled_os) # good

plot(ggpredict(mod_invest_filled_os, terms = c("DP_os_prior", "Nest_parasitized"))) 


# MODEL 3b_ii - Megachile investment #

# Investment into a nest if other active nests or the nest itself is parasitized

mod_invest_filled_meg <- glmer(cbind(Filled_length, Straw_length) ~ scale(DP_meg_prior) * Nest_parasitized + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = nesting_investment_meg, family = binomial, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1000000)))

plot(simulateResiduals(mod_invest_filled_meg)) # poor
check_model(mod_invest_filled_meg) # fine
summary(mod_invest_filled_meg) # significant effect of detectable parasites in neighbouring nests, boardlerine effect of if focal nest is parasitized, day of year, site, and interaction of detectable parasite * focal nest parasitized
vif(mod_invest_filled_meg) # good


plot(ggpredict(mod_invest_filled_meg, terms = c("DP_meg_prior", "Nest_parasitized"))) 


# MODEL 3c_i - Osmia nest termination #

mod_last_cell_os <- glmer(Cell_position_bi ~ Cell_parasitized + poly(Day_of_year,2) + Year + Site + (1|Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_last_cell_os)) # good
check_model(mod_last_cell_os) # high VIF for year and site
vif(mod_last_cell_os) # TERRIBLE - exclude 2024?
summary(mod_last_cell_os) # significant effect of day of year

ggpredict(mod_last_cell_os, terms = "Cell_parasitized [all]") %>%
  plot(jitter = .04)


# MODEL 3c_ii - Megachile nest termination #

mod_last_cell_meg <- glmer(Cell_position_bi ~ Cell_parasitized + poly(Day_of_year,2) + Year + Site + (1|Nest_ID), data = nesting_cells_meg, family = binomial)

plot(simulateResiduals(mod_last_cell_meg)) # good
check_model(mod_last_cell_meg) # good
summary(mod_last_cell_meg) # significant effect of day of year, almost site
vif(mod_last_cell_meg) # good

ggpredict(mod_last_cell_meg, terms = "Cell_parasitized [all]") %>%
  plot(jitter = .04)
# slight trend for more likely to best last cell if its parasitized


# MODEL 3d_i - Osmia abandoned inner walls including pseudo #

mod_inner_wall_os <- glmer(Abandoned_inner_wall ~ scale(DP_os_prior) * scale(NDP_ps_os_prior) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = nesting_os, family = binomial)

plot(simulateResiduals(mod_inner_wall_os)) # good
check_model(mod_inner_wall_os) # good
summary(mod_inner_wall_os) # significant effect of day of year
vif(mod_inner_wall_os) # good

ggpredict(mod_inner_wall_os, terms = "DP_os_prior [all]") %>%
  plot(jitter = .04)


# MODEL 3d_ii - Megachile abandoned inner walls #

mod_inner_wall_meg <- glmer(Abandoned_inner_wall ~ scale(DP_meg_prior) * scale(NDP_ps_meg_prior) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = nesting_meg, family = binomial)

plot(simulateResiduals(mod_inner_wall_meg)) # fine
check_model(mod_inner_wall_meg) # high colinearity 
vif(mod_inner_wall_meg) # TERRIBLE - small sample size 
summary(mod_inner_wall_meg) # no detected effects


```

```{r Ignore, include=FALSE}



## checking unscaled estimates --- this is old as of May 2025, not sure where these numbers (0.2578, etc.) came from

mean_meg_no_ps_prior <- mean(nesting_meg$Meg_no_ps_prior)
sd_meg_no_ps_prior <- sd(nesting_meg$Meg_no_ps_prior)
scaled_estimate_meg <- 0.2578
unscaled_estimate_meg <- scaled_estimate_meg * sd_meg_no_ps_prior + mean_meg_no_ps_prior
unscaled_estimate_meg

mean_os_no_ps_prior <- mean(nesting_os$Osmia_no_ps_prior)
sd_os_no_ps_prior <- sd(nesting_os$Osmia_no_ps_prior)
scaled_estimate_os <- 0.36480
unscaled_estimate_os <- scaled_estimate_os * mean_os_no_ps_prior + mean_os_no_ps_prior
unscaled_estimate_os

mean_os_ps_prior <- mean(nesting_os$Osmia_ps_prior)
sd_os_ps_prior <- sd(nesting_os$Osmia_ps_prior)
scaled_estimate_os_ps <- 0.85564
unscaled_estimate_os_ps <- scaled_estimate_os_ps * mean_os_ps_prior + mean_os_ps_prior
unscaled_estimate_os_ps


```
