---
title: "Analyses"
format: 
 html:
  self-contained: true
  toc: true
editor: visual
bibliography: references.bib
---

**Response variables:**

-   New_nests_os: number of new osmia nests for each trapnest and date combination.
-   New_nests_meg: number of new megachile nests for each trapnest and date combination.
-   Prop_filled: proportion of the straw filled with cells, used as a measure of investment into a nest.

**Predictor variables:**

-   Available_cavities_prior: number of available cavities that could have had a new nest built between visits.
-   Osmia_ps_prior: number of existing osmia in each trapnest per date, including pseudo nests.
-   Osmia_no_ps_prior: number of existing osmia in each trapnest per date, excluding pseudo nests.
-   Meg_ps_prior: number of existing megachile in each trapnest per date, including pseudo nests.
-   Meg_no_ps_prior: number of existing megachile in each trapnest per date, excluding pseudo nests.
-   Par_in_trapnest: number of parasites present in the trapnest when I log new bee nests (Bee_choice).

**Additional variables:**

-   Good_days: number of days between visits with good enough weather for bee activity.
-   Day_of_year: the day of the year, a stand in for date.
-   Site: three sites in total.
-   Subsite: six subsites for each site, unique to each site.
-   Trapnest: 72 trapnests total, ach subsite contains three, unique to each subsite.

```{r Load}



#### CHECK IF BINARIES ARE CODED AS 1 AND 0

## Loading packages

library(glmmTMB)
library(DHARMa)
library(lme4)
library(car)
library(see)
library(performance)
library(sjPlot) # for plot_model
library(ggplot2)
library(ggeffects)
library(effects)
library(tidyverse)
library(interactions) # for interact_plot
library(betareg)


## Loading data

nesting_os <- read_csv('data/nesting_cleaned_osmia.csv')
nesting_meg <- read_csv('data/nesting_cleaned_megachile.csv') 
nesting_cells_os <- read_csv('data/nesting_cells_cleaned_osmia.csv')
nesting_cells_meg <- read_csv('data/nesting_cells_cleaned_megachile.csv')
nesting_investment_os <- read_csv('data/investment_cleaned_osmia.csv')
nesting_investment_meg <- read_csv('data/investment_cleaned_megachile.csv')

```

```{r Inspecting}


## Checking form of Day_of_year

plot(I(New_nests_os/ (Osmia_no_ps_prior)) ~ (Day_of_year), data=nesting_os)
plot(I(New_nests_meg/ (Meg_no_ps_prior)) ~ (Day_of_year), data=nesting_meg)

### check date sin other dfs too?




## Verifying my treatments: How related are my treatments to the number of available cavities?

 ## Osmia ##
confirm_treatment_data_os <- nesting_os %>%
  distinct(Treatment, Subsite, Day_of_year, .keep_all = TRUE)
# mean and SD
confirm_treatment_data_os %>% 
  group_by(Treatment) %>% 
  summarise(Mean_avail = mean(Available_cavities_prior),
            SD_avail = sd(Available_cavities_prior))

# treatment   mean       SD
  # Low	  17.330623	  1.0910565		
  # Mid	  11.596206	  0.9826648	
  # High	6.655827	  0.8468795	

# range 
confirm_treatment_data_os %>% 
  filter(Treatment == 'Low') %>% 
  summarise(Min_low = min(Available_cavities_prior),
            Max_low = max(Available_cavities_prior)) # 12, 18
confirm_treatment_data_os %>% 
  filter(Treatment == 'Mid') %>% 
  summarise(Min_mid = min(Available_cavities_prior),
            Max_mid = max(Available_cavities_prior)) # 5, 13
confirm_treatment_data_os %>% 
  filter(Treatment == 'High') %>% 
  summarise(Min_high = min(Available_cavities_prior),
            Max_high = max(Available_cavities_prior)) # 1, 9

 ## Megachile ##
confirm_treatment_data_meg <- nesting_meg %>%
  distinct(Treatment, Subsite, Day_of_year, .keep_all = TRUE)
# mean and SD
confirm_treatment_data_meg %>% 
  group_by(Treatment) %>% 
  summarise(Mean_avail = mean(Available_cavities_prior),
            SD_avail = sd(Available_cavities_prior))

# treatment  mean        SD
  # High	6.655827	  0.8468795		
  # Low	  17.330623	  1.0910565		
  # Mid	  11.596206	  0.9826648	

# range 
confirm_treatment_data_meg %>% 
  filter(Treatment == 'Low') %>% 
  summarise(Min_low = min(Available_cavities_prior),
            Max_low = max(Available_cavities_prior)) # 12, 18
confirm_treatment_data_meg %>% 
  filter(Treatment == 'Mid') %>% 
  summarise(Min_mid = min(Available_cavities_prior),
            Max_mid = max(Available_cavities_prior)) # 5, 13
confirm_treatment_data_meg %>% 
  filter(Treatment == 'High') %>% 
  summarise(Min_high = min(Available_cavities_prior),
            Max_high = max(Available_cavities_prior)) # 1, 9



## Notes
# in ggpredict(), the percentages on the y-axis are the predicted probabilities of the response variable (e.g., new nests) at each of the x axis values (e.g., number of osmia in the trapnest)
# in interact_plot(), the y-axis are the predicted values of the outcome variable (e.g., proportion of straw filled) with the different combinations of the two main predictors (e.g., number of parasitized unsealed nests and if the focal nest is parasitized)


```

```{r Prep_for_Simulations}


## Create the function to simulate data

# simulate the null hypothesis that bees are treating each cavity equally, thus nesting would be based just on availability
# simulation without replacement so that if there are multiple new osmia nests, the number of available cavities goes down by one for each sequential new nest in that subsite-group combo

# create a function to simulate the distribution of nests based on the null hypothesis that bees are nesting according to available cavities 
simulate_nests <- function(avail_cavities, total_nests) {
  n <- length(avail_cavities) # number of available cavities
  nest_counts <- rep(0, n) # start with zero nests in each trapnest

  # keep as zeros is no nests are built
  if (total_nests == 0) {
    return(nest_counts)
  }

  # look through each of the new nests to be randomly assigned
  for (i in 1:total_nests) {
    if (sum(avail_cavities) == 0) break  # stop if all cavities used
    probability <- avail_cavities / sum(avail_cavities) # calculate probability proportional to the number of available cavities
    chosen <- sample(1:n, size = 1, prob = probability) # randomly choose a trapnest for the new nest based on available cavities
    nest_counts[chosen] <- nest_counts[chosen] + 1 # record it
    avail_cavities[chosen] <- avail_cavities[chosen] - 1 # reduce the number of available cavities by 1, when we apply the function the loop will run again until no new nests are left
  }
  return(nest_counts) # return the vector with how many nests were randomly assigned
}

# indicate how many simulated data columns we want for the model
n_sim <- 100 
```

```{r Congeneric_Attraction_Osmia}


## Apply the function

nesting_os_null <- nesting_os %>%
  group_by(Subsite, Date) %>%
  group_modify(~ { # simulate random nest distribution for subsite-date combo
    for (i in 1:n_sim){ # make each of the columns we indicated that we want 
    nest_counts <- simulate_nests(.x$Available_cavities_prior, sum(.x$New_nests_os)) # use available cavities as weights and sum the new nests
    .x[[paste0('Null_new_nests', i)]] <- nest_counts # make the new column
    }
    .x # return the modified group
  }
  ) %>%
  ungroup()


# MODEL 1.i - Osmia including pseudo #

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
sim_list_os_ps <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Osmia_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = nesting_os_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_os_ps <- c(unlist(lapply(sim_list_os_ps, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))

# plot the simulated coefficients
hist(sim_coefs_os_ps)


## Making the model part 2: Observed data 

mod_choice_consp_os_ps <- glmer(New_nests_os ~ scale(Osmia_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_os_null, family = poisson)
summary(mod_choice_consp_os_ps)

# extract my observed coefficient
obvs_coef_os_ps <- summary(mod_choice_consp_os_ps)$coefficients[2,1]

# plot my observed values with the simulated coefficients
abline(v=obvs_coef_os_ps) 

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_os_ps, c(0.025, 0.975))

# convert to dataframe
df_coefs_os_ps <- data.frame(sim_coefs_os_ps = sim_coefs_os_ps)

ggplot(df_coefs_os_ps, aes(x = sim_coefs_os_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_os_ps, color = 'red', linetype = 'dashed', size = 1) 


# MODEL 1.ii - Osmia excluding pseudo #

## Simulated data 

# run the model on each of the simulated data columns
sim_list_os_no_ps <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = nesting_os_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_os_no_ps <- c(unlist(lapply(sim_list_os_no_ps, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))

# plot the simulated coefficients
hist(sim_coefs_os_no_ps)


## Observed data 

mod_choice_consp_os_no_ps <- glmer(New_nests_os ~ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_os_null, family = poisson)
summary(mod_choice_consp_os_no_ps)

# extract my observed coefficient
obvs_coef_os_no_ps <- summary(mod_choice_consp_os_no_ps)$coefficients[2,1]

# plot my observed values with the simulated coefficients
abline(v=obvs_coef_os_no_ps) 

# does my observed value fall outside this range, and we reject the null?
quantile(sim_coefs_os_no_ps, c(0.025, 0.975))

# convert to dataframe
df_coefs_os_no_ps <- data.frame(sim_coefs_os_no_ps = sim_coefs_os_no_ps)

# plot
ggplot(df_coefs_os_no_ps, aes(x = sim_coefs_os_no_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_os_no_ps, color = 'red', linetype = 'dashed', size = 1) 



```

```{r Congeneric_Attraction_Megachile}


## Apply the function

nesting_meg_null <- nesting_meg %>%
  group_by(Subsite, Date) %>%
  group_modify(~ { # simulate random nest distribution for subsite-date combo
    for (i in 1:n_sim){ # make each of the columns we indicated that we want 
    nest_counts <- simulate_nests(.x$Available_cavities_prior, sum(.x$New_nests_meg)) # use available cavities as weights and sum the new nests
    .x[[paste0('Null_new_nests', i)]] <- nest_counts # make the new column
    }
    .x # return the modified group
  }
  ) %>%
  ungroup()


# MODEL 1.iii - Megachile including pseudo #

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
sim_list_meg_ps <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Meg_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = nesting_meg_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_meg_ps <- c(unlist(lapply(sim_list_meg_ps, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))

# plot the simulated coefficients
hist(sim_coefs_meg_ps)


## Making the model part 2: Observed data 

mod_choice_consp_meg_ps <- glmer(New_nests_meg ~ scale(Meg_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_meg_null, family = poisson)
summary(mod_choice_consp_meg_ps)

# extract my observed coefficient
obvs_coef_meg_ps <- summary(mod_choice_consp_meg_ps)$coefficients[2,1]

# plot my observed values with the simulated coefficients
abline(v=obvs_coef_meg_ps) 

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_meg_ps, c(0.025, 0.975))

# convert to dataframe
df_coefs_meg_ps <- data.frame(sim_coefs_meg_ps = sim_coefs_meg_ps)

ggplot(df_coefs_meg_ps, aes(x = sim_coefs_meg_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_meg_ps, color = 'red', linetype = 'dashed', size = 1) 


# MODEL 1.iv - Megachile excluding pseudo #

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns
sim_list_meg_no_ps <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Meg_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = nesting_meg_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_meg_no_ps <- c(unlist(lapply(sim_list_meg_no_ps, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))

# plot the simulated coefficients
hist(sim_coefs_meg_no_ps)


## Making the model part 2: Observed data 

mod_choice_consp_meg_no_ps <- glmer(New_nests_meg ~ scale(Meg_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_meg_null, family = poisson)
summary(mod_choice_consp_meg_no_ps)

# extract my observed coefficient
obvs_coef_meg_no_ps <- summary(mod_choice_consp_meg_no_ps)$coefficients[2,1]

# plot my observed values with the simulated coefficients
abline(v=obvs_coef_meg_no_ps) 

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_meg_no_ps, c(0.025, 0.975))

# convert to dataframe
df_coefs_meg_no_ps <- data.frame(sim_coefs_meg_no_ps = sim_coefs_meg_no_ps)

ggplot(df_coefs_meg_no_ps, aes(x = sim_coefs_meg_no_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_meg_no_ps, color = 'red', linetype = 'dashed', size = 1) 




```

```{r Parasites_and_Aggregations, include=TRUE}

# Relation between parasites and aggregation densities

## PREDICTION 2

# MODEL 7
# Relation between parasites and aggregation densities, with pseudo

mod_par_agg_ps <- glmer(Cell_parasitized ~ scale(Osmia_ps) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_ps)) # good
check_model(mod_par_agg_ps) # good
testDispersion(mod_par_agg_ps, type = 'DHARMa') # good
testZeroInflation(mod_par_agg_ps) # good
summary(mod_par_agg_ps) # no detected effects
vif(mod_par_agg_ps)

# plot
values_osmia_ps_no_iw <- seq(min(nesting_cells_os$Osmia_ps_no_iw, na.rm = TRUE), 
               max(nesting_cells_os$Osmia_ps_no_iw, na.rm = TRUE), 
               length.out = 100)
marginal_effects_os_ps_no_iw <- ggpredict(mod_par_agg_ps, 
               terms = "Osmia_ps_no_iw [all]", 
               conditions = list(Osmia_ps_no_iw = values_osmia_ps_no_iw))
plot(marginal_effects_os_ps_no_iw, rawdata = TRUE, jitter = 0.04)


# MODEL 8
# Relation between parasites and aggregation densities, without pseudo

mod_par_agg_no_ps <- glmer(cbind(Sum_cell_parasitized, Sum_cell_not_parasitized) ~ scale(Osmia_no_ps_no_iw) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_no_ps)) # WEIRD!
check_model(mod_par_agg_no_ps) # fine
testDispersion(mod_par_agg_no_ps, type = 'DHARMa') # good
testZeroInflation(mod_par_agg_no_ps) # good
summary(mod_par_agg_no_ps) # no detected effects
vif(mod_par_agg_no_ps)

table(nesting_cells_os$Site) # very few CP

# plot
values_osmia_no_ps_no_iw <- seq(min(nesting_cells_os$Osmia_no_ps_no_iw, na.rm = TRUE), 
                max(nesting_cells_os$Osmia_no_ps_no_iw, na.rm = TRUE), 
                length.out = 100)
marginal_effects_os_no_ps_no_iw <- ggpredict(mod_par_agg_no_ps, 
               terms = "Osmia_no_ps_no_iw [all]", 
               conditions = list(Osmia_no_ps_no_iw = values_osmia_no_ps_no_iw))
plot(marginal_effects_os_no_ps_no_iw, rawdata = TRUE, jitter = 0.04)


# MODEL 9
# Relation between parasites and aggregation densities, unsealed

mod_par_agg_unsealed_os <- glmer(cbind(Sum_cell_parasitized, Sum_cell_not_parasitized) ~ scale(Sum_unsealed_os_no_iw) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_par_agg_unsealed_os)) # good
check_model(mod_par_agg_unsealed_os) # fine
testDispersion(mod_par_agg_unsealed_os, type = 'DHARMa') # good
testZeroInflation(mod_par_agg_unsealed_os) # good
summary(mod_par_agg_unsealed_os) 
vif(mod_par_agg_unsealed_os) # no detected effects

# plot
values_unsealed_os_no_iw <- seq(min(nesting_cells_os$Sum_unsealed_os_no_iw, na.rm = TRUE), 
                   max(nesting_cells_os$Sum_unsealed_os_no_iw, na.rm = TRUE), 
                   length.out = 100)
marginal_effects_unsealed_os_no_iw <- ggpredict(mod_par_agg_unsealed_os, 
               terms = "Sum_unsealed_os_no_iw [all]", 
               conditions = list(Sum_unsealed_os_no_iw = values_unsealed_os_no_iw))
plot(marginal_effects_unsealed_os_no_iw, rawdata = TRUE, jitter = 0.04)

```

```{r Preferences_and_Parasites, include=TRUE}

## PREDICTION 3a

# run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
sim_list_2 <- lapply(1:n_sim, function(x){
  my_formula <- paste0("Null_new_nests", x, " ~ scale(DP_os_prior) * scale(NDP_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)") 
  output <- glmer(as.formula(my_formula), data = nesting_os_null, family = poisson())
  summary(output)
} )

# take all of the simulated coefficients
sim_coefs_2 <- c(unlist(lapply(sim_list_2, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))

# plot the simulated coefficients
hist(sim_coefs_2)


## Making the model part 2: Observed data 

mod_choice_consp_par_os_ps <- glmer(New_nests_os ~ scale(DP_os_prior) * scale(NDP_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_os_null, family = poisson)
summary(mod_choice_consp_par_os_ps)



# extract my observed coefficient
obvs_coef <- summary(mod_choice_consp_par_os_ps)$coefficients[2,1]

# plot my observed values with the simulated coefficients
abline(v=obvs_coef)

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_2, c(0.025, 0.975))

# Convert to dataframe
df <- data.frame(sim_coefs = sim_coefs)

ggplot(df, aes(x = sim_coefs_2)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef, color = "red2", linetype = "dashed", size = 1) 
















### old - before simulated data ###
# MODEL 10
# How the presence of parasites effects nesting preferences, with pseudo

mod_choice_par_ps_os <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(DP_os_prior) * scale(NDP_ps_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_os, family = binomial, control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1000000))) 
# having subsite doesnt change model

plot(simulateResiduals(mod_choice_par_ps_os)) # fine
check_model(mod_choice_par_ps_os) # fine
testDispersion(mod_choice_par_ps_os, type = 'DHARMa') # good
testZeroInflation(mod_choice_par_ps_os) # fine
summary(mod_choice_par_ps_os) # significant effect of parasitized unsealed nests, congenerics, good days, day of year, and site
vif(mod_choice_par_ps_os) # good

interact_plot(mod_choice_par_ps_os, 
              pred = NDP_ps_os_prior, 
              modx = DP_os_prior, 
              data = nesting_os,
              colors = c("salmon1", "salmon2", "salmon3"))
#  y-axis represents the predicted probability of observing new nests out of the total available cavities, given the values of the predictor variables

# Model 11
# How the presence of parasites effects nesting preferences, without pseudo

mod_choice_par_no_ps_os <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(DP_os_prior) * scale(NDP_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_os, family = binomial, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1000000))) 

plot(simulateResiduals(mod_choice_par_no_ps_os)) # good
check_model(mod_choice_par_no_ps_os) # fine
testDispersion(mod_choice_par_no_ps_os, type = 'DHARMa') # good
testZeroInflation(mod_choice_par_no_ps_os) # good
summary(mod_choice_par_no_ps_os) # significant effect of parasitized unsealed nests, congenerics, parasitized unsealed nests*congenerics, good days, day of year, and site
vif(mod_choice_par_no_ps_os)

interact_plot(mod_choice_par_no_ps_os, 
              pred = NDP_os_prior, 
              modx = DP_os_prior, 
              data = nesting_os,
              colors = c("salmon1", "salmon2", "salmon3"))


```

```{r Investment, include=TRUE}

## PREDICTION 3b

# MODEL 13
# Investment into a nest if other active nests or the nest itself is parasitized

mod_invest_filled <- glmer(cbind(Filled_length, Straw_length) ~ scale(DP_os_prior) * Nest_parasitized + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = investment_os, family = binomial, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1000000)))

plot(simulateResiduals(mod_invest_filled)) # fine
check_model(mod_invest_filled) # fine
summary(mod_invest_filled) # significant effect of parasitized unsealed nests, if focal nest is parasitized, good days, day of year, and site
vif(mod_invest_filled) # good


plot(ggpredict(mod_invest_filled, terms = c("DP_os_prior", "Nest_parasitized"))) 


## mean straw length:
mean(investment_os$Straw_length_mm)
sd(investment_os$Straw_length_mm)



## PREDICTION 3c

# MODEL 12
# Likelyhood of of terminating nest investment after a cell was parasitized

mod_last_cell <- glmer(Last_cell ~ Cell_parasitized + poly(Day_of_year,2) + Site + (1|Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_last_cell)) # good
summary(mod_last_cell) # significant effect of day of year
vif(mod_last_cell) # good

ggpredict(mod_last_cell, c('Cell_parasitized')) %>% plot(rawdata = TRUE, jitter = .04) 




## PREDICTION 3d



```

```{r Ignore, include=TRUE}

## Old models
# I originally used cbind() to create a matrix to account for number of available cavities, but it doesn't really work as well as I want. E.g., 1 new nest into a trapnest with 5 available cavities = 0.2 while 1 new nest into a trapnest with 17 available cavities = 0.06
# so now I'm using simulated data instead
# here are the old models


 # Available Cavities #

# # MODEL 1 
# ### New nests per available cavities - Osmia ###
# mod_choice_cav_os <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Available_cavities_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_os, family = binomial(), control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1000000))) # ignore boudnary singular fit warning, means random effects don't account for much
# 
# plot(simulateResiduals(mod_choice_cav_os)) # fine
# check_model(mod_choice_cav_os) # fine
# testDispersion(mod_choice_cav_os, type = 'DHARMa') # good
# testZeroInflation(mod_choice_cav_os) # good
# vif(mod_choice_cav_os) # good
# summary(mod_choice_cav_os) # Significant effect of good days, site, day of year
# 
# ggpredict(mod_choice_cav_os, c('Available_cavities_prior')) %>% plot(rawdata = TRUE, jitter = .04)
# 
# # MODEL 2
# ### New nests per available cavities - Megachile ###
# mod_choice_cav_meg <- glmer(cbind(New_nests_meg, Available_cavities_prior - New_nests_sum) ~ scale(Available_cavities_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_meg, family = binomial(), control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1000000)))
# 
# plot(simulateResiduals(mod_choice_cav_meg)) # fine
# check_model(mod_choice_cav_meg) # fine
# testDispersion(mod_choice_cav_meg, type = 'DHARMa') # fine
# testZeroInflation(mod_choice_cav_meg) # good
# summary(mod_choice_cav_meg) # significant effect of site
# vif(mod_choice_cav_meg)
# 
# # Plot model
# ggpredict(mod_choice_cav_meg, c('Available_cavities_prior')) %>% plot(rawdata = TRUE, jitter = .04) 

  # congeneric Presence #
 
# # MODEL 3 
# ### New nests per congenerics Osmia, with pseudo ###
# 
# mod_choice_consp_os_ps <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Osmia_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_os, family = binomial) # ignore boundary singular warning, means little to no influence of random effects
# 
# plot(simulateResiduals(mod_choice_consp_os_ps)) # fine
# check_model(mod_choice_consp_os_ps) # fine
# testDispersion(mod_choice_consp_os_ps, type = 'DHARMa') # good
# testZeroInflation(mod_choice_consp_os_ps) # some zero-inflation, but not detrimental
# summary(mod_choice_consp_os_ps) # significant effect of congenerics, good days, date, and site
# vif(mod_choice_consp_os_ps) # good
# 
# # specify my range of values for Osmia_ps_prior
# values_osmia_ps <- seq(min(nesting_os$Osmia_ps_prior), 
#           max(nesting_os$Osmia_ps_prior), 
#           length.out = 100)
# # compute marginal effects at specific values of Osmia_ps_prior
# marginal_effects_os_ps <- ggpredict(mod_choice_consp_os_ps, 
#                terms = "Osmia_ps_prior [all]", 
#                conditions = list(Osmia_ps_prior = values_osmia_ps))
# # plot
# plot(marginal_effects_os_ps, rawdata = TRUE, jitter = 0.04)
# 
# # MODEL 4 
# ### New nests per congenerics Osmia, without pseudo ###
# 
# # Fit the GLMM model with custom glmerControl
# mod_choice_consp_os_no_ps <- glmer(cbind(New_nests_os, Available_cavities_prior - New_nests_sum) ~ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year, 2) + Site + (1|Subsite) + (1|Trapnest), data = nesting_os, family = binomial(), control = glmerControl(optimizer = "nloptwrap", optCtrl = list(maxeval = 100000)))
# 
# plot(simulateResiduals(mod_choice_consp_os_no_ps)) # fine
# testDispersion(mod_choice_consp_os_no_ps) # good
# testZeroInflation(mod_choice_consp_os_no_ps) # fine
# summary(mod_choice_consp_os_no_ps) # significant effect of congenerics, good days, date, and site
# vif(mod_choice_consp_os_no_ps) # good
# 
# # specify my range of values for Osmia_no_ps_prior
# values_osmia_no_ps <- seq(min(nesting_os$Osmia_no_ps_prior), 
#           max(nesting_os$Osmia_no_ps_prior), 
#           length.out = 100)
# # compute marginal effects at specific values of Osmia_ps_prior
# marginal_effects_os_no_ps <- ggpredict(mod_choice_consp_os_no_ps, 
#                terms = "Osmia_no_ps_prior [all]", 
#                conditions = list(Osmia_no_ps_prior = values_osmia_no_ps))
# # plot
# plot(marginal_effects_os_no_ps, rawdata = TRUE, jitter = 0.04)
#        
# 
# # MODEL 5
# ### New nests per congenerics Megachile, with pseudo ###
# mod_choice_consp_meg_ps <- glmer(cbind(New_nests_meg, Available_cavities_prior - New_nests_sum) ~ scale(Meg_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_meg, family = binomial(), control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1000000)))
# 
# plot(simulateResiduals(mod_choice_consp_meg_ps)) # great
# check_model(mod_choice_consp_meg_ps) # good
# testDispersion(mod_choice_consp_meg_ps, type = 'DHARMa') # good
# testZeroInflation(mod_choice_consp_meg_ps) # good
# summary(mod_choice_consp_meg_ps) # significant effect of site (CP) and day of year
# vif(mod_choice_consp_meg_ps)
# 
# # plot
# values_meg_ps <- seq(min(nesting_meg$Meg_ps_prior), 
#           max(nesting_meg$Meg_ps_prior), 
#           length.out = 100)
# marginal_effects_meg_ps <- ggpredict(mod_choice_consp_meg_ps, 
#                terms = "Meg_ps_prior [all]", 
#                conditions = list(Meg_ps_prior = values_meg_ps))
# plot(marginal_effects_meg_ps, rawdata = TRUE, jitter = 0.04)
# 
# # MODEL 6
# ### New nests per congenerics Megachile, without pseudo ###
# mod_choice_consp_meg_no_ps <- glmer(cbind(New_nests_meg, Available_cavities_prior - New_nests_sum) ~ scale(Meg_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_meg, family = binomial(), control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1000000)))
# 
# plot(simulateResiduals(mod_choice_consp_meg_no_ps)) # great
# check_model(mod_choice_consp_meg_no_ps) # good
# testDispersion(mod_choice_consp_meg_no_ps, type = 'DHARMa') # good 
# testZeroInflation(mod_choice_consp_meg_no_ps) # good
# summary(mod_choice_consp_meg_no_ps) # significant effect of congenerics, sites and day of year
# vif(mod_choice_consp_meg_no_ps) # good
# 
# # plot
# values_meg_no_ps <- seq(min(nesting_meg$Meg_no_ps_prior), 
#           max(nesting_meg$Meg_no_ps_prior), 
#           length.out = 100)
# marginal_effects_meg_no_ps <- ggpredict(mod_choice_consp_meg_no_ps, 
#                terms = "Meg_no_ps_prior [all]", 
#                conditions = list(Meg_no_ps_prior = values_meg_no_ps))
# plot(marginal_effects_meg_no_ps, rawdata = TRUE, jitter = 0.04)












## checking unscaled estimates --- this is old as of May 2025, not sure where these numbers (0.2578, etc.) came from

mean_meg_no_ps_prior <- mean(nesting_meg$Meg_no_ps_prior)
sd_meg_no_ps_prior <- sd(nesting_meg$Meg_no_ps_prior)
scaled_estimate_meg <- 0.2578
unscaled_estimate_meg <- scaled_estimate_meg * sd_meg_no_ps_prior + mean_meg_no_ps_prior
unscaled_estimate_meg

mean_os_no_ps_prior <- mean(nesting_os$Osmia_no_ps_prior)
sd_os_no_ps_prior <- sd(nesting_os$Osmia_no_ps_prior)
scaled_estimate_os <- 0.36480
unscaled_estimate_os <- scaled_estimate_os * mean_os_no_ps_prior + mean_os_no_ps_prior
unscaled_estimate_os

mean_os_ps_prior <- mean(nesting_os$Osmia_ps_prior)
sd_os_ps_prior <- sd(nesting_os$Osmia_ps_prior)
scaled_estimate_os_ps <- 0.85564
unscaled_estimate_os_ps <- scaled_estimate_os_ps * mean_os_ps_prior + mean_os_ps_prior
unscaled_estimate_os_ps


```
