---
title: 'Analyses'
format: 
 html:
  self-contained: true
  toc: true
execute:
  warning: false
editor: visual
---

```{r Load, include=FALSE}

## Loading packages

library(glmmTMB)
library(DHARMa)
library(lme4)
library(car)
library(see)
library(performance)
library(sjPlot) 
library(ggplot2)
library(ggeffects)
library(effects)
library(tidyverse)
library(interactions) 
library(betareg)

# set theme
theme_set(theme_classic(base_size = 13))

## Loading data

  # nest-level data for all cavities
nesting_os <- read_csv('data/nesting_cleaned_osmia.csv')
nesting_meg <- read_csv('data/nesting_cleaned_megachile.csv') 
  # trapnest-level data
trapnesting_os <- read_csv('data/trapnesting_cleaned_osmia.csv')
trapnesting_meg <- read_csv('data/trapnesting_cleaned_megachile.csv')
  # cell-level data
nesting_cells_os <- read_csv('data/nesting_cells_cleaned_osmia.csv')
nesting_cells_meg <- read_csv('data/nesting_cells_cleaned_megachile.csv')
  # nest-level data for investment measures
nesting_investment_os <- read_csv('data/investment_cleaned_osmia.csv')
nesting_investment_meg <- read_csv('data/investment_cleaned_megachile.csv')


## Inspecting data 

# checking form of Day_of_year
plot(I(New_nests_os/ (Osmia_no_ps_prior)) ~ (Day_of_year), data = trapnesting_os)
plot(I(New_nests_meg/ (Meg_no_ps_prior)) ~ (Day_of_year), data = trapnesting_meg)
# inspect correlation for detectable and non detectable parasites
plot(jitter(nesting_os$NDP_os_prior) ~ jitter(nesting_os$DP_os_prior))
cor(nesting_os$NDP_os_prior, nesting_os$DP_os_prior) # good
plot(jitter(nesting_meg$NDP_meg_prior) ~ jitter(nesting_meg$DP_meg_prior))
cor(nesting_meg$NDP_meg_prior, nesting_meg$DP_meg_prior) # good


## Cleaning data to reduce large standard error values from small sample sizes

# how many Osmia in each year?
# note the numbers may vary slightly because of what nests were dissected, moldy, eaten, or did not emerge 
nesting_os %>% 
  group_by(Year) %>% 
  filter(Status != 'pseudo' & Status != 'empty') %>% 
  summarise(Count = length(unique(Nest_ID)))
# 162 in 202023 and only 5 in 2024, remove
nesting_os <- nesting_os %>% filter(Year != 'year_2024')
# same for trapnesting data?
trapnesting_os %>% 
  group_by(Year) %>% 
  summarise(Count = sum(New_nests_os)) # yes, 5
trapnesting_os <- trapnesting_os %>% filter(Year != 'year_2024')
# same for cell data?
nesting_cells_os %>% 
  group_by(Year) %>% 
  summarise(Count = length(unique(Nest_ID))) # yes, 4
nesting_cells_os <- nesting_cells_os %>% filter(Year != 'year_2024')
# same for investment data?
nesting_investment_os %>% 
  group_by(Year) %>% 
  summarise(Count = length(unique(Nest_ID))) # yes, 1
nesting_investment_os <- nesting_investment_os %>% filter(Year != 'year_2024')

# how many Osmia at each site?
nesting_os %>% 
  group_by(Site) %>% 
  filter(Status != 'pseudo' & Status != 'empty') %>% 
  summarise(Count = length(unique(Nest_ID)))
# before previous removal, 109 at BP, 54 at NN, only 4 at CP, remove
nesting_os <- nesting_os %>% filter(Site != 'CP')
# same for trapnesting data?
trapnesting_os %>% 
  group_by(Site) %>% 
  summarise(Count = sum(New_nests_os))
# before previous removal, 109 at BP, 54 at NN, only 4 at CP, remove
trapnesting_os <- trapnesting_os %>% filter(Site != 'CP')
# same for cell data?
nesting_cells_os %>% 
  group_by(Site) %>% 
  summarise(Count = length(unique(Nest_ID))) # yes, 4
nesting_cells_os <- nesting_cells_os %>% filter(Site != 'CP')
# same for investment data?
nesting_investment_os %>% 
  group_by(Site) %>% 
  summarise(Count = length(unique(Nest_ID))) # all good, no CP


# how many Megachile in each year?
nesting_meg %>% 
  group_by(Year) %>% 
  filter(Status != 'pseudo' & Status != 'empty') %>% 
  summarise(Count = length(unique(Nest_ID)))
# 64 in 2023, 76 in 2024, good spread
trapnesting_meg %>% 
  group_by(Year) %>% 
  summarise(Count = sum(New_nests_meg)) # yes, 64 and 76
# same for cell data?
nesting_cells_meg %>% 
  group_by(Year) %>% 
  summarise(Count = length(unique(Nest_ID))) # yes, 44 and 64
# same for investment data?
nesting_investment_meg %>% 
  group_by(Year) %>% 
  summarise(Count = length(unique(Nest_ID))) # fine, so few because of 40+ nests weren't sealed on the last day

# how many Megachile at each site?
nesting_meg %>% 
  group_by(Site) %>% 
  filter(Status != 'pseudo' & Status != 'empty') %>% 
  summarise(Count = length(unique(Nest_ID)))
# BP has 98, CP has 39, only 3 at NN, remove
nesting_meg <- nesting_meg %>% filter(Site != 'NN')
# same with trapnesting?
trapnesting_meg %>% 
  group_by(Site) %>% 
  summarise(Count = sum(New_nests_meg)) # 3 at NN
trapnesting_meg <- trapnesting_meg %>% filter(Site != 'NN')
# same for cell data?
nesting_cells_meg %>% 
  group_by(Site) %>% 
  summarise(Count = length(unique(Nest_ID))) # 2 at NN
nesting_cells_meg <- nesting_cells_meg %>% filter(Site != 'NN')
# same for investment data?
nesting_investment_meg %>% 
  group_by(Site) %>% 
  summarise(Count = length(unique(Nest_ID))) # good


```

```{r Prep_for_Simulations, include=FALSE}


## Create the function to simulate data

# simulate the null hypothesis that bees are treating each cavity equally, thus nesting would be based just on availability
# simulation without replacement so that if there are multiple new osmia nests, the number of available cavities goes down by one for each sequential new nest in that subsite-group combo

# create a function to simulate the distribution of nests based on the null hypothesis that bees are nesting according to available cavities 
simulate_nests <- function(avail_cavities, total_nests) {
  n <- length(avail_cavities) # number of available cavities
  nest_counts <- rep(0, n) # start with zero nests in each trapnest

  # keep as zeros is no nests are built
  if (total_nests == 0) {
    return(nest_counts)
  }

  # look through each of the new nests to be randomly assigned
  for (i in 1:total_nests) {
    if (sum(avail_cavities) == 0) break  # stop if all cavities used
    probability <- avail_cavities / sum(avail_cavities) # calculate probability proportional to the number of available cavities
    chosen <- sample(1:n, size = 1, prob = probability) # randomly choose a trapnest for the new nest based on available cavities
    nest_counts[chosen] <- nest_counts[chosen] + 1 # record it
    avail_cavities[chosen] <- avail_cavities[chosen] - 1 # reduce the number of available cavities by 1, when we apply the function the loop will run again until no new nests are left
  }
  return(nest_counts) # return the vector with how many nests were randomly assigned
}

# indicate how many simulated data columns we want for the model
n_sim <- 100 

```

## Congeneric Attraction - Osmia

Model 1_i - including pseudo

Model 1_ii - excluding pseudo

```{r Models_1_Congeneric_Attraction_Osmia, include=TRUE}


## Apply the function

trapnesting_os_null <- trapnesting_os %>%
  group_by(Subsite, Date) %>%
  group_modify(~ { # simulate random nest distribution for subsite-date combo
    for (i in 1:n_sim){ # make each of the columns we indicated that we want 
    nest_counts <- simulate_nests(.x$Available_cavities_prior, sum(.x$New_nests_os)) # use available cavities as weights and sum the new nests
    .x[[paste0('Null_new_nests', i)]] <- nest_counts # make the new column
    }
    .x # return the modified group
  }
  ) %>%
  ungroup()


# MODEL 1.i - Osmia including pseudo #

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
sim_list_os_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Osmia_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_os_ps <- c(unlist(lapply(sim_list_os_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_choice_consp_os_ps <- glmer(New_nests_os ~ scale(Osmia_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_choice_consp_os_ps)

# extract my observed coefficient
obvs_coef_os_ps <- summary(mod_choice_consp_os_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_os_ps, c(0.025, 0.975))
obvs_coef_os_ps # falls outside of range

# convert to dataframe
df_coefs_os_ps <- data.frame(sim_coefs_os_ps = sim_coefs_os_ps)

hist_1i <- ggplot(df_coefs_os_ps, aes(x = sim_coefs_os_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_os_ps, color = 'red2', linetype = 'dashed', size = 1) + 
  labs(y = 'Count', x = 'Coefficient')
# reject the null - Osmia are attracted to congenerics 


## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_os_ps < df_coefs_os_ps) # none are larger
# save
ggsave("Figures/hist_1i.png", plot = hist_1i, width = 4.5, height = 3.5)


# MODEL 1.ii - Osmia excluding pseudo #

## Simulated data 

# run the model on each of the simulated data columns
sim_list_os_no_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_os_no_ps <- c(unlist(lapply(sim_list_os_no_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Observed data 

mod_choice_consp_os_no_ps <- glmer(New_nests_os ~ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_choice_consp_os_no_ps)

# extract my observed coefficient
obvs_coef_os_no_ps <- summary(mod_choice_consp_os_no_ps)$coefficients[2,1]

# does my observed value fall outside this range, and we reject the null?
quantile(sim_coefs_os_no_ps, c(0.025, 0.975))
obvs_coef_os_no_ps # falls outside of range

# convert to dataframe
df_coefs_os_no_ps <- data.frame(sim_coefs_os_no_ps = sim_coefs_os_no_ps)

# plot
hist_1ii <- ggplot(df_coefs_os_no_ps, aes(x = sim_coefs_os_no_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_os_no_ps, color = 'red2', linetype = 'dashed', size = 1) +
  labs(y = 'Count', x = 'Coefficient')
# reject the null - Osmia are attracted to congenerics


## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_os_no_ps < df_coefs_os_no_ps) # none are larger
# save
ggsave("Figures/hist_1ii.png", plot = hist_1ii, width = 4.5, height = 3.5)

# sample size
sum(trapnesting_os_null$New_nests_os)

```

## Congeneric Attraction - Megachile

Model 1_iii - including pseudo

Model 1_iv - excluding pseudo

```{r Models_1_Congeneric_Attraction_Megachile, include=TRUE}


## Apply the function

trapnesting_meg_null <- trapnesting_meg %>%
  group_by(Subsite, Date) %>%
  group_modify(~ { # simulate random nest distribution for subsite-date combo
    for (i in 1:n_sim){ # make each of the columns we indicated that we want 
    nest_counts <- simulate_nests(.x$Available_cavities_prior, sum(.x$New_nests_meg)) # use available cavities as weights and sum the new nests
    .x[[paste0('Null_new_nests', i)]] <- nest_counts # make the new column
    }
    .x # return the modified group
  }
  ) %>%
  ungroup()


# MODEL 1.iii - Megachile including pseudo #

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns
sim_list_meg_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Meg_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_meg_ps <- c(unlist(lapply(sim_list_meg_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))

## Making the model part 2: Observed data 

mod_choice_consp_meg_ps <- glmer(New_nests_meg ~ scale(Meg_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
summary(mod_choice_consp_meg_ps)

# extract my observed coefficient
obvs_coef_meg_ps <- summary(mod_choice_consp_meg_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_meg_ps, c(0.025, 0.975))
obvs_coef_meg_ps # not within this range

# convert to dataframe
df_coefs_meg_ps <- data.frame(sim_coefs_meg_ps = sim_coefs_meg_ps)

hist_1iii <- ggplot(df_coefs_meg_ps, aes(x = sim_coefs_meg_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_meg_ps, color = 'red2', linetype = 'dashed', size = 1) +
  labs(y = 'Count', x = 'Coefficient')
# reject the null - Megachile are attracted to congenerics 


## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_meg_ps < df_coefs_meg_ps) # none are larger
# save
ggsave("Figures/hist_1iii.png", plot = hist_1iii, width = 4.5, height = 3.5)


# MODEL 1.iv - Megachile excluding pseudo #

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns
sim_list_meg_no_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Meg_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_meg_no_ps <- c(unlist(lapply(sim_list_meg_no_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_choice_consp_meg_no_ps <- glmer(New_nests_meg ~ scale(Meg_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
summary(mod_choice_consp_meg_no_ps)

# extract my observed coefficient
obvs_coef_meg_no_ps <- summary(mod_choice_consp_meg_no_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_meg_no_ps, c(0.025, 0.975))
obvs_coef_meg_no_ps # falls outside this range

# convert to dataframe
df_coefs_meg_no_ps <- data.frame(sim_coefs_meg_no_ps = sim_coefs_meg_no_ps)

hist_1iv <- ggplot(df_coefs_meg_no_ps, aes(x = sim_coefs_meg_no_ps)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_meg_no_ps, color = 'red2', linetype = 'dashed', size = 1) +
  labs(y = 'Count', x = 'Coefficient')
# reject the null - Megachile are attracted to congenerics 


## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_meg_no_ps < df_coefs_meg_no_ps) # none are larger
# save
ggsave("Figures/hist_1iv.png", plot = hist_1iv, width = 4.5, height = 3.5)

# sample size
sum(trapnesting_meg_null$New_nests_meg)

```

## Parasite Preferences - Osmia (Sapyga)

Model 2_i - including pseudo

Model 2_ii - excluding pseudo

Model 2_iii - active nests

```{r Models_2_Parasites_Attraction_Osmia, include=TRUE}

# Relation between parasites and aggregation densities


# MODEL 2_i - Osmia parasite aggregation preferences, including pseudo #

mod_2i_par_agg_os_ps <- glmer(Cell_parasitized ~ scale(Osmia_ps) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_2i_par_agg_os_ps)) # good
check_model(mod_2i_par_agg_os_ps) # good
testDispersion(mod_2i_par_agg_os_ps, type = 'DHARMa') # good
testZeroInflation(mod_2i_par_agg_os_ps) # good
summary(mod_2i_par_agg_os_ps) ### 
vif(mod_2i_par_agg_os_ps) # good

# plot
values_osmia_ps <- seq(min(nesting_cells_os$Osmia_ps, na.rm = TRUE), 
               max(nesting_cells_os$Osmia_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_os_ps <- ggpredict(mod_2i_par_agg_os_ps, 
               terms = 'Osmia_ps [all]', 
               conditions = list(Osmia_ps = values_osmia_ps))
plot(marginal_effects_os_ps)  # trend for MORE parasitism with increasing osmia numbers


# MODEL 2_ii - Osmia parasite aggregation preferences, excluding pseudo #

mod_2ii_par_agg_os_no_ps <- glmer(Cell_parasitized ~ scale(Osmia_no_ps) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_2ii_par_agg_os_no_ps)) # good
check_model(mod_2ii_par_agg_os_no_ps) # good
testDispersion(mod_2ii_par_agg_os_no_ps, type = 'DHARMa') # good
testZeroInflation(mod_2ii_par_agg_os_no_ps) # good
summary(mod_2ii_par_agg_os_no_ps) # all effects significant 
vif(mod_2ii_par_agg_os_no_ps)

# plot
values_osmia_no_ps <- seq(min(nesting_cells_os$Osmia_no_ps, na.rm = TRUE), 
               max(nesting_cells_os$Osmia_no_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_os_no_ps <- ggpredict(mod_2ii_par_agg_os_no_ps, 
               terms = 'Osmia_no_ps [all]', 
               conditions = list(Osmia_no_ps = values_osmia_no_ps))
plot(marginal_effects_os_no_ps) # trend with less parasitism for more osmia 


# MODEL 2_iii - Osmia parasite aggregation preferences for active nests #

mod_2iii_par_agg_os_active <- glmer(Cell_parasitized ~ scale(Osmia_active_mothers) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_2iii_par_agg_os_active)) # good
check_model(mod_2iii_par_agg_os_active) # good
testDispersion(mod_2iii_par_agg_os_active, type = 'DHARMa') # good
testZeroInflation(mod_2iii_par_agg_os_active) # good
summary(mod_2iii_par_agg_os_active) # effect of day of year
vif(mod_2iii_par_agg_os_active) # good

# plot
values_osmia_active <- seq(min(nesting_cells_os$Osmia_active_mothers, na.rm = TRUE), 
               max(nesting_cells_os$Osmia_active_mothers, na.rm = TRUE), 
               length.out = 100)
marginal_effects_os_actives <- ggpredict(mod_2iii_par_agg_os_active, 
               terms = 'Osmia_active_mothers [all]', 
               conditions = list(Osmia_active_mothers = values_osmia_active))
plot(marginal_effects_os_actives) # trend with less parasitism for more active osmia mothers


# save models
saveRDS(mod_2i_par_agg_os_ps, file = 'Models/mod_2i_par_agg_os_ps.rds')
saveRDS(mod_2ii_par_agg_os_no_ps, file = 'Models/mod_2ii_par_agg_os_no_ps.rds')
saveRDS(mod_2iii_par_agg_os_active, file = 'Models/mod_2iii_par_agg_os_active.rds')

# sample size
table(nesting_cells_os$Cell_parasitized) # 174 + 112 = 286


```

## Parasite Preferences - Megachile (Meloid, Coelioxys, Chalcid)

Model 2_iv - including pseudo

Model 2_v - excluding pseudo

Model 2_vi - active nests

```{r Models_2_Parasites_Attraction_Megachile, include=TRUE}

# Relation between parasites and aggregation densities (small sample size for Megachile)

# MODEL 2_iv - Megachile parasite aggregation preferences, including pseudo #

mod_2iv_par_agg_meg_ps <- glmer(Cell_parasitized ~ scale(Meg_ps) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_meg, family = binomial)

plot(simulateResiduals(mod_2iv_par_agg_meg_ps)) # fine
check_model(mod_2iv_par_agg_meg_ps) # fine
testDispersion(mod_2iv_par_agg_meg_ps, type = 'DHARMa') # good
testZeroInflation(mod_2iv_par_agg_meg_ps) # good
summary(mod_2iv_par_agg_meg_ps) # no detected effects
vif(mod_2iv_par_agg_meg_ps) # good

# plot
values_meg_ps <- seq(min(nesting_cells_meg$Meg_ps, na.rm = TRUE), 
               max(nesting_cells_meg$Meg_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_meg_ps <- ggpredict(mod_2iv_par_agg_meg_ps, 
               terms = 'Meg_ps [all]', 
               conditions = list(Meg_ps = values_meg_ps))
plot(marginal_effects_meg_ps) # no trend


# MODEL 2_v - Megachile parasite aggregation preferences, excluding pseudo #

mod_2v_par_agg_meg_no_ps <- glmer(Cell_parasitized ~ scale(Meg_no_ps) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_meg, family = binomial)

plot(simulateResiduals(mod_2v_par_agg_meg_no_ps)) # good
check_model(mod_2v_par_agg_meg_no_ps) # fine
testDispersion(mod_2v_par_agg_meg_no_ps, type = 'DHARMa') # good
testZeroInflation(mod_2v_par_agg_meg_no_ps) # good
summary(mod_2v_par_agg_meg_no_ps) # no detected effects
vif(mod_2v_par_agg_meg_no_ps) # good

# plot
values_meg_no_ps <- seq(min(nesting_cells_meg$Meg_no_ps, na.rm = TRUE), 
               max(nesting_cells_meg$Meg_no_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_meg_no_ps <- ggpredict(mod_2v_par_agg_meg_no_ps, 
               terms = 'Meg_no_ps [all]', 
               conditions = list(Meg_no_ps = values_meg_no_ps))
plot(marginal_effects_meg_no_ps) # no trend 


# MODEL 2_vi - Megachile parasite aggregation preferences for active nests #

mod_2vi_par_agg_meg_active <- glmer(Cell_parasitized ~ scale(Meg_active_mothers) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_meg, family = binomial)

plot(simulateResiduals(mod_2vi_par_agg_meg_active)) # fine
check_model(mod_2vi_par_agg_meg_active) # fine
testDispersion(mod_2vi_par_agg_meg_active, type = 'DHARMa') # good
testZeroInflation(mod_2vi_par_agg_meg_active) # good 
summary(mod_2vi_par_agg_meg_active) # no detected effects, close for active mothers
vif(mod_2vi_par_agg_meg_active) # good

# plot
values_meg_active <- seq(min(nesting_cells_meg$Meg_active_mothers, na.rm = TRUE), 
               max(nesting_cells_meg$Meg_active_mothers, na.rm = TRUE), 
               length.out = 100)
marginal_effects_meg_active <- ggpredict(mod_2vi_par_agg_meg_active, 
               terms = 'Meg_active_mothers [all]', 
               conditions = list(Meg_active_mothers = values_meg_active))
plot(marginal_effects_meg_active) # trend for more parasites with less active megachile


# save models
saveRDS(mod_2iv_par_agg_meg_ps, file = 'Models/mod_2iv_par_agg_meg_ps.rds')
saveRDS(mod_2v_par_agg_meg_no_ps, file = 'Models/mod_2v_par_agg_meg_no_ps.rds')
saveRDS(mod_2vi_par_agg_meg_active, file = 'Models/mod_2vi_par_agg_meg_active.rds')

# sample size
table(nesting_cells_meg$Cell_parasitized) # 123 + 19 = 142

```

## Parasite Avoidance - Osmia

Model 3a_i - including pseudo

Model 3a_ii - excluding pseudo

```{r Models_3_Parasite_Avoidance_Osmia, include=TRUE}

# MODEL 3a_i - Osmia avoidance of parasites, including pseudo #

# run the model on each of the simulated data columns
sim_list_os_ps_avoid <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(DP_os_prior) * scale(NDP_ps_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson())
  summary(output)
} )

# take all of the simulated coefficients
sim_coefs_os_ps_avoid <- c(unlist(lapply(sim_list_os_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_choice_consp_par_os_ps <- glmer(New_nests_os ~ scale(DP_os_prior) * scale(NDP_ps_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_choice_consp_par_os_ps)

# extract my observed coefficient
obvs_coef_os_ps_avoid <- summary(mod_choice_consp_par_os_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_os_ps_avoid, c(0.025, 0.975))
obvs_coef_os_ps_avoid # close but falls within this range

# Convert to dataframe
df_coefs_os_ps_avoid <- data.frame(sim_coefs_os_ps_avoid = sim_coefs_os_ps_avoid)

hist_3ai <- ggplot(df_coefs_os_ps_avoid, aes(x = sim_coefs_os_ps_avoid)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_os_ps_avoid, color = 'red2', linetype = 'dashed', size = 1) +
  labs(y = 'Count', x = 'Coefficient')
# fail to reject the null - no evidence for Osmia avoiding nests with parasites


## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_os_ps_avoid < df_coefs_os_ps_avoid) # 15 are larger
# save
ggsave("Figures/hist_3ai.png", plot = hist_3ai, width = 4.5, height = 3.5)

# plot
plot(ggpredict(mod_choice_consp_par_os_ps, terms = c('DP_os_prior', 'NDP_ps_os_prior'))) 


# MODEL 3a_ii - Osmia avoidance of parasites, excluding pseudo #

# run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
sim_list_os_no_ps_avoid <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(DP_os_prior) * scale(NDP_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson())
  summary(output)
} )

# take all of the simulated coefficients
sim_coefs_os_no_ps_avoid <- c(unlist(lapply(sim_list_os_no_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_choice_consp_par_os_no_ps <- glmer(New_nests_os ~ scale(DP_os_prior) * scale(NDP_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_choice_consp_par_os_no_ps)

# extract my observed coefficient
obvs_coef_os_no_ps_avoid <- summary(mod_choice_consp_par_os_no_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_os_no_ps_avoid, c(0.025, 0.975))
obvs_coef_os_no_ps_avoid # falls outside this range

# Convert to dataframe
df_coefs_os_no_ps_avoid <- data.frame(sim_coefs_os_no_ps_avoid = sim_coefs_os_no_ps_avoid)

hist_3aii <- ggplot(df_coefs_os_no_ps_avoid, aes(x = sim_coefs_os_no_ps_avoid)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_os_no_ps_avoid, color = 'red2', linetype = 'dashed', size = 1) +
  labs(y = 'Count', x = 'Coefficient')
# small effect but reject the null - Osmia avoid nesting near parasites when pseudo are excluded


## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_os_no_ps_avoid < df_coefs_os_no_ps_avoid) # 1 is larger
# save
ggsave("Figures/hist_3aii.png", plot = hist_3aii, width = 4.5, height = 3.5)

# plot
plot(ggpredict(mod_choice_consp_par_os_no_ps, terms = c('DP_os_prior', 'NDP_os_prior'))) 

# sample size
sum(trapnesting_os_null$New_nests_os)

```

## Parasite Avoidance - Megachile

Model 3a_iii - including pseudo

Model 3a_iv - excluding pseudo

```{r Models_3_Parasite_Avoidance_Megachile, include=TRUE}


# MODEL 3a_iii - Megachile avoidance of parasites, including pseudo #

# run the model on each of the simulated data columns
sim_list_meg_ps_avoid <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(DP_meg_prior) * scale(NDP_ps_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson())
  summary(output)
} )

# take all of the simulated coefficients
sim_coefs_meg_ps_avoid <- c(unlist(lapply(sim_list_meg_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))

## Making the model part 2: Observed data 

mod_choice_consp_par_meg_ps <- glmer(New_nests_meg ~ scale(DP_meg_prior) * scale(NDP_ps_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
summary(mod_choice_consp_par_meg_ps)

# extract my observed coefficient
obvs_coef_meg_ps_avoid <- summary(mod_choice_consp_par_meg_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_meg_ps_avoid, c(0.025, 0.975))
obvs_coef_meg_ps_avoid # falls within this range

# Convert to dataframe
df_coefs_meg_ps_avoid <- data.frame(sim_coefs_meg_ps_avoid = sim_coefs_meg_ps_avoid)

hist_3aiii <- ggplot(df_coefs_meg_ps_avoid, aes(x = sim_coefs_meg_ps_avoid)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = obvs_coef_meg_ps_avoid, color = 'red2', linetype = 'dashed', size = 1) +
  labs(y = 'Count', x = 'Coefficient')
# fail to reject the null - Megachile do not avoid nesitng near parasites


## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_meg_ps_avoid < df_coefs_meg_ps_avoid) # 49 is larger
# save
ggsave("Figures/hist_3aiii.png", plot = hist_3aiii, width = 4.5, height = 3.5)

# sample size
sum(trapnesting_meg_null$New_nests_meg)

# MODEL 3a_iv - Megachile avoidance of parasites, excluding pseudo #

# running the simulation gets the error 'Error: PIRLS loop resulted in NaN value'
# too little data to run it

# # run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
# sim_list_meg_no_ps_avoid <- lapply(1:n_sim, function(x){
# my_formula <- paste0('Null_new_nests', x, ' ~ scale(DP_meg_prior) * scale(NDP_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)')
# output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson())
# summary(output)
# } )
#
# # take all of the simulated coefficients
# sim_coefs_meg_no_ps_avoid <- c(unlist(lapply(sim_list_meg_no_ps_avoid, function(x) {
# x$coefficients[2,1] # extract second row, first column for the estimate
# })))
#
#
# ## Making the model part 2: Observed data
#
# mod_choice_consp_par_meg_no_ps <- glmer(New_nests_meg ~ scale(DP_meg_prior) * scale(NDP_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
# summary(mod_choice_consp_par_meg_no_ps)
#
# # extract my observed coefficient
# obvs_coef_meg_no_ps_avoid <- summary(mod_choice_consp_par_meg_no_ps)$coefficients[2,1]
#
# # plot my observed values with the simulated coefficients
# # abline(v=obvs_coef_meg_no_ps_avoid)
#
# # range of 95%, i.e., if my observed value falls within this range, its not different from the null
# quantile(sim_coefs_meg_no_ps_avoid, c(0.025, 0.975))
#
# # Convert to dataframe
# df_coefs_meg_no_ps_avoid <- data.frame(sim_coefs_meg_no_ps_avoid = sim_coefs_meg_no_ps_avoid)
#
# ggplot(df_coefs_meg_no_ps_avoid, aes(x = sim_coefs_meg_no_ps_avoid)) +
# geom_histogram(binwidth = 0.1) +
# geom_vline(xintercept = obvs_coef_meg_no_ps_avoid, color = 'red2', linetype = 'dashed', size = 1)

```

## Parasite Defences

Model 3b_i - length of straw filled, Osmia

Model 3b_ii - length of straw filled, Megachile

Model 3c_i - nest termination, Osmia

Model 3c_ii - nest termination, Megachile

Model 3d_i - abandoned inner wall, Osmia

Model 3d_ii - abandoned inner wall, Megachile

```{r Models_3b_Nest_Investment, include=TRUE}

# MODEL 3b_i - Osmia investment #
## try beta-binomial distribution with glmTMB

# fit as guassion, total-straw length as covariate in the model as fixed 


# Investment into a nest if other active nests or the nest itself is parasitized
library(lmerTest)

mod_3bi_invest_filled_os <- lmer(Filled_length ~ scale(DP_os_prior) * Nest_parasitized + Straw_length + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_investment_os)

check_model(mod_3bi_invest_filled_os) # good
summary(mod_3bi_invest_filled_os) # significant effect of if focal nest is parasitized, day of year, and year
vif(mod_3bi_invest_filled_os) # good

plot(ggpredict(mod_3bi_invest_filled_os, terms = c('DP_os_prior', 'Nest_parasitized'))) 


# MODEL 3b_ii - Megachile investment #

# Investment into a nest if other active nests or the nest itself is parasitized

mod_3bii_invest_filled_meg <- lmer(Filled_length ~ scale(DP_meg_prior) * Nest_parasitized + Straw_length + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = nesting_investment_meg)

check_model(mod_3bii_invest_filled_meg) # good
summary(mod_3bii_invest_filled_meg) # significant effect of day of year 
vif(mod_3bii_invest_filled_meg) # good

plot(ggpredict(mod_3bii_invest_filled_meg, terms = c('DP_meg_prior', 'Nest_parasitized'))) 


# save models
saveRDS(mod_3bi_invest_filled_os, file = 'Models/mod_3bi_invest_filled_os.rds')
saveRDS(mod_3bii_invest_filled_meg, file = 'Models/mod_3bii_invest_filled_meg.rds')

# sample size
nesting_investment_os %>%
  filter(!is.na(Nest_parasitized)) %>%
  summarise(n_unique_nests = n_distinct(Nest_ID)) # 74
nesting_investment_meg %>%
  filter(!is.na(Nest_parasitized)) %>%
  summarise(n_unique_nests = n_distinct(Nest_ID)) # 40

```

```{r Models_3c_Nest_Termination, include=TRUE}

# Remove NAs from unknown cell positions
nesting_cells_os_noNA <- nesting_cells_os %>% filter(!is.na(Cell_position_bi))
nesting_cells_meg_noNA <- nesting_cells_meg %>% filter(!is.na(Cell_position_bi))

# MODEL 3c_i - Osmia nest termination #

mod_3ci_last_cell_os <- glmer(Cell_position_bi ~ Cell_parasitized + poly(Day_of_year,2) + Site + (1|Nest_ID), data = nesting_cells_os_noNA, family = binomial)

plot(simulateResiduals(mod_3ci_last_cell_os)) # good
check_model(mod_3ci_last_cell_os) # good
vif(mod_3ci_last_cell_os) # good
summary(mod_3ci_last_cell_os) # significant effect of day of year

ggpredict(mod_3ci_last_cell_os, terms = 'Cell_parasitized [all]') %>%
  plot(jitter = .04)


# MODEL 3c_ii - Megachile nest termination #

mod_3cii_last_cell_meg <- glmer(Cell_position_bi ~ Cell_parasitized + poly(Day_of_year,2) + Year + Site + (1|Nest_ID), data = nesting_cells_meg_noNA, family = binomial)

plot(simulateResiduals(mod_3cii_last_cell_meg)) # good
check_model(mod_3cii_last_cell_meg) # good
summary(mod_3cii_last_cell_meg) # significant effect of day of year
vif(mod_3cii_last_cell_meg) # good

ggpredict(mod_3cii_last_cell_meg, terms = 'Cell_parasitized [all]') %>%
  plot(jitter = .04)
# slight trend for more likely to best last cell if its parasitized


# save models
saveRDS(mod_3ci_last_cell_os, file = 'Models/mod_3ci_last_cell_os.rds')
saveRDS(mod_3cii_last_cell_meg, file = 'Models/mod_3cii_last_cell_meg.rds')

# sample size
length(unique(nesting_cells_os_noNA$Nest_ID)) # 75 nests
length(unique(nesting_cells_meg_noNA$Nest_ID)) # 52 nests

# sample size including unknown cell positions
length(unique(nesting_cells_os$Nest_ID)) # 81 nests
length(unique(nesting_cells_meg$Nest_ID)) # 56 nests

```

```{r Models_3d_Nest_Abandonment, include=TRUE}

# MODEL 3d_i - Osmia abandoned inner walls including pseudo #

mod_3di_inner_wall_os <- glmer(Abandoned_inner_wall ~ scale(DP_os_prior) * scale(NDP_ps_os_prior) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_os, family = binomial)

plot(simulateResiduals(mod_3di_inner_wall_os)) # good
check_model(mod_3di_inner_wall_os) # good
summary(mod_3di_inner_wall_os) # significant effect of day of year
vif(mod_3di_inner_wall_os) # good

ggpredict(mod_3di_inner_wall_os, terms = 'DP_os_prior [all]') %>%
  plot(jitter = .04)


# MODEL 3d_ii - Megachile abandoned inner walls #

# nesting_meg %>% filter(Abandoned_inner_wall == 1) # no parasites when walls are abandoned


# save models
saveRDS(mod_3di_inner_wall_os, file = 'Models/mod_3di_inner_wall_os.rds')
# saveRDS(mod_3dii_inner_wall_meg, file = 'Models/mod_3dii_inner_wall_meg.rds')

# sample size
nesting_os %>% filter(!is.na(Abandoned_inner_wall)) # 168
nesting_os %>%
  summarise(Count = sum(Abandoned_inner_wall, na.rm = TRUE)) # 28
nesting_meg %>% filter(!is.na(Abandoned_inner_wall)) # 145
nesting_meg %>%
  summarise(Count = sum(Abandoned_inner_wall, na.rm = TRUE)) # 13


```
