---
title: 'Analyses'
format: 
 html:
  self-contained: true
  toc: true
execute:
  warning: false
editor: visual
---

```{r Load, include=FALSE}

## Loading packages

library(tidyverse)
library(glmmTMB)
library(DHARMa)
library(lme4)
library(performance) # for check_model
library(car) # for vif(), just to double check check_model vif plots
library(patchwork) # for adding histomgram panels into one figure
library(ggeffects) # for ggpredict


# set theme
theme_set(theme_classic(base_size = 13))

## Loading data

  # nest-level data for all cavities
nesting_os <- read_csv('data/nesting_cleaned_osmia.csv')
nesting_meg <- read_csv('data/nesting_cleaned_megachile.csv') 
  # trapnest-level data
trapnesting_os <- read_csv('data/trapnesting_cleaned_osmia.csv')
trapnesting_meg <- read_csv('data/trapnesting_cleaned_megachile.csv')
  # cell-level data
nesting_cells_os <- read_csv('data/nesting_cells_cleaned_osmia.csv')
nesting_cells_meg <- read_csv('data/nesting_cells_cleaned_megachile.csv')
  # nest-level data for investment measures
nesting_investment_os <- read_csv('data/investment_cleaned_osmia.csv')
nesting_investment_meg <- read_csv('data/investment_cleaned_megachile.csv')


## Inspecting data 

# checking form of Day_of_year
plot(I(New_nests_os/ (Osmia_no_ps_prior)) ~ (Day_of_year), data = trapnesting_os)
plot(I(New_nests_meg/ (Meg_no_ps_prior)) ~ (Day_of_year), data = trapnesting_meg)
# inspect correlation for detectable and non detectable parasites
plot(jitter(nesting_os$NDP_os_prior) ~ jitter(nesting_os$DP_os_prior))
cor(nesting_os$NDP_os_prior, nesting_os$DP_os_prior) # good
plot(jitter(nesting_meg$NDP_meg_prior) ~ jitter(nesting_meg$DP_meg_prior))
cor(nesting_meg$NDP_meg_prior, nesting_meg$DP_meg_prior) # good


```

```{r Cleaning, include=FALSE}


## Cleaning data to reduce large standard error values from small sample sizes

# how many Osmia in each year?
# note the numbers may vary slightly because of what nests were dissected, moldy, eaten, or did not emerge 
nesting_os %>% 
  group_by(Year) %>% 
  filter(Status != 'pseudo' & Status != 'empty') %>% 
  summarise(Count = length(unique(Nest_ID)))
# 162 in 202023 and only 5 in 2024, remove
nesting_os <- nesting_os %>% filter(Year != 'year_2024')
# same for trapnesting data?
trapnesting_os %>% 
  group_by(Year) %>% 
  summarise(Count = sum(New_nests_os)) # yes, 5
trapnesting_os <- trapnesting_os %>% filter(Year != 'year_2024')
# same for cell data?
nesting_cells_os %>% 
  group_by(Year) %>% 
  summarise(Count = length(unique(Nest_ID))) # yes, 4
nesting_cells_os <- nesting_cells_os %>% filter(Year != 'year_2024')
# same for investment data?
nesting_investment_os %>% 
  group_by(Year) %>% 
  summarise(Count = length(unique(Nest_ID))) # yes, 1
nesting_investment_os <- nesting_investment_os %>% filter(Year != 'year_2024')

# how many Osmia at each site?
nesting_os %>% 
  group_by(Site) %>% 
  filter(Status != 'pseudo' & Status != 'empty') %>% 
  summarise(Count = length(unique(Nest_ID)))
# after previous removal, 107 at BP, 54 at NN, only 1 at CP, remove CP
nesting_os <- nesting_os %>% filter(Site != 'CP')
# same for trapnesting data?
trapnesting_os %>% 
  group_by(Site) %>% 
  summarise(Count = sum(New_nests_os))
# before previous removal, 107 at BP, 54 at NN, only 1 at CP, remove CP
trapnesting_os <- trapnesting_os %>% filter(Site != 'CP')
# same for cell data?
nesting_cells_os %>% 
  group_by(Site) %>% 
  summarise(Count = length(unique(Nest_ID))) # yes, 4
nesting_cells_os <- nesting_cells_os %>% filter(Site != 'CP')
# same for investment data?
nesting_investment_os %>% 
  group_by(Site) %>% 
  summarise(Count = length(unique(Nest_ID))) # all good, no CP

# how many Megachile in each year?
nesting_meg %>% 
  group_by(Year) %>% 
  filter(Status != 'pseudo' & Status != 'empty') %>% 
  summarise(Count = length(unique(Nest_ID)))
# 64 in 2023, 76 in 2024, good spread
trapnesting_meg %>% 
  group_by(Year) %>% 
  summarise(Count = sum(New_nests_meg)) # yes, 64 and 76
# same for cell data?
nesting_cells_meg %>% 
  group_by(Year) %>% 
  summarise(Count = length(unique(Nest_ID))) # yes, 44 and 64
# same for investment data?
nesting_investment_meg %>% 
  group_by(Year) %>% 
  summarise(Count = length(unique(Nest_ID))) # fine, so few because of 40+ nests weren't sealed on the last day

# how many Megachile at each site?
nesting_meg %>% 
  group_by(Site) %>% 
  filter(Status != 'pseudo' & Status != 'empty') %>% 
  summarise(Count = length(unique(Nest_ID)))
# BP has 98, CP has 39, only 3 at NN, remove
nesting_meg <- nesting_meg %>% filter(Site != 'NN')
# same with trapnesting?
trapnesting_meg %>% 
  group_by(Site) %>% 
  summarise(Count = sum(New_nests_meg)) # 3 at NN
trapnesting_meg <- trapnesting_meg %>% filter(Site != 'NN')
# same for cell data?
nesting_cells_meg %>% 
  group_by(Site) %>% 
  summarise(Count = length(unique(Nest_ID))) # 2 at NN
nesting_cells_meg <- nesting_cells_meg %>% filter(Site != 'NN')
# same for investment data?
nesting_investment_meg %>% 
  group_by(Site) %>% 
  summarise(Count = length(unique(Nest_ID))) # good

nesting_cells_meg %>% filter(Cell_occupant == 'osmia')
```

```{r Descriptive, include=FALSE}


# ## How many cells developed? 
# table(nesting_cells_os$Cell_occupant)
# table(nesting_cells_meg$Cell_occupant)

nesting_cells_os %>% filter(Cell_occupant == 'megachile' | Cell_occupant == 'osmia' | Cell_occupant == 'parasite')
length(nesting_cells_os$Cell_occupant)
# 287 / 749 were identifiable

nesting_cells_meg %>% filter(Cell_occupant == 'megachile' | Cell_occupant == 'osmia' | Cell_occupant == 'parasite')
length(nesting_cells_meg$Cell_occupant)
# 142 / 424 were identifiable
# (286 + 142) / (723 + 408)

# how many nests in the emergence data
length(unique(nesting_cells_os$Nest_ID)) # 82
length(unique(nesting_cells_meg$Nest_ID)) # 56


## How many inner wall only nests, and how many were considered empty?

nesting_os %>% 
  filter(!is.na(Abandoned_inner_wall) & Status == 'empty' | Status == 'inner wall' | Status == 'inner wall + end cap') %>% 
  group_by(Abandoned_inner_wall, Status) %>% 
  summarise(count = length(unique(Nest_ID)))
# 7 osmia considered empty 
# 21 osmia considered bee choices that never developed into nests
nesting_meg %>% 
  filter(!is.na(Abandoned_inner_wall) & Status == 'empty' | Status == 'inner wall' | Status == 'inner wall + end cap') %>% 
  group_by(Abandoned_inner_wall, Status) %>% 
  summarise(count = length(unique(Nest_ID)))
# 8 megachile considered empty 
# 5 megachile considered bee choices that never developed into nests (the NA is a duplicate of one already counted)


## How many nests were parasitized?

nesting_cells_os_short <- nesting_cells_os %>% select(Treatment, Trapnest, Nest_ID, Cell_ID, Date, Cell_occupant, Species_or_subtype, Cell_parasitized, Nest_parasitized)

nesting_cells_os_short %>%
  group_by(Nest_ID) %>%
  summarise(n_types = n_distinct(Nest_parasitized, na.rm = TRUE)) %>%
  filter(n_types > 1)
# some nests have both parasitized and not parasitized, because they became parasitized later one: 13_12a, 13_8a, 25_6b, 36_6a, 6_15a
# so arrange and remove if parasitized if they say both

nesting_cells_os_short <- nesting_cells_os_short %>%
  group_by(Nest_ID) %>%
  arrange(desc(Cell_occupant == "parasite")) %>%
  slice(1) %>%
  ungroup()

table(nesting_cells_os_short$Cell_occupant)
  
nesting_cells_os_short %>% 
  group_by(Species_or_subtype) %>% 
  summarise(count =  length(unique(Nest_ID)))

# osmia pumila	  6			
# sapyga	        52			
# small parasite	3			
# osmia tersula	  9			
# osmia         	12


nesting_cells_meg_short <- nesting_cells_meg %>% select(Treatment, Trapnest, Nest_ID, Cell_ID, Date, Cell_occupant, Species_or_subtype, Cell_parasitized, Nest_parasitized)

nesting_cells_meg_short %>%
  group_by(Nest_ID) %>%
  summarise(n_types = n_distinct(Nest_parasitized, na.rm = TRUE)) %>%
  filter(n_types > 1)
# 98_16a has both parasitized and not parasitized
# so arrange and remove if parasitized 

nesting_cells_meg_short <- nesting_cells_meg_short %>%
  group_by(Nest_ID) %>%
  arrange(desc(Cell_occupant == "parasite")) %>%
  slice(1) %>%
  ungroup()

table(nesting_cells_meg_short$Cell_occupant)
  
nesting_cells_meg_short %>% 
  group_by(Species_or_subtype) %>% 
  summarise(count =  length(unique(Nest_ID)))

nest_n_meg <- nesting_cells_meg %>%
  group_by(Nest_ID) %>%
  arrange(
    desc(Nest_parasitized == "parasitized" & Cell_occupant == 'parasite')) %>%
  slice(1) %>%
  ungroup()
nest_n_meg %>% 
  group_by(Species_or_subtype) %>% 
  summarise(count =  length(unique(Nest_ID)))
# all the NAs are megachile not developed enough to ID to species -> nest_n_meg %>% filter(is.na(Species_or_subtype))

# chalcid	                3			
# coelioxys	              6			
# meloidae	              3		
# melittobia              2
# megachile centuncularis	1	
# megachile pugnata	      18			
# megachile relativa    	16			
# megachile rotundata   	1			
# megachile             	6


## How many cells were parasitized?

nesting_cells_os %>% 
  group_by(Cell_occupant, Species_or_subtype) %>% 
  summarise(count = length(unique(Cell_ID)))
# osmia pumila - 80
# osmia tersula - 49
# osmia NA - 45
# spayga - 110
# melittobia - 3
# 287 cells total

nesting_cells_meg %>% 
  group_by(Cell_occupant, Species_or_subtype) %>% 
  summarise(count = length(unique(Cell_ID)))
# megachile centuncularis - 1
# megachile mendica - 1
# megachile pugnata - 32
# megachile relativa - 77
# megachile rotundata - 1
# megachile unknown - 10
# chalcid - 5
# coelioxys - 7
# melittobia - 3
# meloidae - 5
# 142 cells total


## How many nests per treatment?
trapnesting_os %>% 
  group_by(Treatment) %>% 
  summarize(Sum_nests = sum(New_nests_os))
# Osmia
  # low 56
  # mid 51
  # high 54


trapnesting_meg %>% 
  group_by(Treatment) %>% 
  summarize(Sum_nests = sum(New_nests_meg))
# Megachile
  # low 41
  # mid 50
  # high 46


```

```{r Simulations, include=FALSE}


## Create the function to simulate data

# simulate the null hypothesis that bees are treating each cavity equally, thus nesting would be based just on availability
# simulation without replacement so that if there are multiple new osmia nests, the number of available cavities goes down by one for each sequential new nest in that subsite-group combo

# create a function to simulate the distribution of nests based on the null hypothesis that bees are nesting according to available cavities 
simulate_nests <- function(avail_cavities, total_nests) {
  n <- length(avail_cavities) # number of available cavities
  nest_counts <- rep(0, n) # start with zero nests in each trapnest

  # keep as zeros is no nests are built
  if (total_nests == 0) {
    return(nest_counts)
  }

  # look through each of the new nests to be randomly assigned
  for (i in 1:total_nests) {
    if (sum(avail_cavities) == 0) break  # stop if all cavities used
    probability <- avail_cavities / sum(avail_cavities) # calculate probability proportional to the number of available cavities
    chosen <- sample(1:n, size = 1, prob = probability) # randomly choose a trapnest for the new nest based on available cavities
    nest_counts[chosen] <- nest_counts[chosen] + 1 # record it
    avail_cavities[chosen] <- avail_cavities[chosen] - 1 # reduce the number of available cavities by 1, when we apply the function the loop will run again until no new nests are left
  }
  return(nest_counts) # return the vector with how many nests were randomly assigned
}

# indicate how many simulated data columns we want for the model
n_sim <- 250 


## Apply the function for Osmia 

trapnesting_os_null <- trapnesting_os %>%
  group_by(Subsite, Date) %>%
  group_modify(~ { # simulate random nest distribution for subsite-date combo
    for (i in 1:n_sim){ # make each of the columns we indicated that we want 
    nest_counts <- simulate_nests(.x$Available_cavities_prior, sum(.x$New_nests_os)) # use available cavities as weights and sum the new nests
    .x[[paste0('Null_new_nests', i)]] <- nest_counts # make the new column
    }
    .x # return the modified group
  }
  ) %>%
  ungroup()


## Apply the function for Megachile

trapnesting_meg_null <- trapnesting_meg %>%
  group_by(Subsite, Date) %>%
  group_modify(~ { # simulate random nest distribution for subsite-date combo
    for (i in 1:n_sim){ # make each of the columns we indicated that we want 
    nest_counts <- simulate_nests(.x$Available_cavities_prior, sum(.x$New_nests_meg)) # use available cavities as weights and sum the new nests
    .x[[paste0('Null_new_nests', i)]] <- nest_counts # make the new column
    }
    .x # return the modified group
  }
  ) %>%
  ungroup()


```

## Congeneric Attraction - Osmia

Model 1_i - including pseudo

Model 1_ii - excluding pseudo

```{r Models_1_Congeneric_Attraction_Osmia, include=TRUE}


### MODEL 1.i - Osmia including pseudo ###

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
sim_list_os_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Osmia_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_os_ps <- c(unlist(lapply(sim_list_os_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_1i_choice_consp_os_ps <- glmer(New_nests_os ~ scale(Osmia_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_1i_choice_consp_os_ps)

# extract my observed coefficient
obvs_coef_os_ps <- summary(mod_1i_choice_consp_os_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_os_ps, c(0.025, 0.975))
obvs_coef_os_ps # falls outside of range

# convert to dataframe
df_coefs_os_ps <- data.frame(sim_coefs_os_ps = sim_coefs_os_ps)

hist_1i <- ggplot(df_coefs_os_ps, aes(x = sim_coefs_os_ps)) +
  geom_histogram(binwidth = 0.1, fill = 'grey50')  +
  geom_vline(xintercept = obvs_coef_os_ps, color = 'red3', linetype = 'dashed', size = 1) + 
  labs(y = 'Count', x = 'Coefficient')
# reject the null - Osmia are attracted to congenerics 
hist_1i


## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_os_ps < df_coefs_os_ps) # none are larger


### MODEL 1.ii - Osmia excluding pseudo ###

## Simulated data 

# run the model on each of the simulated data columns
sim_list_os_no_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_os_no_ps <- c(unlist(lapply(sim_list_os_no_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Observed data 

mod_1ii_choice_consp_os_no_ps <- glmer(New_nests_os ~ scale(Osmia_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_1ii_choice_consp_os_no_ps)

# extract my observed coefficient
obvs_coef_os_no_ps <- summary(mod_1ii_choice_consp_os_no_ps)$coefficients[2,1]

# does my observed value fall outside this range, and we reject the null?
quantile(sim_coefs_os_no_ps, c(0.025, 0.975))
obvs_coef_os_no_ps # falls outside of range

# convert to dataframe
df_coefs_os_no_ps <- data.frame(sim_coefs_os_no_ps = sim_coefs_os_no_ps)

# plot
hist_1ii <- ggplot(df_coefs_os_no_ps, aes(x = sim_coefs_os_no_ps)) +
  geom_histogram(binwidth = 0.1, fill = 'grey50')  +
  geom_vline(xintercept = obvs_coef_os_no_ps, color = 'red3', linetype = 'dashed', size = 1) +
  labs(y = 'Count', x = 'Coefficient')
# reject the null - Osmia are attracted to congenerics
hist_1ii

## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_os_no_ps < df_coefs_os_no_ps) # none are larger

# save
hist_1_os <- hist_1i + hist_1ii + plot_annotation(tag_levels = 'A')
ggsave('Figures/hist_1_i_and_ii.png', plot = hist_1_os, width = 8, height = 3.5)
saveRDS(mod_1i_choice_consp_os_ps, file = 'Models/mod_1i_choice_consp_os_ps.rds')
saveRDS(mod_1ii_choice_consp_os_no_ps, file = 'Models/mod_1ii_choice_consp_os_no_ps.rds')


# sample size
sum(trapnesting_os_null$New_nests_os)

```

## Congeneric Attraction - Megachile

Model 1_iii - including pseudo

Model 1_iv - excluding pseudo

```{r Models_1_Congeneric_Attraction_Megachile, include=TRUE}


### MODEL 1.iii - Megachile including pseudo ###

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns
sim_list_meg_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Meg_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_meg_ps <- c(unlist(lapply(sim_list_meg_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))

## Making the model part 2: Observed data 

mod_1iii_choice_consp_meg_ps <- glmer(New_nests_meg ~ scale(Meg_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
summary(mod_1iii_choice_consp_meg_ps)

# extract my observed coefficient
obvs_coef_meg_ps <- summary(mod_1iii_choice_consp_meg_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_meg_ps, c(0.025, 0.975))
obvs_coef_meg_ps # not within this range

# convert to dataframe
df_coefs_meg_ps <- data.frame(sim_coefs_meg_ps = sim_coefs_meg_ps)

hist_1iii <- ggplot(df_coefs_meg_ps, aes(x = sim_coefs_meg_ps)) +
  geom_histogram(binwidth = 0.1, fill = 'grey50')  +
  geom_vline(xintercept = obvs_coef_meg_ps, color = 'red3', linetype = 'dashed', size = 1) +
  labs(y = 'Count', x = 'Coefficient')
# reject the null - Megachile are attracted to congenerics 
hist_1iii


## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_meg_ps < df_coefs_meg_ps) # none are larger


### MODEL 1.iv - Megachile excluding pseudo ###

## Making the model part 1: Simulated data 

# run the model on each of the simulated data columns
sim_list_meg_no_ps_attr <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(Meg_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson)
  summary(output)
} ) 

# take all of the simulated coefficients
sim_coefs_meg_no_ps <- c(unlist(lapply(sim_list_meg_no_ps_attr, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_1iv_choice_consp_meg_no_ps <- glmer(New_nests_meg ~ scale(Meg_no_ps_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
summary(mod_1iv_choice_consp_meg_no_ps)

# extract my observed coefficient
obvs_coef_meg_no_ps <- summary(mod_1iv_choice_consp_meg_no_ps)$coefficients[2,1]

# range of 95%, i.e., if my observed value falls within this range, its not different from the null
quantile(sim_coefs_meg_no_ps, c(0.025, 0.975))
obvs_coef_meg_no_ps # falls outside this range

# convert to dataframe
df_coefs_meg_no_ps <- data.frame(sim_coefs_meg_no_ps = sim_coefs_meg_no_ps)

hist_1iv <- ggplot(df_coefs_meg_no_ps, aes(x = sim_coefs_meg_no_ps)) +
  geom_histogram(binwidth = 0.1, fill = 'grey50')  +
  geom_vline(xintercept = obvs_coef_meg_no_ps, color = 'red3', linetype = 'dashed', size = 1) +
  labs(y = 'Count', x = 'Coefficient')
# reject the null - Megachile are attracted to congenerics 
hist_1iv


## Reporting

# what proportion of the simulated coefs fall above mine?
table(obvs_coef_meg_no_ps < df_coefs_meg_no_ps) # none are larger

# save
hist_1_meg <- hist_1iii + hist_1iv + plot_annotation(tag_levels = 'A')
ggsave('Figures/hist_1_iii_and_iv.png', plot = hist_1_meg, width = 8, height = 3.5)
saveRDS(mod_1iii_choice_consp_meg_ps, file = 'Models/mod_1iii_choice_consp_meg_ps.rds')
saveRDS(mod_1iv_choice_consp_meg_no_ps, file = 'Models/mod_1iv_choice_consp_meg_no_ps.rds')

# sample size
sum(trapnesting_meg_null$New_nests_meg)

```


## Parasite Preferences - Osmia (Sapyga)

Model 2_i - including pseudo

Model 2_ii - excluding pseudo

Model 2_iii - active nests

```{r Models_2_Parasites_Attraction_Osmia, include=TRUE, fig.width=10, fig.height=10}


### MODEL 2_i - Osmia parasite aggregation preferences, including pseudo ###

mod_2i_par_agg_os_ps <- glmer(Cell_parasitized ~ scale(Osmia_ps) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_2i_par_agg_os_ps)) # good
check_model(mod_2i_par_agg_os_ps) # some influential points but ultimately fine
testDispersion(mod_2i_par_agg_os_ps, type = 'DHARMa') # good
testZeroInflation(mod_2i_par_agg_os_ps) # good
vif(mod_2i_par_agg_os_ps) # good
summary(mod_2i_par_agg_os_ps) # effect of day of year

# plot
values_osmia_ps <- seq(min(nesting_cells_os$Osmia_ps, na.rm = TRUE), 
               max(nesting_cells_os$Osmia_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_os_ps <- ggpredict(mod_2i_par_agg_os_ps, 
               terms = 'Osmia_ps [all]', 
               conditions = list(Osmia_ps = values_osmia_ps))
plot(marginal_effects_os_ps)  # trend for MORE parasitism with increasing osmia numbers



### MODEL 2_ii - Osmia parasite aggregation preferences, excluding pseudo ###

mod_2ii_par_agg_os_no_ps <- glmer(Cell_parasitized ~ scale(Osmia_no_ps) + poly(Day_of_year,2) + Site + (1|Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_2ii_par_agg_os_no_ps)) # good
check_model(mod_2ii_par_agg_os_no_ps) # some influential points but fine
testDispersion(mod_2ii_par_agg_os_no_ps, type = 'DHARMa') # good
testZeroInflation(mod_2ii_par_agg_os_no_ps) # good
vif(mod_2ii_par_agg_os_no_ps)
summary(mod_2ii_par_agg_os_no_ps) # all effects significant 


# plot
values_osmia_no_ps <- seq(min(nesting_cells_os$Osmia_no_ps, na.rm = TRUE), 
               max(nesting_cells_os$Osmia_no_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_os_no_ps <- ggpredict(mod_2ii_par_agg_os_no_ps, 
               terms = 'Osmia_no_ps [all]', 
               conditions = list(Osmia_no_ps = values_osmia_no_ps))
plot(marginal_effects_os_no_ps) # trend with less parasitism for more osmia 



### MODEL 2_iii - Osmia parasite aggregation preferences for active nests ###

mod_2iii_par_agg_os_active <- glmer(Cell_parasitized ~ scale(Osmia_active_mothers) + poly(Day_of_year,2) + Site + (1|Nest_ID), data = nesting_cells_os, family = binomial)

plot(simulateResiduals(mod_2iii_par_agg_os_active)) # good
check_model(mod_2iii_par_agg_os_active) # good
testDispersion(mod_2iii_par_agg_os_active, type = 'DHARMa') # good
testZeroInflation(mod_2iii_par_agg_os_active) # good
vif(mod_2iii_par_agg_os_active) # good
summary(mod_2iii_par_agg_os_active) # effect of osmia mothers and day of year 

# plot
values_osmia_active <- seq(min(nesting_cells_os$Osmia_active_mothers, na.rm = TRUE), 
               max(nesting_cells_os$Osmia_active_mothers, na.rm = TRUE), 
               length.out = 100)
marginal_effects_os_actives <- ggpredict(mod_2iii_par_agg_os_active, 
               terms = 'Osmia_active_mothers [all]', 
               conditions = list(Osmia_active_mothers = values_osmia_active))
plot(marginal_effects_os_actives) # trend with less parasitism for more active osmia mothers


# save models
saveRDS(mod_2i_par_agg_os_ps, file = 'Models/mod_2i_par_agg_os_ps.rds')
saveRDS(mod_2ii_par_agg_os_no_ps, file = 'Models/mod_2ii_par_agg_os_no_ps.rds')
saveRDS(mod_2iii_par_agg_os_active, file = 'Models/mod_2iii_par_agg_os_active.rds')

# sample size
table(nesting_cells_os$Cell_parasitized) # 174 + 113 = 287


```

## Parasite Preferences - Megachile (Meloid, Coelioxys, Chalcid)

Model 2_iv - including pseudo

Model 2_v - excluding pseudo

Model 2_vi - active nests

```{r Models_2_Parasites_Attraction_Megachile, include=TRUE, fig.width=10, fig.height=10}


### MODEL 2_iv - Megachile parasite aggregation preferences, including pseudo ###

mod_2iv_par_agg_meg_ps <- glmer(Cell_parasitized ~ scale(Meg_ps) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_meg, family = binomial)

plot(simulateResiduals(mod_2iv_par_agg_meg_ps)) # fine
check_model(mod_2iv_par_agg_meg_ps) # fine
testDispersion(mod_2iv_par_agg_meg_ps, type = 'DHARMa') # good
testZeroInflation(mod_2iv_par_agg_meg_ps) # good
vif(mod_2iv_par_agg_meg_ps) # good
summary(mod_2iv_par_agg_meg_ps) # no detected effects


# plot
values_meg_ps <- seq(min(nesting_cells_meg$Meg_ps, na.rm = TRUE), 
               max(nesting_cells_meg$Meg_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_meg_ps <- ggpredict(mod_2iv_par_agg_meg_ps, 
               terms = 'Meg_ps [all]', 
               conditions = list(Meg_ps = values_meg_ps))
plot(marginal_effects_meg_ps) # no trend


# MODEL 2_v - Megachile parasite aggregation preferences, excluding pseudo #

mod_2v_par_agg_meg_no_ps <- glmer(Cell_parasitized ~ scale(Meg_no_ps) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_meg, family = binomial)

plot(simulateResiduals(mod_2v_par_agg_meg_no_ps)) # good
check_model(mod_2v_par_agg_meg_no_ps) # fine
testDispersion(mod_2v_par_agg_meg_no_ps, type = 'DHARMa') # good
testZeroInflation(mod_2v_par_agg_meg_no_ps) # good
vif(mod_2v_par_agg_meg_no_ps) # good
summary(mod_2v_par_agg_meg_no_ps) # no detected effects


# plot
values_meg_no_ps <- seq(min(nesting_cells_meg$Meg_no_ps, na.rm = TRUE), 
               max(nesting_cells_meg$Meg_no_ps, na.rm = TRUE), 
               length.out = 100)
marginal_effects_meg_no_ps <- ggpredict(mod_2v_par_agg_meg_no_ps, 
               terms = 'Meg_no_ps [all]', 
               conditions = list(Meg_no_ps = values_meg_no_ps))
plot(marginal_effects_meg_no_ps) # no trend 



### MODEL 2_vi - Megachile parasite aggregation preferences for active nests ###

mod_2vi_par_agg_meg_active <- glmer(Cell_parasitized ~ scale(Meg_active_mothers) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest/Nest_ID), data = nesting_cells_meg, family = binomial)

plot(simulateResiduals(mod_2vi_par_agg_meg_active)) # fine
check_model(mod_2vi_par_agg_meg_active) # fine
testDispersion(mod_2vi_par_agg_meg_active, type = 'DHARMa') # good
testZeroInflation(mod_2vi_par_agg_meg_active) # good 
vif(mod_2vi_par_agg_meg_active) # good
summary(mod_2vi_par_agg_meg_active) # no detected effects, close for active mothers


# plot
values_meg_active <- seq(min(nesting_cells_meg$Meg_active_mothers, na.rm = TRUE), 
               max(nesting_cells_meg$Meg_active_mothers, na.rm = TRUE), 
               length.out = 100)
marginal_effects_meg_active <- ggpredict(mod_2vi_par_agg_meg_active, 
               terms = 'Meg_active_mothers [all]', 
               conditions = list(Meg_active_mothers = values_meg_active))
plot(marginal_effects_meg_active) # trend for more parasites with less active megachile


# save models
saveRDS(mod_2iv_par_agg_meg_ps, file = 'Models/mod_2iv_par_agg_meg_ps.rds')
saveRDS(mod_2v_par_agg_meg_no_ps, file = 'Models/mod_2v_par_agg_meg_no_ps.rds')
saveRDS(mod_2vi_par_agg_meg_active, file = 'Models/mod_2vi_par_agg_meg_active.rds')

# sample size
table(nesting_cells_meg$Cell_parasitized) # 122 + 20 = 142

```

## Parasite Avoidance - Osmia

Model 3a_i - including pseudo

Model 3a_ii - excluding pseudo

```{r Models_3a_Parasite_Avoidance_Osmia, include=TRUE}


### MODEL 3a_i - Osmia avoidance of parasites, including pseudo ###

# run the model on each of the simulated data columns
sim_list_3ai_os_ps_avoid <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(DP_os_prior) * scale(NDP_ps_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson())
  summary(output)
} )

# take all of the simulated coefficients for my three variable sof interest (DP, NDP, DP*NDP)
sim_coefs_3ai_DP <- c(unlist(lapply(sim_list_3ai_os_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate for DP
})))
sim_coefs_3ai_NDP <- c(unlist(lapply(sim_list_3ai_os_ps_avoid, function(x) {
  x$coefficients[3,1] # extract third row, first column for the estimate for NDP
})))
sim_coefs_3ai_DPxNDP <- c(unlist(lapply(sim_list_3ai_os_ps_avoid, function(x) {
  x$coefficients[8,1] # extract eigth row, first column for the estimate for interaction
})))


## Making the model part 2: Observed data 

mod_3ai_choice_consp_par_os_ps <- glmer(New_nests_os ~ scale(DP_os_prior) * scale(NDP_ps_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_3ai_choice_consp_par_os_ps)

# plot
plot(ggpredict(mod_3ai_choice_consp_par_os_ps, terms = c('DP_os_prior', 'NDP_ps_os_prior'))) 


## DP_os_prior expected vs observed

  # extract my observed coefficient
  obvs_coef_3ai_DP <- summary(mod_3ai_choice_consp_par_os_ps)$coefficients[2,1]
  
  # range of 95%, i.e., if my observed value falls within this range, its not different from the null
  quantile(sim_coefs_3ai_DP, c(0.025, 0.975))
  obvs_coef_3ai_DP # close but falls within this range
  
  # convert to dataframe
  df_coefs_3ai_DP <- data.frame(sim_coefs_3ai_DP = sim_coefs_3ai_DP)
  
  hist_3ai_DP <- ggplot(df_coefs_3ai_DP, aes(x = sim_coefs_3ai_DP)) +
    geom_histogram(binwidth = 0.1, fill = 'grey50')  +
    geom_vline(xintercept = obvs_coef_3ai_DP, color = 'red3', linetype = 'dashed', size = 1) +
    labs(y = 'Count', x = 'Coefficient') + 
    theme(plot.margin = margin(1, 1, 1, 1))
  # fail to reject the null - no evidence for Osmia avoiding nests with parasites
  hist_3ai_DP
  
  # what proportion of the simulated coefs fall above mine?
  table(obvs_coef_3ai_DP < df_coefs_3ai_DP) # 31 are larger (varies everytime we rerun simulations)
  
  
## NDP_os_prior expected vs observed
  
  # extract my observed coefficient
  obvs_coef_3ai_NDP <- summary(mod_3ai_choice_consp_par_os_ps)$coefficients[3,1]
  
  # range of 95%, i.e., if my observed value falls within this range, its not different from the null
  quantile(sim_coefs_3ai_NDP, c(0.025, 0.975))
  obvs_coef_3ai_NDP # significant, falls outside this range
  
  # convert to dataframe
  df_coefs_3ai_NDP <- data.frame(sim_coefs_3ai_NDP = sim_coefs_3ai_NDP)
  
  hist_3ai_NDP <- ggplot(df_coefs_3ai_NDP, aes(x = sim_coefs_3ai_NDP)) +
    geom_histogram(binwidth = 0.1, fill = 'grey50')  +
    geom_vline(xintercept = obvs_coef_3ai_NDP, color = 'red3', linetype = 'dashed', size = 1) +
    labs(y = 'Count', x = 'Coefficient') + 
    theme(plot.margin = margin(1, 1, 1, 1))
  #reject the null - Osmia attracted to nests without parasites
  hist_3ai_NDP
  
  # what proportion of the simulated coefs fall above mine?
  table(obvs_coef_3ai_NDP < df_coefs_3ai_NDP) # 0 are larger

  
## DP_os_prior * NDP_os_prior expected vs observed
  
  # extract my observed coefficient
  obvs_coef_3ai_DPxNDP <- summary(mod_3ai_choice_consp_par_os_ps)$coefficients[8,1]
  
  # range of 95%, i.e., if my observed value falls within this range, its not different from the null
  quantile(sim_coefs_3ai_DPxNDP, c(0.025, 0.975))
  obvs_coef_3ai_DPxNDP # falls within this range
  
  # convert to dataframe
  df_coefs_3ai_DPxNDP <- data.frame(sim_coefs_3ai_DPxNDP = sim_coefs_3ai_DPxNDP)
  
  hist_3ai_DPxNDP <- ggplot(df_coefs_3ai_DPxNDP, aes(x = sim_coefs_3ai_DPxNDP)) +
    geom_histogram(binwidth = 0.1, fill = 'grey50')  +
    geom_vline(xintercept = obvs_coef_3ai_DPxNDP, color = 'red3', linetype = 'dashed', size = 1) +
    labs(y = 'Count', x = 'Coefficient') + 
    theme(plot.margin = margin(1, 1, 1, 1))
  # fail to reject the null - no evidence for Osmia avoiding nests with parasites
  hist_3ai_DPxNDP
  
  # what proportion of the simulated coefs fall above mine?
  table(obvs_coef_3ai_DPxNDP < df_coefs_3ai_DPxNDP) # 214 are larger
 
# save all three
hist_3ai <- hist_3ai_DP + hist_3ai_NDP + hist_3ai_DPxNDP + plot_annotation(tag_levels = 'A')
ggsave('Figures/hist_3ai.png', plot = hist_3ai, width = 8, height = 3)
  
  
  
### MODEL 3a_ii - Osmia avoidance of parasites, excluding pseudo ###

# run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
sim_list_3aii_os_no_ps_avoid <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(DP_os_prior) * scale(NDP_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_os_null, family = poisson())
  summary(output)
} )

# take all of the simulated coefficients
sim_coefs_3aii_DP <- c(unlist(lapply(sim_list_3aii_os_no_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))
sim_coefs_3aii_NDP <- c(unlist(lapply(sim_list_3aii_os_no_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))
sim_coefs_3aii_DPxNDP <- c(unlist(lapply(sim_list_3aii_os_no_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate 
})))


## Making the model part 2: Observed data 

mod_3aii_choice_consp_par_os_no_ps <- glmer(New_nests_os ~ scale(DP_os_prior) * scale(NDP_os_prior) + scale(Good_days) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = trapnesting_os_null, family = poisson)
summary(mod_3aii_choice_consp_par_os_no_ps)

# plot
plot(ggpredict(mod_3aii_choice_consp_par_os_no_ps, terms = c('DP_os_prior', 'NDP_os_prior'))) 


## DP_os_prior expected vs observed

  # extract my observed coefficient
  obvs_coef_3aii_DP <- summary(mod_3aii_choice_consp_par_os_no_ps)$coefficients[2,1]
  
  # range of 95%, i.e., if my observed value falls within this range, its not different from the null
  quantile(sim_coefs_3aii_DP, c(0.025, 0.975))
  obvs_coef_3aii_DP # outside this range, significant 
  
  # convert to dataframe
  df_coefs_3aii_DP <- data.frame(sim_coefs_3aii_DP = sim_coefs_3aii_DP)
  
  hist_3aii_DP <- ggplot(df_coefs_3aii_DP, aes(x = sim_coefs_3aii_DP)) +
    geom_histogram(binwidth = 0.1, fill = 'grey50')  +
    geom_vline(xintercept = obvs_coef_3aii_DP, color = 'red3', linetype = 'dashed', linewidth = 1) +
    labs(y = 'Count', x = 'Coefficient') + 
    theme(plot.margin = margin(1, 1, 1, 1))
  # not too far out 
  hist_3aii_DP
  
  # what proportion of the simulated coefs fall above mine?
  table(obvs_coef_3aii_DP < df_coefs_3aii_DP) # 3 are larger (varies with each new set of simulations)

  
## NDP_os_prior expected vs observed
  
  # extract my observed coefficient
  obvs_coef_3aii_NDP <- summary(mod_3aii_choice_consp_par_os_no_ps)$coefficients[3,1]
  
  # range of 95%, i.e., if my observed value falls within this range, its not different from the null
  quantile(sim_coefs_3aii_NDP, c(0.025, 0.975))
  obvs_coef_3aii_NDP # significant, falls outside this range
  
  # convert to dataframe
  df_coefs_3aii_NDP <- data.frame(sim_coefs_3aii_NDP = sim_coefs_3aii_NDP)
  
  hist_3aii_NDP <- ggplot(df_coefs_3aii_NDP, aes(x = sim_coefs_3aii_NDP)) +
    geom_histogram(binwidth = 0.1, fill = 'grey50')  +
    geom_vline(xintercept = obvs_coef_3aii_NDP, color = 'red3', linetype = 'dashed', size = 1) +
    labs(y = 'Count', x = 'Coefficient') + 
    theme(plot.margin = margin(1, 1, 1, 1))
  # reject the null - Osmia attracted to nests without parasites
  hist_3aii_NDP
  
  # what proportion of the simulated coefs fall above mine?
  table(obvs_coef_3aii_NDP < df_coefs_3aii_NDP) # 0 are larger

  
## DP_os_prior * NDP_os_prior expected vs observed
  
  # extract my observed coefficient
  obvs_coef_3aii_DPxNDP <- summary(mod_3aii_choice_consp_par_os_no_ps)$coefficients[8,1]
  
  # range of 95%, i.e., if my observed value falls within this range, its not different from the null
  quantile(sim_coefs_3aii_DPxNDP, c(0.025, 0.975))
  obvs_coef_3aii_DPxNDP # falls within this range
  
  # convert to dataframe
  df_coefs_3aii_DPxNDP <- data.frame(sim_coefs_3aii_DPxNDP = sim_coefs_3aii_DPxNDP)
  
  hist_3aii_DPxNDP <- ggplot(df_coefs_3aii_DPxNDP, aes(x = sim_coefs_3aii_DPxNDP)) +
    geom_histogram(binwidth = 0.1, fill = 'grey50')  +
    geom_vline(xintercept = obvs_coef_3aii_DPxNDP, color = 'red3', linetype = 'dashed', size = 1) +
    labs(y = 'Count', x = 'Coefficient') + 
    theme(plot.margin = margin(1, 1, 1, 1))
  # fail to reject the null - no evidence for Osmia avoiding nests with parasites
  hist_3aii_DPxNDP
  
  # what proportion of the simulated coefs fall above mine?
  table(obvs_coef_3aii_DPxNDP < df_coefs_3aii_DPxNDP) # 224 are larger 

# save all three
hist_3aii <- hist_3aii_DP + hist_3aii_NDP + hist_3aii_DPxNDP + plot_annotation(tag_levels = 'A')
ggsave('Figures/hist_3aii.png', plot = hist_3aii, width = 8, height = 3)
saveRDS(mod_3ai_choice_consp_par_os_ps, file = 'Models/mod_3ai_choice_consp_par_os_ps.rds')
saveRDS(mod_3aii_choice_consp_par_os_no_ps, file = 'Models/mod_3aii_choice_consp_par_os_no_ps.rds')

# sample size
sum(trapnesting_os_null$New_nests_os) # 161


```

## Parasite Avoidance - Megachile

Model 3a_iii - including pseudo

Model 3a_iv - excluding pseudo

```{r Models_3a_Parasite_Avoidance_Megachile, include=TRUE}


### MODEL 3a_iii - Megachile avoidance of parasites, including pseudo ###

# run the model on each of the simulated data columns
sim_list_3aiii_meg_ps_avoid <- lapply(1:n_sim, function(x){
  my_formula <- paste0('Null_new_nests', x, ' ~ scale(DP_meg_prior) * scale(NDP_ps_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)') 
  output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson())
  summary(output)
} )

# take all of the simulated coefficients for my three variable sof interest (DP, NDP, DP*NDP)
sim_coefs_3aiii_DP <- c(unlist(lapply(sim_list_3aiii_meg_ps_avoid, function(x) {
  x$coefficients[2,1] # extract second row, first column for the estimate for DP
})))
sim_coefs_3aiii_NDP <- c(unlist(lapply(sim_list_3aiii_meg_ps_avoid, function(x) {
  x$coefficients[3,1] # extract third row, first column for the estimate for NDP
})))
sim_coefs_3aiii_DPxNDP <- c(unlist(lapply(sim_list_3aiii_meg_ps_avoid, function(x) {
  x$coefficients[9,1] # extract eigth row, first column for the estimate for interaction
})))


## Making the model part 2: Observed data 

mod_3aiii_choice_consp_par_meg_ps <- glmer(New_nests_meg ~ scale(DP_meg_prior) * scale(NDP_ps_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
summary(mod_3aiii_choice_consp_par_meg_ps)

# plot 
plot(ggpredict(mod_3aiii_choice_consp_par_meg_ps, terms = c('NDP_ps_meg_prior [all]', 'DP_meg_prior'))) 

# save for using in figures doc
saveRDS(mod_3aiii_choice_consp_par_meg_ps, file = 'Models/mod_3aiii_choice_consp_par_meg_ps.rds')

## DP_meg_prior expected vs observed

  # extract my observed coefficient
  obvs_coef_3aiii_DP <- summary(mod_3aiii_choice_consp_par_meg_ps)$coefficients[2,1]
  
  # range of 95%, i.e., if my observed value falls within this range, its not different from the null
  quantile(sim_coefs_3aiii_DP, c(0.025, 0.975))
  obvs_coef_3aiii_DP # falls within this range
  
  # convert to dataframe
  df_coefs_3aiii_DP <- data.frame(sim_coefs_3aiii_DP = sim_coefs_3aiii_DP)
  
  hist_3aiii_DP <- ggplot(df_coefs_3aiii_DP, aes(x = sim_coefs_3aiii_DP)) +
    geom_histogram(binwidth = 0.1, fill = 'grey50')  +
    geom_vline(xintercept = obvs_coef_3aiii_DP, color = 'red3', linetype = 'dashed', size = 1) +
    labs(y = 'Count', x = 'Coefficient') + 
    theme(plot.margin = margin(1, 1, 1, 1))
  # fail to reject the null - no evidence for Megachile avoiding nests with parasites
  hist_3aiii_DP
  
  # what proportion of the simulated coefs fall above mine?
  table(obvs_coef_3aiii_DP < df_coefs_3aiii_DP) # 96 are larger (varies with simulations)

  
## NDP_meg_prior expected vs observed
  
  # extract my observed coefficient
  obvs_coef_3aiii_NDP <- summary(mod_3aiii_choice_consp_par_meg_ps)$coefficients[3,1]
  
  # range of 95%, i.e., if my observed value falls within this range, its not different from the null
  quantile(sim_coefs_3aiii_NDP, c(0.025, 0.975))
  obvs_coef_3aiii_NDP # significant, falls outside this range
  
  # convert to dataframe
  df_coefs_3aiii_NDP <- data.frame(sim_coefs_3aiii_NDP = sim_coefs_3aiii_NDP)
  
  hist_3aiii_NDP <- ggplot(df_coefs_3aiii_NDP, aes(x = sim_coefs_3aiii_NDP)) +
    geom_histogram(binwidth = 0.1, fill = 'grey50')  +
    geom_vline(xintercept = obvs_coef_3aiii_NDP, color = 'red3', linetype = 'dashed', size = 1) +
    labs(y = 'Count', x = 'Coefficient') + 
    theme(plot.margin = margin(1, 1, 1, 1))
  # reject the null - Megachile attracted to nests without parasites
  hist_3aiii_NDP
  
  # what proportion of the simulated coefs fall above mine?
  table(obvs_coef_3aiii_NDP < df_coefs_3aiii_NDP) # 1 is larger

  
## DP_meg_prior * NDP_meg_prior expected vs observed
  
  # extract my observed coefficient
  obvs_coef_3aiii_DPxNDP <- summary(mod_3aiii_choice_consp_par_meg_ps)$coefficients[9,1]
  
  # range of 95%, i.e., if my observed value falls within this range, its not different from the null
  quantile(sim_coefs_3aiii_DPxNDP, c(0.025, 0.975))
  obvs_coef_3aiii_DPxNDP # falls outsdie this range, smaller
  
  # convert to dataframe
  df_coefs_3aiii_DPxNDP <- data.frame(sim_coefs_3aiii_DPxNDP = sim_coefs_3aiii_DPxNDP)
   
  hist_3aiii_DPxNDP <- ggplot(df_coefs_3aiii_DPxNDP, aes(x = sim_coefs_3aiii_DPxNDP)) +
    geom_histogram(binwidth = 0.1, fill = 'grey50') +
    geom_vline(xintercept = obvs_coef_3aiii_DPxNDP, color = 'red3', linetype = 'dashed', size = 1) +
    labs(y = 'Count', x = 'Coefficient') + 
    theme(plot.margin = margin(1, 1, 1, 1))
  # reject the null - Megachile avoid nests with parasites
  hist_3aiii_DPxNDP
  
  # what proportion of the simulated coefs fall above mine?
  table(obvs_coef_3aiii_DPxNDP < df_coefs_3aiii_DPxNDP) # 125 are larger
 
# save all three
hist_3aiii <- hist_3aiii_DP + hist_3aiii_NDP + hist_3aiii_DPxNDP + plot_annotation(tag_levels = 'A')
ggsave('Figures/hist_3aiii.png', plot = hist_3aiii, width = 8, height = 3)
saveRDS(mod_3aiii_choice_consp_par_meg_ps, file = 'Models/mod_3aiii_choice_consp_par_meg_ps.rds')



### MODEL 3a_iv - Megachile avoidance of parasites, excluding pseudo ###

# run only observed data model
# running the simulation gets the error 'Error: PIRLS loop resulted in NaN value' or 'Error in pwrssUpdate(pp, resp, tol = tolPwrss, GQmat = GQmat, compDev = compDev,  : Downdated VtV is not positive definite'
# too little data to run it

# run the model on each of the simulated data columns, which will create the summary outputs of each simulated data model
# sim_list_meg_no_ps_avoid <- lapply(1:n_sim, function(x){
# my_formula <- paste0('Null_new_nests', x, ' ~ scale(DP_meg_prior) * scale(NDP_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest)')
# output <- glmer(as.formula(my_formula), data = trapnesting_meg_null, family = poisson())
# summary(output)
# } )


## Making observed data model

mod_3aiv_choice_consp_par_meg_no_ps <- glmer(New_nests_meg ~ scale(DP_meg_prior) * scale(NDP_meg_prior) + scale(Good_days) + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = trapnesting_meg_null, family = poisson)
summary(mod_3aiv_choice_consp_par_meg_no_ps)

# save
saveRDS(mod_3aiv_choice_consp_par_meg_no_ps, file = 'Models/mod_3aiv_choice_consp_par_meg_no_ps.rds')


```

## Parasite Defences

Model 3b_i - length of straw filled, Osmia

Model 3b_ii - length of straw filled, Megachile

Model 3c_i - nest termination, Osmia

Model 3c_ii - nest termination, Megachile

Model 3d_i - abandoned inner wall, Osmia

Model 3d_ii - abandoned inner wall, Megachile

```{r Models_3b_Nest_Investment, include=TRUE, fig.width=10, fig.height=10}

# MODEL 3b_i - Osmia investment #
## try beta-binomial distribution with glmTMB

# fit as guassion, total-straw length as covariate in the model as fixed 


# Investment into a nest if other active nests or the nest itself is parasitized
library(lmerTest)

mod_3bi_invest_filled_os <- lmer(Filled_length ~ scale(DP_os_prior) * Nest_parasitized + Straw_length + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_investment_os)

check_model(mod_3bi_invest_filled_os) # good
vif(mod_3bi_invest_filled_os) # good
summary(mod_3bi_invest_filled_os) # close to significant for if focal nest is parasitized, and day of year


plot(ggpredict(mod_3bi_invest_filled_os, terms = c('DP_os_prior', 'Nest_parasitized'))) 


# MODEL 3b_ii - Megachile investment #

# Investment into a nest if other active nests or the nest itself is parasitized

mod_3bii_invest_filled_meg <- lmer(Filled_length ~ scale(DP_meg_prior) * Nest_parasitized + Straw_length + poly(Day_of_year,2) + Year + Site + (1|Subsite/Trapnest), data = nesting_investment_meg)

check_model(mod_3bii_invest_filled_meg) # good
vif(mod_3bii_invest_filled_meg) # good
summary(mod_3bii_invest_filled_meg) # significant effect of day of year and site


plot(ggpredict(mod_3bii_invest_filled_meg, terms = c('DP_meg_prior', 'Nest_parasitized'))) 


# save models
saveRDS(mod_3bi_invest_filled_os, file = 'Models/mod_3bi_invest_filled_os.rds')
saveRDS(mod_3bii_invest_filled_meg, file = 'Models/mod_3bii_invest_filled_meg.rds')

# sample size
nesting_investment_os %>%
  filter(!is.na(Nest_parasitized)) %>%
  summarise(n_unique_nests = n_distinct(Nest_ID)) # 74
nesting_investment_meg %>%
  filter(!is.na(Nest_parasitized)) %>%
  summarise(n_unique_nests = n_distinct(Nest_ID)) # 40

```

```{r Models_3c_Nest_Termination, include=TRUE, fig.width=10, fig.height=10}


# Remove NAs from unknown cell positions
nesting_cells_os_noNA <- nesting_cells_os %>% filter(!is.na(Cell_position_bi))
nesting_cells_meg_noNA <- nesting_cells_meg %>% filter(!is.na(Cell_position_bi))


# MODEL 3c_i - Osmia nest termination #

mod_3ci_last_cell_os <- glmer(Cell_position_bi ~ Cell_parasitized + poly(Day_of_year,2) + Site + (1|Nest_ID), data = nesting_cells_os_noNA, family = binomial)

plot(simulateResiduals(mod_3ci_last_cell_os)) # good
check_model(mod_3ci_last_cell_os) # fine
vif(mod_3ci_last_cell_os) # good
summary(mod_3ci_last_cell_os) # no significant effects

ggpredict(mod_3ci_last_cell_os, terms = 'Cell_parasitized [all]') %>%
  plot(jitter = .04)


# MODEL 3c_ii - Megachile nest termination #

mod_3cii_last_cell_meg <- glmer(Cell_position_bi ~ Cell_parasitized + poly(Day_of_year,2) + Year + Site + (1|Nest_ID), data = nesting_cells_meg_noNA, family = binomial)

plot(simulateResiduals(mod_3cii_last_cell_meg)) # good
check_model(mod_3cii_last_cell_meg) # fine
summary(mod_3cii_last_cell_meg) # significant effect of day of year
vif(mod_3cii_last_cell_meg) # good

ggpredict(mod_3cii_last_cell_meg, terms = 'Cell_parasitized [all]') %>%
  plot(jitter = .04)
# slight trend for more likely to best last cell if its parasitized


# save models
saveRDS(mod_3ci_last_cell_os, file = 'Models/mod_3ci_last_cell_os.rds')
saveRDS(mod_3cii_last_cell_meg, file = 'Models/mod_3cii_last_cell_meg.rds')

# sample size
length(unique(nesting_cells_os_noNA$Nest_ID)) # 75 nests
length(unique(nesting_cells_meg_noNA$Nest_ID)) # 52 nests

# sample size including unknown cell positions
length(unique(nesting_cells_os$Nest_ID)) # 81 nests
length(unique(nesting_cells_meg$Nest_ID)) # 56 nests

```

```{r Models_3d_Nest_Abandonment, include=TRUE, fig.width=10, fig.height=10}

# MODEL 3d_i - Osmia abandoned inner walls including pseudo #

mod_3di_inner_wall_os <- glmer(Abandoned_inner_wall ~ scale(DP_os_prior) * scale(NDP_ps_os_prior) + poly(Day_of_year,2) + Site + (1|Subsite/Trapnest), data = nesting_os, family = binomial)

plot(simulateResiduals(mod_3di_inner_wall_os)) # good
check_model(mod_3di_inner_wall_os) # good
summary(mod_3di_inner_wall_os) # significant effect of day of year
vif(mod_3di_inner_wall_os) # good

ggpredict(mod_3di_inner_wall_os, terms = 'DP_os_prior [all]') %>%
  plot(jitter = .04)


# MODEL 3d_ii - Megachile abandoned inner walls #

# no parasites when walls are abandoned
nesting_meg %>% filter(Abandoned_inner_wall == 1) 


# save model
saveRDS(mod_3di_inner_wall_os, file = 'Models/mod_3di_inner_wall_os.rds')
# saveRDS(mod_3dii_inner_wall_meg, file = 'Models/mod_3dii_inner_wall_meg.rds')

# sample size
nesting_os %>% filter(!is.na(Abandoned_inner_wall)) # 168
nesting_os %>%
  summarise(Count = sum(Abandoned_inner_wall, na.rm = TRUE)) # 28
nesting_meg %>% filter(!is.na(Abandoned_inner_wall)) # 145
nesting_meg %>%
  summarise(Count = sum(Abandoned_inner_wall, na.rm = TRUE)) # 13


```
```{r AIC_Comparing_Models}

## COmapring AIC between models including and excluding pseudo nests to examine how pseudo nests were perceived 


# Prediction 1: Congeneric attraction

  # Osmia - difference of 6.4
AIC(mod_1i_choice_consp_os_ps) - AIC(mod_1ii_choice_consp_os_no_ps)

  # Megachile - difference of 14.4
AIC(mod_1iii_choice_consp_meg_ps) - AIC(mod_1iv_choice_consp_meg_no_ps)


# Prediction 2: Risk of nesting in aggregations

  # Osmia - difference of 2.7
AIC(mod_2i_par_agg_os_ps) - AIC(mod_2ii_par_agg_os_no_ps)

  # Megachile - difference of 1.5
AIC(mod_2iv_par_agg_meg_ps) - AIC(mod_2v_par_agg_meg_no_ps)


# Prediction 3: Avoidance of parasites

  # Osmia - difference of 14.1
AIC(mod_3ai_choice_consp_par_os_ps) - AIC(mod_3aii_choice_consp_par_os_no_ps)

  # Megachile - difference of 44.6
AIC(mod_3aiii_choice_consp_par_meg_ps) - AIC(mod_3aiv_choice_consp_par_meg_no_ps)



```


