---
title: "Nest-Site Selection in Non-Eusocial Cavity-Nesting Bees"
format: 
  html:
    self-contained: true
    toc: true
  docx: default
editor: visual
bibliography: references.bib
---

## Background

Strategies for raising offspring exist on a spectrum from cooperative breeding in tightly knit social groups to raising young in isolation who may not even encounter their parents or siblings after emerging. Raising young in proximity with others has benefits such as increased protection from predators and enhanced knowledge of local resources. However, proximity can increase the circulation of pathogens and foster competition if resources become scarce. Bees have adopted the entirety of this spectrum, as species range from solitary (e.g., most leafcutter bees: *Megachile* spp.) to advanced eusociality (e.g., honeybees: *Apis* spp.). Some species along this spectrum are facultatively social, meaning that within a population, they exhibit social polymorphism for their nesting strategy and thus can be solitary or social (e.g., some carpenter bees: *Xylocopa* spp.)

Bees will nest in the ground or in above-ground cavities, depending on the species. Nesting in proximity to others is common among ground-nesting bees. There are a few outlined explanations for why ground-nesting bee aggregations may occur. Bees of the same species are likely to prefer the same site characteristics; thus, one location may be appealing to several individuals. Additionally, resource limitation can encourage aggregations, as bees nest near floral resources, and nesting locations themselves are resources. Philopatry, the behaviour of returning to have offspring at the location of birth, could also be a driving factor in aggregations. Finally, individuals could be attracted to conspecifics, as their presence could signal good-quality habitat. Nesting aggregation preferences have yet to be explicitely tested in cavity0nesting bees, despite trapnests and bee hotels commonly being used for research and citizen-based conservation. While there is evidence that cavity-nesting *Osmia* bees in one Colorado populationdo not consider presence of other individuals while choosing where to nest [@groulx2018], it remains unclear how aggregation density and parasitism interact in nest-site selection of cavity-nesting bees.

## Objectives

I began to address this research gap by studying nest-site selection of non-eusocial cavity-nesting bees by measuring their preferences for aggregations and avoidance of parasites. In the Greenbelt of Ottawa, I tested nesting preferences of orchard bees (*Osmia* spp.) and leafcutter bees (*Megachile* spp.). My two primary questions are:

1.  **Do cavity-nesting bees have aggregation density preferences**? If bees have aggregation density preferences, they could prefer nesting in higher density aggregations because the presence of nesting conspecifics could indicate the presence of high-quality nesting habitat. Alternatively, bees could avoid the presence of conspecifics if density increases parasitism risk, thus prefer nesting in lower aggregation densities. If cavity-nesting bees do not consider aggregation density while selecting a nesting location, bees must use other nest-site characteristics when choosing where to build a nest.

2.  **How does parasitism interact with nesting preferences**?

    A. I will investigate the **relation between parasites and aggregation densities**. Parasites might target higher densities of nests because the presence of many host nests indicates more opportunities for parasitism. On the other hand, parasites may target lower densities of nests because there could be fewer chances of directly encountering a host, which could interrupt the act of parasitizing a nest.

    B. I will measure **how parasitism rates affect nesting preferences** of bees. If parasites are density-dependent, bees may alter their own preferences in attempt to reduce chances of parasitism. However, regardless if parasites are density-dependent, their presence alone may influence nest-site selection by bees directly avoiding parasites.

    C. I will measure **how the presence of parasites effects nest investment**. If bees can detect recent parasite activity in their own nest or adjacent nests, perhaps they reduce their investment into building their nest.

## Methods

I deployed trapnests containing 18 cavities each. The first half of the summer was dedicated towards the early-season orchard bees (*Osmia* spp.), while the second half was dedicated towards the later-season leafcutter bees (*Megachile* spp.). For each half of the summer, I deployed 36 trapnests spread evenly between two sites, for a total of 72. Trapnests were divided into three treatments to simulate variations in occupancy:

-   **Low-occupancy treatment**: the baseline of completed nests is zero. Once a bee nest is completed, we removed it for storage at the University of Ottawa in falcon tubes. The trapnest treatment should have little to no bee activity.

-   **Mid-occupancy treatment**: the baseline of completed nests was 6 out of 18 cavities occupied. Occupancy was simulated with pseudo nests built from old nesting material from real orchard or leafcutter bee nests, or the equivalent to what bees would use for building (mud for orchard bee season, leaves for leafcutter bee season). Once real nests were built, pseudo nests were removed to maintain baseline occupancy. If real nests surpassed the baseline occupancy, we removed them for storage in falcon tubes to maintain the baseline density.

-   **High-occupancy treatment**: the baseline of completed nests was 11 out of 18 cavities. Maintaining the baseline occupancy was identical to methods for the mid-occupancy treatment.

In total, I have approximately 164 *Osmia* nests and 58 *Megachile* nests (IDed based on nesting material).

![Figure 1: Example of a trapnest with various types of "occupied" cavities. Pseudo nests are empty inside, but appear to be sealed nests from the outside. Bee nests may either be completed (sealed) or in progress or abandoned (unsealed). Bee nests may also be parasitized. Non-target insects may be occupying the cavties as well, such as wasps or bees other than *Osmia* or *Megachile*.](Figures_files/figure-html/Labelled_nest.png)

```{r Prepping, include=FALSE}
library(ggplot2)
library(ggeffects)
library(kableExtra)
library(lme4)
library(tidyverse)


nesting <- read_csv("data/nesting_cleaned_all_bees.csv")
nesting_os <- read_csv("data/nesting_cleaned_osmia.csv")
nesting_meg <- read_csv("data/nesting_cleaned_megachile.csv") # three columns with parasite and cell info blank, not dissected
nesting_os_xs <- read_csv("data/nesting_cleaned_osmia_xs.csv")
nesting_meg_xs <- read_csv("data/nesting_cleaned_megachile_xs.csv")
investment <- read_csv("data/investment_cleaned.csv")
investment_os <- investment %>% filter(Occupant == "osmia")
parasite_sum_os_contains_cells <- read_csv("Data/parasite_sum_os_contains_cells.csv")
nesting_cells_os <- read_csv("data/nesting_cells_cleaned_osia.csv")


# number of nests
nesting_os %>%
  filter(!is.na(Bee_choice)) %>%
  summarize(nests_total = sum(Bee_choice)) # 164 osmia nests for osmia through the summer
nesting_meg %>%
  filter(!is.na(Bee_choice)) %>%
  summarize(nests_total = sum(Bee_choice)) # 58 megachile nests for osmia through the summer


# making the model that isn't overdispersed that I can show
# load('mod_choice_cav_os.Rdata')


theme_set(theme_classic(base_size = 15))

library("RColorBrewer")
display.brewer.pal(n = 9, name = "Set2")
brewer.pal(n = 9, name = "Oranges")
display.brewer.pal(n = 9, name = "Blues")
brewer.pal(n = 9, name = "Blues")




## MEGACHILE

meg <- nesting %>% filter(Occupant == "megachile" & Status != "pseudo" & Bee_choice == 1)
hist(meg$Day_of_year)

library(ggplot2)

ggplot(meg, aes(x = Day_of_year)) +
  geom_freqpoly(bins = 10)
```

## Nesting per conspecific presence (including pseudo nests)

```{r Conspecifics_with_pseudo, echo=FALSE, warning=FALSE, message=FALSE}
### Probability

# create a probability by successes/ successes + failures
os_fig_data <- nesting_os_xs %>% mutate(Probability_new_nest = New_nests_os / Available_cavities_prior)

## Osmia

os_fig_data %>%
  ggplot(aes(x = Osmia_ps_prior, y = Probability_new_nest)) +
  geom_jitter(colour = "goldenrod2", alpha = 0.5, height = 0.05, width = 0.1) +
  stat_smooth(method = "glm", colour = "goldenrod2", fill = "goldenrod2") +
  labs(x = "Number of " ~ italic(Osmia) ~ " in trapnest", y = "Number of new nests / \n Number of potential nests") +
  theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black")) +
  xlim(0, 18) +
  ylim(0, 1)



### Raw data

## Osmia

nesting_os_xs %>%
  ggplot(aes(Osmia_ps_prior, New_nests_os)) +
  geom_jitter(height = 0.05, alpha = 0.5, colour = "steelblue2") +
  geom_smooth(method = "lm", colour = "indianred3", fill = "indianred3") +
  labs(x = "Osmia", y = "Number of new nests each visit") +
  xlim(0, 18)

## Megachile

nesting_meg_xs %>%
  ggplot(aes(Meg_ps_prior, New_nests_meg)) +
  geom_jitter(height = 0.05, alpha = 0.5, colour = "goldenrod2") +
  geom_smooth(method = "glm", colour = "indianred3", fill = "indianred3") +
  labs(x = "Available cavities", y = "Number of new nests each visit") +
  xlim(0, 18)
```

## Nesting per conspecific presence (excluding pseudo nests)

```{r Conspecifics_without_pseudo, echo=FALSE, warning=FALSE, message=FALSE}
### Osmia

nesting_os_xs %>%
  ggplot(aes(Osmia_no_ps_prior, New_nests_os)) +
  geom_jitter(height = 0.05, alpha = 0.5, colour = "steelblue2") +
  geom_smooth(method = "glm", colour = "rosybrown", fill = "rosybrown") +
  labs(x = "Available cavities", y = "Number of new nests each visit") +
  xlim(0, 18)

### Megachile

nesting_meg_xs %>%
  ggplot(aes(Meg_no_ps_prior, New_nests_meg)) +
  geom_jitter(height = 0.05, alpha = 0.5, colour = "goldenrod2") +
  geom_smooth(method = "glm", colour = "rosybrown", fill = "rosybrown") +
  labs(x = "Available cavities", y = "Number of new nests each visit") +
  xlim(0, 18)
```

## Parasitism

```{r Wrangling Parasitism data, ignore=TRUE}
## Change data to make a stacked bar plot

cells_os <- pivot_longer(nesting_cells_os,
  cols = c(Sum_cell_not_parasitized, Sum_cell_parasitized),
  names_to = "Parasitized",
  values_to = "Cell_count"
)


cells_os$Parasitized <- factor(cells_os$Parasitized, levels = c("Sum_cell_parasitized", "Sum_cell_not_parasitized"))
```

```{r Parasitism, echo=FALSE, warning=FALSE, message=FALSE}
# Relation between parasites and aggregation densities, with pseudo

cells_os %>%
  ggplot(aes(Osmia_ps_no_iw, Cell_count, fill = Parasitized, colour = Parasitized)) +
  geom_bar(stat = "identity") + # to make this a proportion of cells, add "position='fill'"
  scale_fill_manual(values = c("plum", "seagreen3")) +
  scale_colour_manual(values = c("plum", "seagreen3"))


# Relation between parasites and aggregation densities, without pseudo

cells_os %>%
  ggplot(aes(Osmia_no_ps_no_iw, Cell_count, fill = Parasitized, colour = Parasitized)) +
  geom_bar(stat = "identity") + # to make this a proportion of cells, add "position='fill'"
  scale_fill_manual(values = c("plum", "seagreen3")) +
  scale_colour_manual(values = c("plum", "seagreen3"))


# Relation between parasites and aggregation densities, only unsealed

cells_os %>%
  ggplot(aes(Sum_unsealed_os_no_iw, Cell_count, fill = Parasitized, colour = Parasitized)) +
  geom_bar(stat = "identity") + # to make this a proportion of cells, add "position='fill'"
  scale_fill_manual(values = c("tomato2", "goldenrod2"), labels = c("Parasitized", "Not Parasitized"), name = ("Cell status")) +
  scale_colour_manual(values = c("tomato2", "goldenrod2"), labels = c("Parasitized", "Not Parasitized"), name = ("Cell status")) +
  labs(y = "Number of cells", x = "Number of active " ~ italic(Osmia) ~ " nests") +
  theme(legend.title = element_blank()) +
  theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black"))
```

```{r Preferences and Parasites}
# make proportion of new nests
nests_os_xs <- nesting_os_xs %>% mutate(Proportion_new_nests_os = New_nests_os / Available_cavities_prior)

# scale DP_os_prior
nests_os_xs$DP_os_scaled <- scale(nests_os_xs$DP_os_prior)

# make numeric
nests_os_xs$DP_os_scaled <- as.numeric(nests_os_xs$DP_os_scaled)

# make mean + and - 1 SD
nests_os_xs <- nests_os_xs %>%
  mutate(
    mean_DP_os_scaled = mean(DP_os_scaled, na.rm = TRUE),
    sd_DP_os_scaled = sd(DP_os_scaled, na.rm = TRUE),
    DP_os_category = case_when(
      DP_os_scaled < mean_DP_os_scaled - sd_DP_os_scaled ~ "Mean - 1 SD",
      DP_os_scaled > mean_DP_os_scaled + sd_DP_os_scaled ~ "Mean + 1 SD",
      TRUE ~ "Mean"
    )
  ) %>%
  select(-mean_DP_os_scaled, -sd_DP_os_scaled) # Optionally remove the intermediate columns


mean(nests_os_xs$DP_os_scaled)



# How the presence of parasites effects nesting preferences, with pseudo

# With interaction

nests_os_xs %>%
  ggplot(aes(NDP_ps_os_prior, Proportion_new_nests_os, colour = DP_os_category, group = DP_os_category)) +
  geom_jitter(colour = "goldenrod2", height = 0.1, width = 0.1, alpha = 0.6) +
  geom_smooth(method = "glm", aes(colour = DP_os_category, fill = DP_os_category)) +
  scale_colour_manual(values = c("tomato3", "tomato"), name = ("Number of nests with \ndetectable parasites"), breaks = c("Mean + 1 SD", "Mean")) +
  scale_fill_manual(values = c("tomato3", "tomato"), name = ("Number of nests with \ndetectable parasites"), breaks = c("Mean + 1 SD", "Mean")) +
  labs(y = "Number of new nests / \n Number of potential nests", x = "Number of nests without \ndetectable parasites") +
  theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black")) +
  xlim(0, 18) +
  ylim(0, 1)


# Without interaction

nests_os_xs %>%
  ggplot(aes(DP_os_prior, Proportion_new_nests_os)) +
  geom_jitter(colour = "goldenrod2", height = 0.1, width = 0.1, alpha = 0.6) +
  geom_smooth(method = "glm", colour = "tomato2", fill = "tomato2") +
  labs(y = "Number of new nests / \n Number of potential nests", x = "Number of nests with detectable parasites") +
  theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black")) +
  xlim(0, 3.5) +
  ylim(0, 1)
```

The most recent figure shows osmia density of unsealed nests, but shows cumulative parasites and doesn't count the current nest

## Nest Investment

```{r Last Cell, echo=FALSE, warning=FALSE, message=FALSE}
# Is the last cell more likely to be parasitized?
cells_os$Last_cell <- as.factor(cells_os$Last_cell)


cells_os %>%
  ggplot(aes(Last_cell, Cell_count, fill = Parasitized, colour = Parasitized)) +
  geom_bar(stat = "identity") + # to make this a proportion of cells, add "position='fill'"
  scale_fill_manual(values = c("tomato2", "goldenrod2"), labels = c("Parasitized", "Not Parasitized"), name = ("Cell status")) +
  scale_colour_manual(values = c("tomato2", "goldenrod2"), labels = c("Parasitized", "Not Parasitized"), name = ("Cell status")) +
  labs(y = "Number of cells", x = "Cell position") +
  scale_x_discrete(labels = c("Not last cell", "Last cell")) +
  theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black"))
```

```{r Investment, echo=FALSE, warning=FALSE, message=FALSE}
# Editing data for figure

investment_os <- investment_os %>% mutate(Prop_filled = Filled_length / Straw_length)


# Investment into a nest if other active nests or the nest itself is parasitized


investment_os %>%
  filter(!is.na(Nest_parasitized)) %>%
  ggplot(aes(DP_os_prior, Prop_filled, fill = Nest_parasitized)) +
  geom_smooth(method = "glm",
    method.args = list(family = binomial()),
    aes(colour = Nest_parasitized)) +
  geom_jitter(aes(colour = Nest_parasitized), 
    alpha = 0.7, width = 0.05, height = 0.01) +
  scale_colour_manual(values = c("goldenrod2", "tomato2"),
    labels = c("Not parasitized", "Parasitized"),
    name = "Nest status") +
  scale_fill_manual(values = c("goldenrod2", "tomato2"), labels = c("Not parasitized", "Parasitized"), name = "Nest status") +
  labs(y = "Proportion of focal nest filled with cells", x = "Number of surrounding nests with detectable parasites") +
  theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black"))
```

```{r Old Figures, include=FALSE}
#
# ## Make summary for proportion of new nests per number of Available_cavities_prior
#
# ### Osmia
#
# # sum the new nests per each Available_cavities_prior value
# summary_os_avlb <- nesting_os %>%
#   group_by(Available_cavities_prior) %>%
#   mutate(New_nests_by_avlb = sum(Bee_choice, na.rm = TRUE)) %>%
#   ungroup()
# # divide by total number of osmia nests
# summary_os_avlb <- summary_os_avlb %>%
#   mutate(Prop_new_per_available = New_nests_by_avlb/164)
# # deleting duplicates
# summary_os_avlb <- summary_os_avlb[!duplicated(summary_os_avlb[c('Available_cavities_prior','Prop_new_per_available')]),]
#
#
# ### Megachile
#
# # sum the new nests per each Available_cavities_prior value
# summary_meg_avlb <- nesting_meg %>%
#   group_by(Available_cavities_prior) %>%
#   mutate(New_nests_by_avlb = sum(Bee_choice, na.rm = TRUE)) %>%
#   ungroup()
# # divide by total number of megachile nests
# summary_meg_avlb <- summary_meg_avlb %>%
#   mutate(Prop_new_per_available = New_nests_by_avlb/58)
# # deleting duplicates
# summary_meg_avlb <- summary_meg_avlb[!duplicated(summary_meg_avlb[c('Available_cavities_prior', 'Prop_new_per_available')]),]
#
# # Figures
#
# summary_os_avlb %>%
#   ggplot(aes(Available_cavities_prior, Prop_new_per_available)) +
#   geom_jitter(height = 0, width = 0) +
#   stat_smooth(color = 'plum', fill = 'plum') +
#   labs(y = 'Proportion of new Osmia nests', x = 'Number of available cavities prior to nesting') +
#   xlim(0, 18) +
#   ylim(-0.1, 0.5) +
#   geom_text(
#     label = summary_os_avlb$New_nests_by_avlb, nudge_y = 0.01)
#
# summary_meg_avlb %>%
#   ggplot(aes(Available_cavities_prior, Prop_new_per_available)) +
#   geom_jitter(height = 0, width = 0) +
#   stat_smooth(color = 'goldenrod2', fill = 'goldenrod2') +
#   labs(y = 'Proportion of new Megachile nests', x = 'Number of available cavities prior to nesting') +
#   xlim(0, 18) +
#   ylim(-0.1, 0.5) +
#   geom_text(
#     label = summary_meg_avlb$New_nests_by_avlb, nudge_y = 0.01)
#
#
#
# ## Make summary for proportion of new nests per number of available cavities within each X axis value
#
# ### Osmia
#
# # make sum of new nests, then counts (0s and 1s)
# summary_os_avlb_2 <- nesting_os %>%
#   group_by(Available_cavities_prior) %>%
#   mutate(New_nests_by_avlb = sum(Bee_choice, na.rm = TRUE)) %>%
#   ungroup()
# summary_os_avlb_2 <- summary_os_avlb_2 %>%
#   group_by(Available_cavities_prior) %>%
#   mutate(Possible_occurrences_by_avlb = length(Bee_choice)) %>%
#   ungroup()
# # make into a proportion of yeses
# summary_os_avlb_2 <- summary_os_avlb_2 %>%
#   mutate(Prop_yes = New_nests_by_avlb/Possible_occurrences_by_avlb)
# # deleting duplicates
# summary_os_avlb_2 <- summary_os_avlb_2[!duplicated(summary_os_avlb_2[c('Available_cavities_prior','Prop_yes')]),]
#
#
# ### Megachile
#
# # make sum of new nests, then counts (0s and 1s)
# summary_meg_avlb_2 <- nesting_meg %>%
#   group_by(Available_cavities_prior) %>%
#   mutate(New_nests_by_avlb = sum(Bee_choice, na.rm = TRUE)) %>%
#   ungroup()
# summary_meg_avlb_2 <- summary_meg_avlb_2 %>%
#   group_by(Available_cavities_prior) %>%
#   mutate(Possible_occurrences_by_avlb = length(Bee_choice)) %>%
#   ungroup()
# # make into a proportion of yeses
# summary_meg_avlb_2 <- summary_meg_avlb_2 %>%
#   mutate(Prop_yes = New_nests_by_avlb/Possible_occurrences_by_avlb)
# # deleting duplicates
# summary_meg_avlb_2 <- summary_meg_avlb_2[!duplicated(summary_meg_avlb_2[c('Available_cavities_prior','Prop_yes')]),]
#
#
# # Figures
#
# summary_os_avlb_2 %>%
#   ggplot(aes(Available_cavities_prior, Prop_yes)) +
#   geom_jitter(height = 0, width = 0) +
#   geom_smooth(color = 'plum', fill = 'plum') +
#   xlim(0, 18) +
#   ylim(-0.1, 0.5) +
#   labs(y = 'Proportion of visits that are 1s', x = 'Number of available cavities prior to nesting')
#
# summary_meg_avlb_2 %>%
#   ggplot(aes(Available_cavities_prior, Prop_yes)) +
#   geom_jitter(height = 0, width = 0) +
#   geom_smooth(color = 'goldenrod2', fill = 'goldenrod2') +
#   xlim(0, 18) +
#   ylim(-0.1, 0.5) +
#   labs(y = 'Proportion of visits that are 1s', x = 'Number of available cavities prior to nesting')
#
#
#
#
#
```
