---
title: "Analyses"
format: html
editor: visual
---


------------------------------------------------------------------------

```{r Notes, include=FALSE, echo=TRUE, results='hide'}

## To Do List ##

 # make loop for how long the nest is active for
 # NN I 26 1a was a large pseudo until June 22, then a medium after


## Columns ##
# Nest_ID: the trapnest number, cavity number, and straw ID put together 
# Index: you can ignore this. It gave every row a unique number that I needed for a loop.
# Status: the activity of that cavity, e.g., its empty, new cells are made, a nest is sealed off, etc.
# Bee_choice: is a new nest built or not, Yes = 1, no = 0, pseudo nests or nests already in progress or completed get an NA.
# Available_post: number of available cavities for each trapnest and date combo AFTER the new nests for that visit have been started (i.e., how many available cavities that the experiments saw when arriving to the nest)
# New_nests_os: number of new osmia nests for each trapnest and date combo
# Osmia_ps_prior: number of existing osmia in each trapnest per date, including pseudo nests
# Osmia_no_ps: number of existing osmia in each trapnest per date, excluding pseudo nests
# New_nests_meg: number of new megachile nests for each trapnest and date combo
# Meg_ps_prior: number of existing megachile in each trapnest per date, including pseudo nests
# Meg_no_ps: number of existing megachile in each trapnest per date, excluding pseudo nests
# New_nests_sum: total number of new nests including megachile and osmia.
# Available_prior: average number of available cavities for each trapnest and date combo BEFORE the new nests for that visit have been started (i.e., how many available cavities the bees saw when arriving to the nest before building). This is the average, so there are some numbers that have .5 in there. For example, if there are 8 cavities available_post and two new nests were built when we visited that trapnest, the first bee to nest would have seen 10 available cavities, and the second would have seen 9. So we average that out to 9.5 for available_prior.
# Available_post = + Osmia_ps_prior + Meg_ps_prior (+ non-target nests) = 18



```

```{r Cleaning Data, include=FALSE, echo=TRUE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(ggplot2)
library(ggeffects)
library(kableExtra)
library(sqldf)
library(glmmTMB)
library(DHARMa)
library(car)
library(lme4)
# citation() # to see packages


## Load data
data <- read_csv('Data/Aggregate_nesting_data.csv')

## Add in straws that remained empty throughout the experiment

# add empty data frame
empty_data <- read_csv('Data/Empty_data.csv')
# add columns to say 'for binding' or NA to tell data frames apart and prioritize the collected data
empty_data$For_binding <- 0
data$For_binding <- 1
# merge data frames
data <- bind_rows(empty_data, data) 
# make nest ID column
data$Nest_ID <- with(data, paste(Trapnest, paste0(Cavity_number, Straw_ID), sep = '_'))
# make cavity ID column
data$Cavity_ID <- with(data, paste(Trapnest, Cavity_number, sep = '_'))
# make Trapnest a character
data$Trapnest <- paste('trapnest', data$Trapnest, sep = '_')

# to delete duplicate rows, check if it adds up
  length(empty_data$Straw_ID) # 1295 (minus one for the header)
  # length(data$Straw_ID) data when first reading is 1110 (minus header)
  sum(duplicated(data$Nest_ID)) # 701 duplicates
# final data frame should be 1295+1110-701 = 1704
data <- data %>% 
    group_by(Nest_ID) %>% 
    filter(For_binding == max(For_binding))
length(data$Nest_ID) # 1707, few extras.... ### CHECK ###


## Pivot longer and inspect
long_data <- pivot_longer(data, '06-May':'09-Aug', names_to = 'Date', values_to = 'Status')
table(long_data$Status)
table(long_data$Occupant)
# Replace some input for statuses of how they should be treated
long_data$Status <- replace(long_data$Status, long_data$Status=='eaten','empty')
long_data$Status <- replace(long_data$Status, long_data$Status=='inner wall started','empty')
long_data$Status <- replace(long_data$Status, long_data$Status=='non-target inner wall started','empty')
long_data$Status <- replace(long_data$Status, long_data$Status=='megachile?','megachile') # for now, but not working???


## Make the date column actually a date
# divide Date column from day-month into two columns
long_data <- long_data %>% 
  separate(Date, c('day', 'month')) 
# inspect
table(long_data$day)
table(long_data$month)
# replace month names
long_data$month <- replace(long_data$month, long_data$month == 'Jun','June')
long_data$month <- replace(long_data$month, long_data$month == 'Aug', 'August')
long_data$month <- replace(long_data$month, long_data$month == 'Jul', 'July')
# make new column with months as numbers
long_data$month2 <- match(long_data$month, month.name) 
# add column for the year
long_data <- long_data %>% mutate(Year = 2023)
# make a column with the full date that is read by R as a date
long_data$Date <- as.Date(with(long_data, paste(Year, month2, day,sep='-')), '%Y-%m-%d')
# check if R recognizes it as a date
library(lubridate)
sapply(long_data, is.Date)


## Fixing the Occupant column
# at first, this column says the occupant for the entire data collection period, even if the nest if empty for most of it
# make a new column, needs to ne 'NA' of loop wont run
long_data <- long_data %>% mutate(Occupant.v2 = ifelse(Status == 'non_existent', 'NA', ifelse(Status == 'empty', 'none', Occupant)))
long_data <- long_data %>% mutate(Occupant.v3 = ifelse(Status == 'no_visit', 'NA', Occupant.v2))
# add occupant for pseudo nests, Osmia first half of summer, Megachile second half
long_os <- long_data %>% 
  filter(Site == 'NN' | Subsite == 'A' | Subsite == 'B' | Subsite == 'C' | Subsite == 'D' | Subsite == 'E' | Subsite == 'F') %>%
  mutate(Occupant.v3 = ifelse(Status == 'pseudo', 'osmia', Occupant.v3))
# add megachile for pseudo nest type 
long_meg <- long_data %>% 
  filter(Site == 'CP' | Subsite == 'M' | Subsite == 'N' | Subsite == 'O' | Subsite == 'P' | Subsite == 'Q' | Subsite == 'R') %>%
  mutate(Occupant.v3 = ifelse(Status == 'pseudo', 'megachile', Occupant.v3))
# merge back together
long_data <- rbind(long_os, long_meg)
# Take columns that I want
long_data <- long_data %>% select(Site:Straw_ID, Straw_length_mm, Filled_length_mm, Nest_ID, Cavity_ID, Occupant.v3, Date, Status)

# rename new occupant.v2 column
long_data <- rename(long_data, Occupant = Occupant.v3)

# see if each trapnest has 18 existing cavities (unique cavity IDs) for each visit
long_data %>% 
  filter(Status != 'non_existent' & Status != 'no_visit') %>% 
  group_by(Trapnest, Date) %>% 
  summarize(Count = length(unique(Nest_ID))) 


# remove chunks of "non_existent" for first and second half of the summer ### CHECK do i need other bp and cp?
long_data_BP <- long_data %>% filter(Subsite == 'A' | Subsite == 'B' | Subsite == 'C' | Subsite == 'D' | Subsite == 'E' | Subsite == 'F')
long_data_NN <- long_data %>% filter(Subsite == 'G' | Subsite == 'H' | Subsite == 'I' | Subsite == 'J' | Subsite == 'K' | Subsite == 'L')

long_data_BP <- long_data_BP %>% filter(Date != '2023-05-06' & Date != '2023-06-30' & Date != '2023-07-03' & Date != '2023-07-04' & Date != '2023-07-05' & Date != '2023-07-06' & Date != '2023-07-07' & Date != '2023-07-10' & Date != '2023-07-12' & Date != '2023-07-14' & Date != '2023-07-17' & Date != '2023-07-20' & Date != '2023-07-24' & Date != '2023-07-26' & Date != '2023-07-28' & Date != '2023-07-29' & Date != '2023-08-01' & Date != '2023-08-04' & Date != '2023-08-09')
long_data_NN <- long_data_NN %>% filter(Date != '2023-07-07' & Date != '2023-07-10' & Date != '2023-07-12' & Date != '2023-07-14' & Date != '2023-07-17' & Date != '2023-07-20' & Date != '2023-07-24' & Date != '2023-07-26' & Date != '2023-07-28' & Date != '2023-07-29' & Date != '2023-08-01' & Date != '2023-08-04' & Date != '2023-08-09')
long_data_second <- long_data %>% filter(Subsite == 'M' | Subsite == 'N' | Subsite == 'O' | Subsite == 'P' | Subsite == 'Q' | Subsite == 'R' | Subsite == 'S' | Subsite == 'T' | Subsite == 'U' | Subsite == 'V' | Subsite == 'W' | Subsite == 'X')
long_data_second <- long_data_second %>% filter(Date != '2023-05-06' & Date != '2023-05-07' & Date != '2023-05-11' & Date != '2023-05-12' & Date != '2023-05-15' & Date != '2023-05-16' & Date != '2023-05-17' & Date != '2023-05-18' & Date != '2023-05-19' & Date != '2023-05-21' & Date != '2023-05-22' & Date != '2023-05-23' & Date != '2023-05-25' & Date != '2023-05-26' & Date != '2023-05-28' & Date != '2023-05-29' & Date != '2023-05-31' & Date != '2023-06-01' & Date != '2023-06-02' & Date != '2023-06-04' & Date != '2023-06-05' & Date != '2023-06-08' & Date != '2023-06-09' & Date != '2023-06-10' & Date != '2023-06-11' & Date != '2023-06-12' & Date != '2023-06-14' & Date != '2023-06-15' & Date != '2023-06-16' & Date != '2023-06-19' & Date != '2023-06-20' & Date != '2023-06-22' & Date != '2023-06-23' & Date != '2023-06-29' & Date != '2023-06-30')

long_data <- rbind(long_data_BP, long_data_NN, long_data_second)


## Loop for making the column for new nests per visit at the cavity-level

# make an index
index <- (1:length(long_data$Site)) # make index of 1 to last row
long_i <- long_data
long_i$index <- index # making index column to count
long_i <- long_i %>% relocate('index') # puts it in the front
# initialize a table for bee choice
Bee_choice <- as.data.frame(matrix(NA,0,2))
colnames(Bee_choice)[1] <- 'index'
colnames(Bee_choice)[2]<-'Bee_choice'
# saying for i in each of the rows in index

### START
for (i in long_i$index) { 
  # make an i-1 and i-2 and i-3 variable
  i_status <- long_i$Status[i]
  i_minus_one_status <- long_i$Status[i-1]
  i_minus_two_status <- long_i$Status[i-2]
  i_minus_three_status <- long_i$Status[i-3]
  # initialize a choice variable
  choice <- 'blank'
  
  if(i_status == 'non_existent' | i_status == 'no change' | i_status == 'no_visit' | i_status == 'pseudo') {
    choice <- 'NA'
  }else{
    if(i_status == 'empty'){
      choice <- '0'
      }else{
        if(i_status != 'non_existent' & i_status != 'no change' & i_status != 'no_visit' & i_status != 'pseudo' & i_status != 'empty' & i_minus_one_status == 'non_existent' | i_minus_one_status == 'empty') {
        choice <- '1'
      }else{
        if(i_status != 'non_existent' & i_status != 'no change' & i_status != 'no_visit' & i_status != 'pseudo' & i_status != 'empty' & i_minus_one_status == 'no_visit' & i_minus_two_status == 'non_existent' | i_minus_two_status == 'empty') {
          choice <- '1'
        }else{ 
          if(i_status != 'non_existent' & i_status != 'no change' & i_status != 'no_visit' & i_status != 'pseudo' & i_status != 'empty' & i_minus_one_status != 'non_existent' | i_minus_one_status != 'empty') {
          choice <- 'NA' 
          }else{
            if(i_status != 'non_existent' & i_status != 'no change' & i_status != 'no_visit' & i_status != 'pseudo' & i_status != 'empty' & i_minus_one_status == 'no_visit' & i_minus_two_status != 'non_existent' & i_minus_two_status != 'empty') {
            choice <- 'NA'
            }else{ 
              if(i_status != 'non_existent' & i_status != 'no change' & i_status != 'no_visit' & i_status != 'pseudo' & i_status != 'empty' & i_minus_one_status == 'no_visit' & i_minus_two_status == 'no_visit' & i_minus_three_status != 'non_existent' & i_minus_three_status != 'empty') {choice <- 'leftover'}
              else{ 
                choice <- 'leftover'}
      }}}}}}

  occupant_i <- long_i$Occupant[i]
  if(occupant_i == 'wasp' | occupant_i == 'masked bee' | occupant_i == 'resin bee'){
    choice <- 'NA'
  } else {choice <- choice}

  # create Bee_choice row for our big table
  Bee_choice_i <- as.data.frame(matrix(NA,0,2))
  colnames(Bee_choice_i)[1] <- 'index'
  colnames(Bee_choice_i)[2]<-'Bee_choice'
  Bee_choice_i[nrow(Bee_choice_i)+1,] = c(i,choice)
  # bind the row to the big table
  Bee_choice <- rbind(Bee_choice, Bee_choice_i)
} ### END

# Join products together
Bee_choice$index <- as.numeric(Bee_choice$index)
long_choice <- long_i %>% left_join(Bee_choice)
long_choice$Bee_choice <- as.numeric(long_choice$Bee_choice)



```

```{r Osmia Nesting Preferences, include=FALSE, echo=TRUE, results='hide'}

### Osmia ###

### Question 1: Nesting according to available cavities ###

## Count the number of Available Cavities in each trapnest including pseudo
# make a new column to say available (1) or not (0), or NA
long_avail <- long_choice %>% 
  mutate(Available_bin = ifelse(Status == 'no_visit' | Status == 'non_existent', NA, ifelse(Status == 'empty', 1,  0)))
# make a summary column to say the number of available nests in each trapnest by date AFTER that visit's bees began nesting
summ_avail <- long_avail %>% 
  group_by(Trapnest, Date) %>% 
  summarize(Available_post = sum(Available_bin, na.rm = T))
# replace zeros with NA
summ_avail$Available_post <- replace(summ_avail$Available_post, summ_avail$Available_post == 0, NA)
# Available_post: number of available cavities for each trapnest and date combo

## Count number of New Osmia Nests per trapnest 
long_new_os <- filter(long_choice, !is.na(Bee_choice), Occupant == 'osmia' | Occupant == 'empty')
summ_new_os <- long_new_os %>%
  group_by(Trapnest, Date) %>%
  summarize(New_nests_os = sum(Bee_choice == 1))
# New_nests_os: number of new osmia nests for each trapnest and date combo

# Merge together
prefs_os <- merge(long_choice, summ_avail, all.x = TRUE) 
prefs_os <- merge(prefs_os,  summ_new_os, all.x = TRUE)


### Question 2: Nesting according to conspecific presence (with and without pseudo nests) ###

## Make column to say osmia present or not, including pseudo
prefs_os <- prefs_os %>% 
  mutate(Osmia_bi_ps = ifelse(Occupant == 'osmia', 1, 0)) 
## Make column to say osmia present or not, excluding pseudo
prefs_os <- prefs_os %>% 
  mutate(Osmia_bi_no_ps = ifelse(Occupant == 'osmia' & Status != 'pseudo', 1, 0)) 

## Count the number of genera in each nest
# with pseudo nests/ pseudo nests considered osmia occupancies
summ_osmia_ps <- prefs_os %>% 
  group_by(Trapnest, Date) %>% 
  summarize(Osmia_ps = (sum(Osmia_bi_ps)))
# Osmia_ps: number of osmia in each trapnest per date, including pseudo nests

# without pseudo nests (pseudo nests not considered osmia occupancies)
summ_osmia_no_ps <- prefs_os %>% 
  group_by(Trapnest, Date) %>% 
  summarize(Osmia_no_ps = (sum(Osmia_bi_no_ps)))
# Osmia_no_ps: number of osmia in each trapnest per date, excluding pseudo nests

# Merge together
prefs_os <- merge(prefs_os,  summ_osmia_ps, all.x = TRUE)
prefs_os <- merge(prefs_os,  summ_osmia_no_ps, all.x = TRUE)

# Delete columns I dont need
prefs_os <- prefs_os %>% select(-Osmia_bi_ps, -Osmia_bi_no_ps )


```

```{r Megachile Nesting Preferences, include=FALSE, echo=TRUE, results='hide'}

### Megachile ###

### Question 1: Nesting according to available cavities ###

## Count the number of Available Cavities in each nest including pseudo, done in previous chunk
# Available_post: number of available cavities for each trapnest and date combo

## Count number of New Nests per trapnest 
long_new_meg <- filter(long_choice, !is.na(Bee_choice), Occupant == 'megachile' | Occupant == 'empty')
summ_new_meg <- long_new_meg %>%
  group_by(Trapnest, Date) %>%
  summarize(New_nests_meg = sum(Bee_choice == 1))
# New_nests_meg: number of new megachile nests for each trapnest and date combo

# Merge together
prefs_meg <- merge(long_choice, summ_avail, all.x = TRUE) 
prefs_meg <- merge(prefs_meg,  summ_new_meg, all.x = TRUE)


### Question 2: Nesting according to conspecific presence (with and without pseudo nests) ###

## Make column to say osmia present or not, including pseudo
prefs_meg <- prefs_meg %>% 
  mutate(Mega_bi_ps = ifelse(Occupant == 'megachile', 1, 0)) 
## Make column to say osmia present or not, excluding pseudo
prefs_meg <- prefs_meg %>% 
  mutate(Mega_bi_no_ps = ifelse(Occupant == 'megachile' & Status != 'pseudo', 1, 0)) 

## Count the number of genera in each nest
# with pseudo nests/ pseudo nests considered osmia occupancies
summ_Meg_ps <- prefs_meg %>% 
  group_by(Trapnest, Date) %>% 
  summarize(Meg_ps = (sum(Mega_bi_ps)))
# Meg_ps: number of megachile in each trapnest per date, including pseudo nests

# without pseudo nests/ pseudo nests not considered osmia occupancies
summ_Meg_no_ps <- prefs_meg %>% 
  group_by(Trapnest, Date) %>% 
  summarize(Meg_no_ps = (sum(Mega_bi_no_ps)))
# Meg_no_ps: number of megachile in each trapnest per date, excluding pseudo nests

# Merge together
prefs_meg <- merge(prefs_meg,  summ_Meg_ps, all.x = TRUE)
prefs_meg <- merge(prefs_meg,  summ_Meg_no_ps, all.x = TRUE)

# Delete columns I dont need
prefs_meg <- prefs_meg %>% select(-Mega_bi_ps, -Mega_bi_no_ps )

# Join Osmia and Megachile preferences together
prefs <- merge(prefs_os, prefs_meg)



```

```{r Cleaning Data 2}

## Cleaning required after nesting preferences data frame created

# remove rest of no_visits and non_existents (had to leave them in for the loop)
prefs <- prefs %>% filter(Status != 'no_visit' & Status != 'non_existent')


# make sure no duplicates of Bee_choice for new nests 
prefs.1 <- prefs %>% filter(Bee_choice == 1)
prefs.1$Bee_choice[duplicated(prefs.1$Nest_ID)] <- NA
prefs.0 <- prefs %>%
  filter(Bee_choice == 0 | is.na(Bee_choice))
prefs <- rbind(prefs.1, prefs.0)

# 26568 - 26529

```

```{r Weather Data, include=FALSE, echo=TRUE, results='hide'}


## Creating column with number of good days for bee activity based on weather between my visits to the trapnests 

weather <- read_csv('Data/Weather_data.csv')

# remove empty rows
weather <- weather[rowSums(is.na(weather)) != ncol(weather), ]
# remove empty columns
weather <- Filter(function(x)!all(is.na(x)), weather)

# replaces /NA/ with NA fr visit columns
weather <- weather %>% mutate(Visit_NN = ifelse(is.na(Visit_NN), 'NA', Visit_NN))
weather <- weather %>% mutate(Visit_CP = ifelse(is.na(Visit_CP), 'NA', Visit_CP))
weather <- weather %>% mutate(Visit_BP_Os = ifelse(is.na(Visit_BP_Os), 'NA', Visit_BP_Os))
weather <- weather %>% mutate(Visit_BP_Meg = ifelse(is.na(Visit_BP_Meg), 'NA', Visit_BP_Meg))

# replace /NA/ with zeros in weather count columns
weather <- weather %>% mutate(Good_days_NN = ifelse(is.na(Good_days_NN), 0, Good_days_NN))
weather <- weather %>% mutate(Good_days_CP = ifelse(is.na(Good_days_CP), 0, Good_days_CP))
weather <- weather %>% mutate(Good_days_BP_Os = ifelse(is.na(Good_days_BP_Os), 0, Good_days_BP_Os))
weather <- weather %>% mutate(Good_days_BP_Meg = ifelse(is.na(Good_days_BP_Meg), 0, Good_days_BP_Meg))

table(weather$Weather) # 31 bad days, 65 good days for bee activity  
str(weather)


## Bind together

# take only columns I need, for each site
w_NN <- weather %>% select(Date, Good_days_NN)
w_BP_Os <- weather %>% select(Date, Good_days_BP_Os)
w_BP_Meg <- weather %>% select(Date, Good_days_BP_Meg)
w_CP <- weather %>% select(Date, Good_days_CP)

# divide data frames into their sites to properly link up the weather data
NN <- prefs %>% filter(Site == 'NN')
BP_Os <- prefs %>% filter(Subsite == 'A' | Subsite == 'B' | Subsite == 'C' | Subsite == 'D' | Subsite == 'E' | Subsite == 'F')
BP_Meg <- prefs %>% filter(Subsite == 'M' | Subsite == 'N' | Subsite == 'O' | Subsite == 'P' | Subsite == 'Q' | Subsite == 'R')
CP <- prefs %>% filter(Site == 'CP')


# Join weather data 
pref_NN <- left_join(NN, w_NN)
pref_BP_Os <- left_join(BP_Os, w_BP_Os)
pref_BP_Meg <- left_join(BP_Meg, w_BP_Meg)
pref_CP <- left_join(CP, w_CP)
# rename good days column
pref_NN <- rename(pref_NN, Good_days = Good_days_NN)
pref_BP_Os <- rename(pref_BP_Os, Good_days = Good_days_BP_Os)
pref_BP_Meg <- rename(pref_BP_Meg, Good_days = Good_days_BP_Meg)
pref_CP <- rename(pref_CP, Good_days = Good_days_CP)

# bind together
preferences <- rbind(pref_BP_Os, pref_NN)
preferences <- rbind(preferences, pref_BP_Meg)
preferences <- rbind(preferences, pref_CP)

# compare new df to the old intact one, should be the same
length(prefs$Trapnest)
length(preferences$Trapnest)


```

```{r Calculating prior values, include=FALSE, echo=TRUE, results='hide'}


## Dealing with multiple nests built on the same day
# for available cavities
# and for congenera present 

# first, see the max and mean of new nests built on the same day
max(preferences$New_nests_os, na.rm = T) # 5
mean(preferences$New_nests_os, na.rm = T) # 1.54
hist(preferences$New_nests_os)
max(preferences$New_nests_meg, na.rm = T) # 5
mean(preferences$New_nests_meg, na.rm = T) # 1.42
hist(preferences$New_nests_meg)


## Changing available cavities for prior number

# given the infrequency of multiple nests built on the same day, it seems reasonable seems to average the number of new nests and add that to the available nests column. 
# e.g., if there was 4 new nests one visit to a low occupancy treatment trapnest, the Available_post column (# of available nests) should change from 14 to 16.

# Available_post + Osmia_ps + Meg_ps (+ non-target cells, not pictured) = 18

# 1- Add together the megachile and osmia nests on the same day 
preferences <- preferences %>% mutate(New_nest_sum = ifelse(!is.na(New_nests_os) & is.na(New_nests_meg), New_nests_os, ifelse(is.na(New_nests_os) & !is.na(New_nests_meg), New_nests_meg, (New_nests_os + New_nests_meg))))

# 2- Make a column to take the average of the available cavities that each bee could have encountered. E.g., if there were 2 new nests in a low-occupancy treatment with no other active nests, the first bee encountered 18 empty cavities while the second bee encountered 17 empty ones. Use (n+1)/2 which is equivalent to n+(n-1)+(n-2).../n. E.g., (3+1)/2=2, while (3+2+1)/3=2. n being the number of new nests.
# Remove occurences of zeros because (0+1)/2 = 0.5, and include occurences where there was new nest(s) of one genera but not the other

# make the equation
preferences <- preferences %>% mutate(Nest_equation = ifelse(New_nest_sum > 1, ((New_nest_sum + 1)/2), ifelse(New_nests_os == 1 | New_nests_meg == 1 & is.na(New_nest_sum), 1, 'leftover')))

# make nest equation numeric
preferences$Nest_equation <- as.numeric(preferences$Nest_equation)

# add equation and available post to make available prior
preferences$Available_prior <- (preferences$Nest_equation + preferences$Available_post)

# filter out wasps and non-targets from choices, after we counted them as occupied cavities
preferences <- filter(preferences, Occupant != 'wasp' & Occupant !='unkown' & Occupant !='masked bee')
table(preferences$Occupant)


## Changing occupant count for prior number

# need to make theses columns the number of occupencies PRIOR to the bee nesting, e.g., the new nests should not count in Osmia_ps and Osmia_no_ps
# If theres 1 new nest, make new column to say Osmia_(no_)ps - New_nests_os
# If there are multiple nests, to take the average, do Osmia_(no_)ps - (n-1)/2, n being the number of new nests per genera
# e.g., 1 new nest, 9 Osmia_ps = 9-1 = 8
# e.g, 3 new nests, 9 Osmia_ps (so only 6 filled the previous visit) = ((3-1)/2) + (9-3) = 7
# Osmia_ps_prior + Available_prior should usually = 18, unless there's megachile or non-target occupants 

# 1- for 0-1 new nests, (available_post - new nests)
  # osmia pseudo
preferences <- preferences %>% mutate(Osmia_ps_prior_V1 = ifelse(New_nests_os == 1, (Osmia_ps - New_nests_os), ifelse(is.na(Osmia_ps) | is.na(New_nests_os), NA, 'TBD'))) 
  # osmia no pseudo
preferences <- preferences %>% mutate(Osmia_no_ps_prior_V1 = ifelse(New_nests_os == 1, (Osmia_no_ps - New_nests_os), ifelse(is.na(Osmia_no_ps) | is.na(New_nests_os), NA, 'TBD')))  
  # megachile pseudo
preferences <- preferences %>% mutate(Meg_ps_prior_V1 = ifelse(New_nests_meg == 1, (Meg_ps - New_nests_meg), ifelse(is.na(Meg_ps) | is.na(New_nests_meg), NA, 'TBD'))) 
  # megachile no pseudo
preferences <- preferences %>% mutate(Meg_no_ps_prior_V1 = ifelse(New_nests_meg == 1, (Meg_no_ps - New_nests_meg), ifelse(is.na(Meg_no_ps) | is.na(New_nests_meg), NA, 'TBD'))) 

# 2 A- for multiple new nests, (n-1)/2
  # osmia
preferences <- preferences %>% mutate(Osmia_equ = ifelse(New_nests_os > 1, ((New_nests_os - 1)/2), NA)) 
  # megachile
preferences <- preferences %>% mutate(Meg_equ = ifelse(New_nests_meg > 1, ((New_nests_meg - 1)/2), NA)) 

# B- add equation + (available_post - new nests)
  # osmia pseudo
preferences <- preferences %>% mutate(Osmia_ps_prior = ifelse(Osmia_ps_prior_V1 == 'TBD', (Osmia_equ + Osmia_ps - New_nests_os), ifelse(is.na(Osmia_equ) & !is.na(Osmia_ps_prior_V1) | Osmia_ps_prior_V1 != 'TBD', Osmia_ps_prior_V1, 'leftover' ))) 
  # osmia no pseudo
preferences <- preferences %>% mutate(Osmia_no_ps_prior = ifelse(Osmia_no_ps_prior_V1 == 'TBD', (Osmia_equ + Osmia_no_ps - New_nests_os), ifelse(is.na(Osmia_equ) & !is.na(Osmia_no_ps_prior_V1) | Osmia_no_ps_prior_V1 != 'TBD', Osmia_no_ps_prior_V1, 'leftover' ))) 
  # megachile pseudo
preferences <- preferences %>% mutate(Meg_ps_prior = ifelse(Meg_ps_prior_V1 == 'TBD', (Meg_equ + Meg_ps - New_nests_meg), ifelse(is.na(Meg_equ) & !is.na(Meg_ps_prior_V1) | Meg_ps_prior_V1 != 'TBD', Meg_ps_prior_V1, 'leftover' ))) 
  # megachile no pseudo
preferences <- preferences %>% mutate(Meg_no_ps_prior = ifelse(Meg_no_ps_prior_V1 == 'TBD', (Meg_equ + Meg_no_ps - New_nests_meg), ifelse(is.na(Meg_equ) & !is.na(Meg_no_ps_prior_V1) | Meg_no_ps_prior_V1 != 'TBD', Meg_no_ps_prior_V1, 'leftover' ))) 

# take only columns I need
preferences <- preferences %>% select(Date, Site, Subsite, Treatment, Trapnest, Cavity_number, Straw_ID, Cavity_size, Filled_length_mm, Straw_length_mm, Nest_ID, Cavity_ID, Status, Occupant, Good_days, Bee_choice, Available_prior, New_nests_os, New_nests_meg, New_nest_sum, Osmia_ps_prior, Osmia_no_ps_prior, Meg_ps_prior, Meg_no_ps_prior)


# write.csv(preferences, 'preferences.csv')

```

```{r Parasitism Data, include=FALSE, echo=TRUE, results='hide'}

# Columns
 # Cumulative_par: cumulative number of parasites for that specific nest. E.g., 6_15a had no parasites when 1 cell was built, then two more cells each had a parasite, then six more cells with three parasites. So cumulative_par is 0, 2, then 5, and 5 until the nest is removed 
 # Number_of_cells: the cumulative number of cells in the nest at the time.
 # Prop_cells_par = Cumulative_par/Number_of_cells which gives the proportion of cells parasitized 

### Question 3: Do parasites affect nesting preferences ###

# load data
os_par <- read_csv('Data/Parasites_Osmia_clean.csv') # 728 rows

table(os_par$Cell_occupant)
# there will be some missing data even though we have the nest info because the nests were too moldy to dissect, and megachile nests are not logged yet e.g., 6_17a and 3_13b

# make nest and cell ID column
os_par$Nest_ID <- with(os_par, paste(Trapnest, paste0(Cavity_number, Straw_ID), sep = '_'))
os_par$Cell_ID <- with(os_par, paste(Nest_ID, Cell_number, sep = '_'))

# take number of cells total
os_par <- os_par %>% # 248 rows
    group_by(Nest_ID, Date) %>% 
    mutate(Number_of_cells = max(Cell_number))

# create column to say if cell is parasitized or not
os_par <- os_par %>% # 721 rows
  mutate(Cell_par = ifelse(Cell_occupant == 'sapyga', 1, 0)) 


## Make column to say nest is parasitized or not for when a parasite is in a cell, from that date onwards

# make column to say for each date, if a cell has parasites, then count cumulative parasites in a new column
os_par <- os_par %>% 
  group_by(Nest_ID) %>% 
  mutate(Cumulative_par = cumsum(Cell_par))
os_par <- os_par %>% 
  select(Trapnest, Date, Nest_ID, Cell_ID, Number_of_cells, Cell_occupant, Cell_par, Cumulative_par) # 721 rows
# bind parasite counts with preferences data
# preferences$Trapnest <- as.numeric(preferences$Trapnest)
os_par_pref <- select(preferences, Trapnest, Site:Straw_ID, Date, Filled_length_mm:Bee_choice, New_nests_os:Meg_no_ps_prior) # 24871 rows ### NOW 26060
os_par$Date <- as.Date(os_par$Date)
os_parasites <- merge(os_par_pref, os_par, all=T) # the status is repeating for each cell # 25356 NOW 25203

### CHECK FOR MISSING STATUSES ### these would be mistakes/ extras

# os_parasites$Trapnest <- as.numeric(os_parasites$Trapnest)
# IM TRYING TO SEE IF REMOVING THE DUPLOCATES REMOVES AGGREGATE DATA
os_parasites %>% filter(!is.na(New_nest_sum)) # with duplicates = 3246 # now 3317

# remove duplicates from repeating combos (more than one cell built in a nest in one day)
os_parasites <- os_parasites %>%  
  group_by(Date, Nest_ID, Cumulative_par) %>%
  distinct(Cumulative_par, .keep_all = TRUE)
os_parasites %>% filter(!is.na(New_nest_sum)) # without duplicates = 2927 Now 2980


# take max number of parasites for a nest_ID - date combo
# e.g., on 2023-06-14 for 13_18a (cells 1 and 2), both are parasitized, so take the highest parasite count for that date, aka say that for 13_18a on 2023-06-14, there's 2 parasites
os_parasites <- os_parasites %>% 
  group_by(Date, Nest_ID) %>% 
  mutate(Cumulative_par_max = max(Cumulative_par)) %>% 
  ungroup()
# remove duplicates of the same nest_ID - date combo, take only max number
os_parasites <- os_parasites %>%  
  group_by(Date, Nest_ID, Cumulative_par_max) %>% 
  distinct(Cumulative_par_max, .keep_all = TRUE) # 26031 NOW 26062


## Loop for making the cumulative_par_max value for the duration the nest is in the trapnest 

# make sure data is arranged chronologically
os_parasites <- os_parasites %>% arrange(Nest_ID, Date)
# initiate new column, then fill it in
os_parasites$Cumulative_par_V2 <- os_parasites$Cumulative_par 

# loop takes too long to do at once, so remove all empty and pseudo, then adding back in after. Tested this in the next chunk with trapnest 12
os_parasites_exclude <- os_parasites %>% filter(Status != 'empty' & Status != 'pseudo') # excluding empty and pseudo

### START
for(i in 1:nrow(os_parasites_exclude)){
  nestID <- os_parasites_exclude$Nest_ID[i] # focal nest for i-th row
  dt <- os_parasites_exclude$Date[i] # date for i-th row
  allprev <- os_parasites_exclude %>% filter(Nest_ID == nestID & Date <= dt) %>% pull(Cumulative_par_max) # all previous parasite counts for that focal nest
  
  check1 <- !all(is.na(allprev))
  check2 <- !os_parasites_exclude$Status[i] %in% c('empty', 'pseudo') 
  # you'll have to adjust this as needed for the intended behaviour & possible values in your larger dataset
  
  if(check1 & check2){
    os_parasites_exclude$Cumulative_par_V2[i] <- max(allprev, na.rm = T)
  }
} ### END



# bind back together
os_parasites_include <- os_parasites %>% filter(Status == 'empty' | Status == 'pseudo') # include empty and pseudo 
os_parasites_v2 <- rbind(os_parasites_exclude, os_parasites_include) # 26031
os_parasites_v2 <- os_parasites_v2 %>% arrange(Nest_ID, Date)
# remove unnecessary columns
os_parasites_v2 <- ungroup(os_parasites_v2)
os_parasites_v2 <- os_parasites_v2 %>% select(-Cell_ID, -Cumulative_par, -Cumulative_par_max, -Cell_par, -Cell_occupant)
# rename column and df
os_parasites_v2 <- os_parasites_v2 %>% rename(Cumulative_par = Cumulative_par_V2)
# x <- left_join(preferences, os_parasites_v2, all=T) # 25981

# checking for differences/ extra data
# preferences_list <- preferences %>% select(Date, Nest_ID, Status)
# os_parasites_list <- os_parasites_v2 %>% select(Date, Nest_ID, Status)
# extra_1 <- sqldf('SELECT * FROM preferences_list EXCEPT SELECT * FROM os_parasites_list') 
# extra_2 <- sqldf('SELECT * FROM os_parasites_list EXCEPT SELECT * FROM preferences_list') 
# figure out whats up with 18_8a/b ### CHECK ###


## Number of cells parasitized per number of cells in the nest 

os_parasites_v2$Number_of_cells <- as.numeric(os_parasites_v2$Number_of_cells)
os_parasites_v2$Osmia_ps_prior <- as.numeric(os_parasites_v2$Osmia_ps_prior)
os_parasites_v2$Osmia_no_ps_prior <- as.numeric(os_parasites_v2$Osmia_no_ps_prior)
preferences$Meg_ps_prior <- as.numeric(preferences$Meg_ps_prior)
preferences$Meg_no_ps_prior <- as.numeric(preferences$Meg_no_ps_prior)

nesting <- os_parasites_v2
nesting <- nesting %>% mutate(Prop_cells_par = (Cumulative_par/Number_of_cells)) 


## Number of parasites in that trapnest - date combo

nesting <- nesting %>%
  group_by(Date, Trapnest) %>%
  mutate(Par_in_trapnest = sum(!is.na(Cumulative_par))) %>%
  ungroup()



```

```{r Investment Data, include=FALSE, echo=TRUE, results='hide'}


### Question 4: Does parasitism affect nest investment ###

# make data frames

nesting$Straw_length_mm <- as.numeric(nesting$Straw_length_mm)
nesting$Filled_length_mm <- as.numeric(nesting$Filled_length_mm)


investment <- nesting %>%  
  filter(!is.na(Straw_length_mm) & Occupant != 'NA' & Occupant != 'none') %>% 
  mutate(Prop_filled = Filled_length_mm/Straw_length_mm) 
# some are still missing now because they mustve been moldy ### CHECK ###

investment_os <- investment %>% filter(Occupant == 'osmia')
investment_meg <- investment %>% filter(Occupant == 'megachile')

investment_os %>% 
  ggplot(aes(Prop_filled, Par_in_trapnest)) + 
  geom_jitter(height = 0, alpha = 0.4) + stat_smooth()

```

```{r Editing data for models, include=FALSE, echo=TRUE, results='hide'}

## Putting preferences, parasites, and investment into one dataframe 

# check structure
str(nesting)
nesting$Straw_length_mm <- as.numeric(nesting$Straw_length_mm)
nesting$Filled_length_mm <- as.numeric(nesting$Filled_length_mm)

# turn dates to day of year, because date wasn't working in the models
nesting <- nesting %>% mutate(Day_of_year = lubridate::yday(as.Date(Date,format='%d/%m/%Y')))
table(nesting$Day_of_year)


## Separate Osmia and Megachile data frames 

nesting_os <- nesting %>% filter(Occupant == 'osmia' | Occupant == 'none')
nesting_meg <- nesting %>% filter(Occupant == 'megachile' | Occupant == 'none') 
# for each model and filter out most off-season dates: Find the first and last dates of activity to measure the active season, and remove off-season data (this is still including the 'full summer' of data, while excluding the inflated 'zeros' for Bee_choice for when osmia and megachile wouldn't have been active)
dates_os <- nesting_os %>% 
  group_by(Date, Occupant) %>% 
  filter(Occupant == 'osmia' & Status != 'pseudo' & Status != 'no change')
  # earliest activity date is 2023-05-11
  # latest activity date is 2023-07-12
dates_meg <- nesting_meg %>% 
  group_by(Date, Occupant) %>% 
  filter(Occupant == 'megachile' & Status != 'pseudo' & Status != 'no change')
  # earliest activity date is 2023-06-23
  # latest activity date is 2023-08-04
list(unique(nesting$Date)) 
dates_exclude_os <- c('2023-07-14', '2023-07-17', '2023-07-20', '2023-07-24', '2023-07-26', '2023-07-28', '2023-07-29', '2023-08-01', '2023-08-04', '2023-08-09') ### CHECK make this the same as first cleaning chunk?
nesting_os <- nesting_os %>%
  filter(!Date %in% dates_exclude_os)
dates_exclude_meg <- c('2023-05-11', '2023-05-12', '2023-05-15', '2023-05-16', '2023-05-17', '2023-05-18', '2023-05-19', '2023-05-21', '2023-05-22', '2023-05-23', '2023-05-25', '2023-05-26', '2023-05-28', '2023-05-29', '2023-05-31', '2023-06-01', '2023-06-02', '2023-06-04', '2023-06-05', '2023-06-08', '2023-06-09', '2023-06-10', '2023-06-11', '2023-06-12', '2023-06-14', '2023-06-15', '2023-06-16', '2023-06-19', '2023-06-20', '2023-06-22', '2023-08-09')
nesting_meg <- nesting_meg %>%
  filter(!Date %in% dates_exclude_meg)


#  #  #  #  #  # 
nesting_os <- nesting_os %>% filter(Occupant == 'osmia' | Occupant == 'none')


#  #  #  #  #  #

```

------------------------------------------------------------------------

## Models

Variables:

-   Nest_ID: the trapnest number, cavity number, and straw ID put together
-   Index: you can ignore this. It gave every row a unique number that I needed for a loop.
-   Status: the activity of that cavity, e.g., its empty, new cells are made, a nest is sealed off, etc.
-   Bee_choice: is a new nest built or not, Yes = 1, no = 0, pseudo nests or nests already in progress or completed get an NA.
-   Available_post: number of available cavities for each trapnest and date combo AFTER the new nests for that visit have been started (i.e., how many available cavities that the experiments saw when arriving to the nest)
-   New_nests_os: number of new osmia nests for each trapnest and date combo
-   Osmia_ps_prior: number of existing osmia in each trapnest per date, including pseudo nests
-   Osmia_no_ps: number of existing osmia in each trapnest per date, excluding pseudo nests
-   New_nests_meg: number of new megachile nests for each trapnest and date combo
-   Meg_ps_prior: number of existing megachile in each trapnest per date, including pseudo nests
-   Meg_no_ps: number of existing megachile in each trapnest per date, excluding pseudo nests
-   New_nests_sum: total number of new nests including megachile and osmia.
-   Available_prior: average number of available cavities for each trapnest and date combo BEFORE the new nests for that visit have been started (i.e., how many available cavities the bees saw when arriving to the nest before building). This is the average, so there are some numbers that have .5 in there. For example, if there are 8 cavities available_post and two new nests were built when we visited that trapnest, the first bee to nest would have seen 10 available cavities, and the second would have seen 9. So we average that out to 9.5 for available_prior.

### Nesting per Available Cavities

```{r Models: Nest per Available Cavities}

### New nests per available cavities Osmia ###
mod_choice_cav_os <- glmer(Bee_choice ~ Available_prior + Good_days + Day_of_year + Site + (1|Subsite/Trapnest/Cavity_ID), data = Osmia_data, family = binomial) # boundary singular warning

plot(simulateResiduals(mod_c_cav_os)) # looks good jan 9
testDispersion(mod_c_cav_os, type = 'DHARMa') # ns
testOutliers(mod_c_cav_os, type = 'bootstrap') # ns p value = 1
summary(mod_c_cav_os)

opitimizers_os <- allFit(mod_c_cav_os) # did this to compare with model below that gets more warnings


### New nests per available cavities Megachile ###
mod_choice_cav_meg <- glmer(Bee_choice ~ scale(Available_prior) + scale(Good_days) + scale(Day_of_year)  + (1|Cavity_ID), data = Megachile_data, family = binomial) # Model failed to converge with max|grad|

# check optimizors and their fits
opitimizers <- allFit(mod_c_cav_meg)
# https://github.com/lme4/lme4/issues/120
relgrad <- with(mod_c_cav_meg@optinfo$derivs,solve(Hessian,gradient))
max(abs(relgrad))

plot(simulateResiduals(mod_c_cav_meg)) # WONKY! jan 9, but using only Cavity_ID is better then subsite, trapnest, and cavity ID
testDispersion(mod_c_cav_meg, type = 'DHARMa') # significant
testOutliers(mod_c_cav_meg, type = 'bootstrap') # ns p value = 1
summary(mod_c_cav_meg)

```

### Nesting per Conspecific Presence

```{r Models: Nest per Conspecifics}

### New nests per conspecifics Osmia ###
## with pseudo ##
mod_choice_consp_os_ps <- glmer(Bee_choice ~ Osmia_ps_prior + Good_days + Day_of_year + Site + (1|Subsite/Trapnest/Cavity_ID), data = Osmia_data, family = binomial) # boundary singular warning 

plot(simulateResiduals(mod_c_consp_os_ps)) # looks good jan 9
testDispersion(mod_c_consp_os_ps, type = 'DHARMa')  # ns
testOutliers(mod_c_consp_os_ps, type = 'bootstrap') # ns p value = 1
summary(mod_c_consp_os_ps)

### New nests per conspecifics Osmia ###
## without pseudo ##
mod_choice_consp_os_no_ps <- glmer(Bee_choice ~ Osmia_no_ps_prior + Good_days + Day_of_year + Site + (1|Subsite/Trapnest/Cavity_ID), data = Osmia_data, family = binomial) # boundary singular warning 

plot(simulateResiduals(mod_c_consp_os_no_ps)) # looks good jan 9
testDispersion(mod_c_consp_os_no_ps, type = 'DHARMa') # ns 
testOutliers(mod_c_consp_os_no_ps, type = 'bootstrap') # ns p value = 1
summary(mod_c_consp_os_no_ps)


### New nests per conspecifics Megachile ###
## with pseudo ##
mod_choice_consp_meg_ps <- glmer(Bee_choice ~ Meg_ps_prior + Good_days + Day_of_year + Site + (1|Subsite/Trapnest/Cavity_ID), data = Megachile_data, family = binomial)  # boundary singular warning 

plot(simulateResiduals(mod_c_consp_meg_ps)) # WONKY! jan 9
testDispersion(mod_c_consp_meg_ps, type = 'DHARMa') # significant
testOutliers(mod_c_consp_meg_ps, type = 'bootstrap') # ns p value = 1
summary(mod_c_consp_meg_ps)

### New nests per conspecifics Megachile ###
## without pseudo ##
mod_choice_consp_meg_no_ps <- glmer(Bee_choice ~ scale(Meg_no_ps_prior) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest/Cavity_ID), data = Megachile_data, family = binomial) # Model failed to converge with max|grad|

plot(simulateResiduals(mod_c_consp_meg_no_ps)) # WONKY! jan 9
testDispersion(mod_c_consp_meg_no_ps, type = 'DHARMa') # significant
testOutliers(mod_c_consp_meg_no_ps, type = 'bootstrap') # ns p value = 1 
summary(mod_c_consp_meg_no_ps)

```

------------------------------------------------------------------------

### Observed and Expected values of new nests

```{r Chi Square: on-season, include=FALSE, echo=TRUE, results='hide'}

### Only on-season data

## Make a summary of the observed new nests in each trapnest

# have to make new dataframe and pivot to converge date and trapnest data (removing cavity IDs)
for.table <- select(preferences, Trapnest, Subsite, Site, Date, Treatment,New_nests_os, New_nests_meg)
for.table <- distinct(for.table, Trapnest, Date, .keep_all = TRUE)

# count total nests
obs.os.1 <- for.table %>% filter(!is.na(New_nests_os)) %>% 
  group_by(Site, Subsite, Treatment) %>% 
  summarize(Total_nests = sum(New_nests_os))
obs.meg.1 <- for.table %>% filter(!is.na(New_nests_meg)) %>% 
  group_by(Site, Subsite, Treatment) %>% 
  summarize(Total_nests = sum(New_nests_meg))

# Osmia: use full season for BP and NN, since osmia had only 2 nests at CP
# Megachile: use full season for BP, since megchiole had only 5 nests at CP, and 3 at NN
obs.os.1 <- obs.os.1 %>% filter(Site == 'BP' | Site == 'NN')
obs.meg.1 <- obs.meg.1 %>% filter(Site == 'BP')

obs.os <- pivot_wider(obs.os.1, names_from = Treatment, values_from = Total_nests)
obs.meg <- pivot_wider(obs.meg.1, names_from = Treatment, values_from = Total_nests)

# replaces NAs
obs.os <- replace(obs.os, is.na(obs.os), 0)
obs.meg <- replace(obs.meg, is.na(obs.meg), 0)

# sum data
obs.os <- obs.os %>%
  group_by(Site) %>% 
  mutate(Total.Low = sum(Low), Total.Mid = sum(Mid), Total.High = sum(High))
obs.meg <- obs.meg %>%
  group_by(Site) %>% 
  mutate(Total.Low = sum(Low), Total.Mid = sum(Mid), Total.High = sum(High))

# remove unwanted columns
obs.os <- obs.os %>% select(Site, Total.Low, Total.Mid, Total.High)
obs.meg <- obs.meg %>% select(Site, Total.Low, Total.Mid, Total.High)

# remove duplicates 
obs.os <- obs.os %>% distinct(Site, .keep_all = TRUE)
obs.meg <- obs.meg %>% distinct(Site, .keep_all = TRUE)

# make sum per site
obs.os$Total <- obs.os$Total.High + obs.os$Total.Mid + obs.os$Total.Low
obs.meg$Total <- obs.meg$Total.High + obs.meg$Total.Mid + obs.meg$Total.Low


## Make a summary of expected new nests in each trapnest

exp.os <- select(obs.os, Site, Total)
exp.meg <- select(obs.meg, Site, Total)

# number nests observed * (available cavities in treatment / 37)
exp.os$Low <- (exp.os$Total * (18/37))
exp.os$Mid <- (exp.os$Total * (12/37))
exp.os$High <- (exp.os$Total * (7/37))

exp.meg$Low <- (exp.meg$Total * (18/37))
exp.meg$Mid <- (exp.meg$Total * (12/37))
exp.meg$High <- (exp.meg$Total * (7/37))

# remove total column
obs.os <- obs.os %>% select(-Total)
obs.meg <- obs.meg %>% select(-Total)
exp.os <- exp.os %>% select(-Total)
exp.meg <- exp.meg %>% select(-Total)

# rename column headings
obs.os <- rename(obs.os, Low = Total.Low, Mid = Total.Mid, High = Total.High)
obs.meg <- rename(obs.meg, Low = Total.Low, Mid = Total.Mid, High = Total.High)

# make data into table for chi test - observed 
chi.os <- pivot_longer(obs.os, 'Low':'High', names_to = 'Treatment', values_to = 'Observed_nests')
chi.os$Site_Treatment <- with(chi.os, paste(Site, Treatment, sep = '_'))
chi.os <- chi.os %>% ungroup() %>% select('Site_Treatment', 'Observed_nests')
chi.meg <- pivot_longer(obs.meg, 'Low':'High', names_to = 'Treatment', values_to = 'Observed_nests')
chi.meg$Site_Treatment <- with(chi.meg, paste(Site, Treatment, sep = '_'))
chi.meg <- chi.meg %>% ungroup() %>% select('Site_Treatment', 'Observed_nests')

# add columns for the expected nests (doing this the easy way for now >:)
chi.os$Expected_nests <- c(45.72973, 30.48649, 17.783784, 23.35135, 15.56757, 9.081081)
chi.meg$Expected_nests <- c(19.94595, 13.2973, 7.756757)

chisq.test(chi.os$Observed_nests, chi.os$Expected_nests)
fisher.test(chi.meg$Observed_nests, chi.meg$Expected_nests)



```

```{r Chi Square Tables: on-season, echo = FALSE}

### Only on-season data

knitr::kable(exp.os, 
             caption = 'Expected new nests: Osmia',
             col.names = c('Site', 'Low', 'Mid', 'High')) %>% 
  kable_styling()

knitr::kable(obs.os, 
             caption = 'Observed new nests: Osmia',
             col.names = c('Site', 'Low', 'Mid', 'High')) %>% 
  kable_styling()

knitr::kable(exp.meg, 
             caption = 'Expected new nests: Megachile',
             col.names = c('Site', 'Low', 'Mid', 'High')) %>% 
  kable_styling()

knitr::kable(obs.meg, 
             caption = 'Observed new nests: Megachile',
             col.names = c('Site', 'Low', 'Mid', 'High')) %>% 
  kable_styling()

```

------------------------------------------------------------------------

### Parasitism

```{r Parasitism Models}

## Something is off about this. I'll need help...

## Does parasitism affect nesting preference?

mod_par_os <- glmer(Bee_choice ~ scale(Osmia_no_ps_prior) * scale(Par_in_trapnest) + scale(Good_days) + scale(Day_of_year) + Site + (1|Subsite/Trapnest/Cavity_ID), data = nesting_os, family = binomial)

plot(simulateResiduals(mod_par_os))
summary(mod_par_os)

# Plot actual model
ggpredict(mod_par_os, c('Par_in_trapnest')) %>% 
  plot(rawdata = TRUE, jitter = .04) 


```

------------------------------------------------------------------------

## Visualizations

The figures below show if a new nest is built when I visit the trapnests, represented in a binary response in which, *of the cavities available for nesting*, 1 is when a new nest is built and 0 is no new nest is built (i.e., the cavity remains empty).

### Measuring the choice to nest per available cavities.

```{r Figures: Available Cavities, echo=FALSE, out.width='50%', out.height='50%', fig.show='hold', fig.align='side', warning=FALSE, message=FALSE}

## Setting up things for figures


theme_set(theme_classic(base_size=12)) 
# make data frames for each season
os_season <- filter(preferences, Subsite == 'A' | Subsite == 'B' | Subsite == 'C' | Subsite == 'D' | Subsite == 'E' | Subsite == 'F' | Subsite == 'G' | Subsite == 'H' | Subsite == 'I' | Subsite == 'J' | Subsite == 'K' | Subsite == 'L')
meg_season <- filter(preferences, Subsite == 'M' | Subsite == 'N' | Subsite == 'O' | Subsite == 'P' | Subsite == 'Q' | Subsite == 'R' | Subsite == 'S' | Subsite == 'T' | Subsite == 'U' | Subsite == 'V' | Subsite == 'W' | Subsite == 'X')


## New nests per available cavities

# Osmia during Osmia season
os_season %>% 
  filter(Bee_choice != 'NA' & Occupant == 'osmia' | Bee_choice == 0) %>%  
  ggplot(aes(Available_prior, Bee_choice)) + 
  geom_jitter(height = 0.05, alpha = 0.5, aes(colour = Occupant)) +
  stat_smooth(method='glm', method.args=list(family='binomial'), colour = 'grey75') +
  scale_color_manual(values=c('steelblue3', 'plum3')) +
  labs(x='Available cavities', y = 'Presence of Presence of new nest for each visit', title = 'Osmia: available cavities (Osmia season)') +
  theme( # removing the y axes labels
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) + 
  xlim(0, 18)


# Osmia during whole of summer
Osmia_data %>% 
  filter(Bee_choice != 'NA' & Occupant == 'osmia' | Bee_choice == 0) %>%  
  ggplot(aes(Available_prior, Bee_choice)) + 
  geom_jitter(height = 0.05, alpha = 0.5, aes(colour = Occupant)) +
  stat_smooth(method='glm', method.args=list(family='binomial'), colour = 'grey75') +
  scale_color_manual(values=c('steelblue3', 'plum3')) +
  labs(x= 'Available cavities', y = 'Presence of new nest for each visit', title = 'Osmia: available cavities (Full season)') + theme( # removing the y axes labels
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) + 
  xlim(0, 18) 

# Plot actual model
ggpredict(mod_c_cav_os, c('Available_prior')) %>% plot(rawdata = TRUE, jitter = .04) + labs(x = 'Available cavities', y = 'Presence of new nest for each visit', title = 'Osmia: available cavities (full season)')


# Megachile during megachile season
meg_season %>% 
  filter(Bee_choice != 'NA' & Occupant == 'megachile' | Bee_choice == 0) %>% 
  ggplot(aes(Available_prior, Bee_choice)) + 
  geom_jitter(height = 0.05, alpha = 0.6, aes(colour = Occupant)) + 
  stat_smooth(method='glm', method.args=list(family='binomial'), colour = 'grey75') + 
  scale_color_manual(values=c('darkgoldenrod2', 'steelblue3')) +
  labs(x= 'Available cavities', y = 'Presence of new nest for each visit', title = 'Megachile: available cavities (Megachile season)') +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) + 
  xlim(0, 18)
  
# Megachile during whole of summer
Megachile_data %>% 
  filter(Bee_choice != 'NA' & Occupant == 'megachile' | Bee_choice == 0) %>% 
  ggplot(aes(Available_prior, Bee_choice)) + 
  geom_jitter(height = 0.05, alpha = 0.6, aes(colour = Occupant)) + 
  stat_smooth(method='glm', method.args=list(family='binomial'), colour = 'grey75') + 
  scale_color_manual(values=c('darkgoldenrod2', 'steelblue3')) +
  labs(x='Available cavities', y='Presence of new nest for each visit', title='Megachile: available cavities (Full season)') + theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) + 
  xlim(0, 18)
#################### SOMETHING ABOUT THIS ISNT RIGHT, ONY 23 DATA POINTS HERE AND 10 ON MEG SEASON 


# Plot actual model
ggpredict(mod_c_cav_meg, c('Available_prior')) %>% plot(rawdata = TRUE, jitter = .04) + labs(x = 'Available cavities', y = 'Presence of new nest for each visit', title = 'Megachile: available cavities (full season)')



```

------------------------------------------------------------------------

### Measuring the choice to nest per presence of conspecifics.

```{r Figures: Conspecific Presence, echo=FALSE, out.width='50%', out.height='50%', fig.show='hold', fig.align='side', warning=FALSE, message=FALSE}

  
## New nests per con-specific occupancies (with and without pseudo nests)

# Osmia # with pseudo 
Osmia_data %>% 
  filter(Occupant == 'osmia' | Occupant == 'none') %>%  
  ggplot(aes(Osmia_ps_prior, Bee_choice)) + 
  geom_jitter(height = 0.05, alpha = 0.6, aes(colour = Occupant)) + 
  stat_smooth(method='glm', method.args=list(family='binomial'), colour = 'grey75') +
  labs(title = 'Osmia occupancies', x = 'Osmia occupancies (including pseudo)', y = 'Presence of new nest for each visit') +
  scale_color_manual(values=c('steelblue3', 'plum3'))  +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) +
  xlim(0, 18)

# Plot actual model
ggpredict(mod_c_consp_os_ps, c('Osmia_ps_prior')) %>% 
  plot(rawdata = TRUE, jitter = .04) + 
  labs(x = 'Osmia occupancies including pseudo', y = 'Presence of new nest for each visit', title = 'Osmia') +
  xlim(0, 18)


# Osmia # without pseudo 
Osmia_data %>% 
  filter(Occupant == 'osmia' | Occupant == 'none') %>%  
  ggplot(aes(Osmia_no_ps_prior, Bee_choice)) + 
  geom_jitter(height = 0.05, alpha = 0.6, aes(colour = Occupant)) + 
  stat_smooth(method='glm', method.args=list(family='binomial'), colour = 'grey75') +
  labs(title = 'Osmia occupancies', x = 'Osmia occupancies (excluding pseudo)', y = 'Presence of new nest for each visit') +
  scale_color_manual(values=c('steelblue3', 'plum3')) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) +
  xlim(0, 18)

# Plot actual model
ggpredict(mod_c_consp_os_no_ps, c('Osmia_no_ps_prior')) %>% 
  plot(rawdata = TRUE, jitter = .04) + 
  labs(x = 'Osmia occupancies excluding pseudo', y = 'Presence of new nest for each visit', title = 'Osmia') +
  xlim(0, 18)


# Megachile # with pseudo 
Megachile_data %>% 
  filter(Occupant == 'megachile' | Occupant == 'none') %>%  
  ggplot(aes(Meg_ps_prior, Bee_choice)) + 
  geom_jitter(height = 0.05, alpha = 0.6, aes(colour = Occupant)) + 
  stat_smooth(method='glm', method.args=list(family='binomial'), colour = 'pink3', fill = 'lightpink') +
  labs(title = 'Megachile occupancies',  x = 'Megachile occupancies (including pseudo)', y = 'Presence of new nest for each visit') +
  scale_color_manual(values=c('darkgoldenrod2', 'steelblue3')) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) +
  xlim(0, 18)

# Plot actual model
ggpredict(mod_c_consp_meg_ps, c('Meg_ps_prior')) %>% 
  plot(rawdata = TRUE, jitter = .04) + 
  labs(x = 'Megachile occupancies including pseudo', y = 'Presence of new nest for each visit', title = 'Megachile')


# Megachile # without pseudo 
Megachile_data %>% 
  filter(Occupant == 'megachile' | Occupant == 'none') %>%  
  ggplot(aes(Meg_no_ps_prior, Bee_choice)) + 
  geom_jitter(height = 0.05, alpha = 0.6, aes(colour = Occupant)) + 
  stat_smooth(method='glm', method.args=list(family='binomial'), colour = 'grey75') +
  labs(title = 'Megachile occupancies',  x = 'Megachile occupancies (excluding pseudo)', y = 'Presence of new nest for each visit') +
  scale_color_manual(values=c('darkgoldenrod2', 'steelblue3')) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) +
  xlim(0, 18)

# Plot actual model
ggpredict(mod_c_consp_meg_no_ps, c('Meg_no_ps_prior')) %>% 
  plot(rawdata = TRUE, jitter = .04) + 
  labs(x='Megachile occupancies excluding pseudo',y='Presence of new nest for each visit', title='Megachile') + 
  xlim(0, 18)


```

------------------------------------------------------------------------

### Parasitism

```{r Parasitism Visuals}

## Something is off about this. I'll need help...

nesting_os %>% 
  ggplot(aes(Par_in_trapnest, Bee_choice)) + 
  geom_jitter(height = 0.03, alpha = 0.4) 

# each data point is again a trapnest and date combo

# Cumulative number of parasites per osmia occupancies
nesting %>% 
  filter(Occupant == 'osmia' & !is.na(Cumulative_par)) %>% 
  ggplot(aes(Osmia_no_ps_prior, Cumulative_par)) + 
  geom_jitter(height = 0.05, alpha = 0.6, aes(colour = Occupant)) + 
  stat_smooth(method='glm', colour = 'grey75') +
  labs(x = 'Osmia occupancies (excluding pseudo)', y = 'Cumulative number of parasites') 

# Proportion of cells in a nest with parasites per osmia occupancies
nesting %>% 
  filter(Occupant == 'osmia' & !is.na(Prop_cells_par)) %>% 
  ggplot(aes(Osmia_no_ps_prior, Prop_cells_par)) +
  geom_jitter() + 
  stat_smooth(method='glm', method.args=list(family='binomial'), colour = 'grey75') +
  labs(x = 'Osmia occupancies (excluding pseudo)', y = 'Proportion of cells in a nest that are parasitized') 

# Plot actual model
ggpredict(mod_par_os, c('Par_in_trapnest')) %>% 
  plot(rawdata = TRUE, jitter = .04) 


```

------------------------------------------------------------------------

### Nest investement

```{r Investment Models, include=FALSE, echo=TRUE, results='hide'}

library(betareg)

hist(investment_os$Prop_filled)

mod_invest_os <- glmer(Prop_filled ~ Par_in_trapnest * Cumulative_par + Day_of_year + Site + (1|Subsite/Cavity_ID), data = investment_os, family = binomial)

plot(simulateResiduals(mod_invest_os))
summary(mod_invest_os)








#### Make a loop for active period

# # make an index
# index_2 <- (1:length(weather$Date)) # make index of 1 to last row
# weather_i <- weather
# weather_i$index <- index_2 # making index column to count
# weather_i <- weather_i %>% relocate('index') # puts it in the front
# 
# # initialize a table for type of day count
# Day_type <- as.data.frame(matrix(NA,0,2))
# colnames(Day_type)[1] <- 'index'
# colnames(Day_type)[2]<-'Day_type'


# ### START
# for (i in weather_i$index) { 
#   
#   # create Day_type row for our big table
#   Day_type_i <- as.data.frame(matrix(NA,0,2))
#   colnames(Day_type_i)[1] <- 'index'
#   colnames(Day_type_i)[2]<-'Day_type_i'
#   
#   Day_type_i[nrow(Day_type_i)+1,] = c(i, XXX )
#   
#   # bind the row to the big table
#   Day_type <- rbind(Day_type, Day_type_i)
# } ### END

# join things together



# 3) Do bees reduce investment in a nest if parasites are present?
# Proportion of volume filled with cells ~ presence of parasites in that trapnest * if the nest is parasitized or not + parasite species + (1|site/subsite/trapnest)



```

```{r Figures: Nest Investment, echo=FALSE, out.width='50%', out.height='50%', fig.show='hold', fig.align='side', warning=FALSE, message=FALSE}







## Plot number of cells with filled length
investment %>% ggplot(aes(Prop_filled, Number_of_cells, label = Nest_ID)) + 
  geom_jitter() + geom_text(aes(colour = Trapnest), size = 2)






investment$Treatment <- factor(investment$Treatment , levels=c('Low', 'Mid', 'High'))

investment %>% filter(Occupant == 'osmia' | Occupant == 'megachile') %>% 
  ggplot(aes(Treatment, Prop_filled, fill = Occupant)) +
  geom_boxplot(alpha = 0.8) +
  scale_fill_manual(values=c('goldenrod2', 'plum3'), labels = c('Megachile', 'Osmia')) +
  labs(y = 'Proportion of straw filled with cells') + 
  theme(legend.text = element_text(face = 'italic'))



# write.csv(preferences, 'preferences.csv')
# write.csv(nesting, 'nesting.csv')


```

```{r Example code}
 
# Testing to make sure my looping method works

# # loop takes too long to do at once
# # so remove all empty and pseudo, then adding back in after
# # test it with trapnest 12 first
# os_parasites_ex1 <- os_parasites %>% filter(Trapnest== 12) # including empty and pseudo
# os_parasites_ex2 <- os_parasites %>% filter(Trapnest== 12 & Status != 'empty' & Status != 'pseudo') # excluding empty and pseudo
# 
# # Test 1
# for(i in 1:nrow(os_parasites_ex1)){
#   nestID <- os_parasites_ex1$Nest_ID[i] # focal nest for i-th row
#   dt <- os_parasites_ex1$Date[i] # date for i-th row
#   allprev <- os_parasites_ex1 %>% filter(Nest_ID == nestID & Date <= dt) %>% pull(Cumulative_par_max) # all previous parasite counts for that focal nest
#   
#   check1 <- !all(is.na(allprev))
#   check2 <- !os_parasites_ex1$Status[i] %in% c('empty', 'pseudo') 
#   # you'll have to adjust this as needed for the intended behaviour & possible values in your larger dataset
#   
#   if(check1 & check2){
#     os_parasites_ex1$Cumulative_par_V2[i] <- max(allprev, na.rm = T)
#   }
# }
# view(os_parasites_ex1) # looks good
# 
# # Test 2
# for(i in 1:nrow(os_parasites_ex2)){
#   nestID <- os_parasites_ex2$Nest_ID[i] # focal nest for i-th row
#   dt <- os_parasites_ex2$Date[i] # date for i-th row
#   allprev <- os_parasites_ex2 %>% filter(Nest_ID == nestID & Date <= dt) %>% pull(Cumulative_par_max) # all previous parasite counts for that focal nest
#   
#   check1 <- !all(is.na(allprev))
#   check2 <- !os_parasites_ex2$Status[i] %in% c('empty', 'pseudo') 
#   # you'll have to adjust this as needed for the intended behaviour & possible values in your larger dataset
#   
#   if(check1 & check2){
#     os_parasites_ex2$Cumulative_par_V2[i] <- max(allprev, na.rm = T)
#   }
# }
# view(os_parasites_ex2) # looks good
# # bind back together
# os_parasites_ex3 <- os_parasites %>% filter(Trapnest== 12 & Status == 'empty' | Trapnest== 12 & Status == 'pseudo') # the empty and pseudo ones
# # the length of ex1 should = ex2 + ex3 --> 80 + 334 
# os_parasites_ex3 <- rbind(os_parasites_ex2, os_parasites_ex3)
# os_parasites_ex3 <- os_parasites_ex3 %>% arrange(Nest_ID, Date)

```
